### 2013-02-06 doc/PCRB.tex
Inital files and documentation
Contains a simple pendulum example (nonlinear) and a motion example
(linear).

\documentclass{article}

\input{../common/latex_includes.tex}
\usepackage[T1]{fontenc}
\usepackage{palatino}

\usepackage{siunitx}
% \sisetup{%
%   mode = math,
% %  per-mode = symbol,
% %  per-mode=repeated-symbol,
%   per-mode=reciprocal,
%   detect-all
% }
% \sisetup{per-mode=reciprocal}

\title{PCRB}
\author{Kelvin Layton}
\date{\today}

\begin{document}

\section{Model}

The model takes the form
\begin{align}
	\vec x_k &= \vec f(\vec x_{k-1}) + \vec v_k \\
	\vec y_k &= \mat H\vec x_{k-1} + \vec w_k
\end{align}
where $\vec v_k$ and $\vec w_k$ are white, uncorrelated, zero mean Gaussian noise vectors with covariances of $\mat Q$ and $\mat R$, respectively.

\subsection{Pendulum example}

Let $\vec x_k = [a_k\quad b_k]^T$ where $a_k$ is the pendulum angle and $b_k$ is the angular velocity,
\begin{align}
	\vec f(\vec x_{k}) &= \left[\begin{array}{c}
								a_k + b_k\Delta \\ -b_k - g\sin(a_k)\Delta
								\end{array}\right]
\end{align}
The linearisation used for the EKF is
\begin{align}
	\mat F(\vec x_{k}) &= \left[\begin{array}{cc}
					1 & \Delta \\
					-g\cos(a_k) & 1 \\
				\end{array}\right]
\end{align}
The observation matrix is $H=[1\quad 0]$, so that only the first state element is measured.

\section{PCRB}

The PCRB is computed from the Fisher information matrix computed on the joint density $p(\vec y,\vec x)$,
\begin{align}
	\mat J &= \mathbb E\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right]
\end{align}
where the expectation is taken over $\vec y$ and $\vec x$. A recursive formula was developed in \cite{Tichavsky1998} and is based on the transition density,
\begin{align}
	p(\vec x_k|\vec x_{k-1}) &= \mathcal N(\vec f(\vec x_{k-1})),\mat Q) \\
	&= (2\pi)^{-d/2} |\mat Q|^{-1/2} \exp\left(-\frac{1}{2}(\vec x_k - \vec f(\vec x_{k-1}))^T \mat Q^{-1} (\vec x_k - \vec f(\vec x_{k-1})) \right)
\end{align}
The PCRB is calculated using the logarithm of the transition density,
\begin{align}
	\log p(\vec x_k|\vec x_{k-1}) = C -\frac{1}{2}(\vec x_k - \vec f(\vec x_{k-1}))^T \mat Q^{-1} (\vec x_k - \vec f(\vec x_{k-1})) 
\end{align}
The derivatives are
\begin{align}
	\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) &= -(\vec x_k - \vec f(\vec x_{k-1}))^T\mat Q^{-1}  \\
	\nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) &= -\mat Q^{-1} \\
	\nabla_{\vec x_{k-1}}\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) &= \mat F(\vec x_{k-1})\mat Q^{-1} \\
	\nabla_{\vec x_{k-1}}^T \log p(\vec x_k|\vec x_{k-1}) &= (\vec x_k - \vec f(\vec x_{k-1}))^T\mat Q^{-1} \mat F(\vec x_{k-1}) \\
	\nabla_{\vec x_{k-1}}\nabla_{\vec x_{k-1}}^T \log p(\vec x_k|\vec x_{k-1}) &= -\mat F(\vec x_{k-1})^T\mat Q^{-1} \mat F(\vec x_{k-1}) 
\end{align}
Now the terms in the recursive formula for the Fisher information matrix are
\begin{align}
	 \mat U_k &= \mathbb E\left[ \mat F(\vec x_{k-1})^T\mat Q^{-1} \mat F(\vec x_{k-1}) \right ] \label{eqn:pcrb_u} \\
	\mat V_k &= \mathbb E\left[ -\mat F(\vec x_{k-1})^T\mat Q^{-1} \right ] \label{eqn:pcrb_v} \\
	\mat W_k &= \mat H^T\mat R^{-1}\mat H + \mat Q^{-1}
\end{align}

\subsection{Monte Carlo approximation}

The expectations in \eqref{eqn:pcrb_u} and \eqref{eqn:pcrb_v} can be approximated by numerical integration using many trajectory realisations. That is,
\begin{align}
	\mat U_k &\approx \frac{1}{M}\sum_{i=1}^M \mat F(\vec x_{k-1}^{(i)})^T\mat Q^{-1} \mat F(\vec x_{k-1}^{(i)}) \\
	\mat V_k &\approx \frac{1}{M}\sum_{i=1}^M \mat F(\vec x_{k-1}^{(i)})^T\mat Q^{-1} 
\end{align}

\subsection{Pendulum simulations}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.75\textwidth]{figures/pendulum_trajectory}
	\caption{Trajectory of a pendulum and EKF estimates}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.75\textwidth]{figures/pendulum_pcrb}
	\caption{MSE of EKF estimates (averaged over 1000 realisations) compared to the PCRB. The suboptimal EKF is dependent on the accuracy of the linearisation at different angles, an optimal filter would achieve the CRB at all times}
\end{figure}

% \small
% \bibliographystyle{ieeetr}
% \bibliography{../library}

\end{document}

### 2013-02-06 code/example_ekf.m

@@ -1,0 +1,164 @@
+%% Simulate pendulum model and EKF estimates
+%
+close all
+
+% Define time
+%
+delta = 0.01;  % Sampling period
+t=0:delta:4;
+N=length(t);
+
+% Transition model
+%
+f = @(x)model_pendulum(x,delta,'transition');
+Ffunc = @(x)model_pendulum(x,delta,'jacobian');
+
+% Observation matrix (only observe angle)
+%
+H = [1 0];
+
+% Define noise vectors
+%
+sigma_v = 0.01;  % Process noise
+sigma_w = 0.1;   % Measurement noise
+Q = sigma_v^2*[1 0; 0 1];
+R = sigma_w^2 .* eye(size(H,1));
+
+w = mvnrnd(zeros(size(R,1),1),R,N)';
+v = mvnrnd(zeros(size(Q,1),1),Q,N)';
+
+% Starting state
+%
+x0 = [pi/4; 0];
+
+% Calculate state trajectory
+%
+x = zeros(length(x0),N);
+x(:,1)=x0;
+for i=2:N
+    x(:,i) = f(x(:,i-1)) + v(:,i);
+end
+
+% Define observation vector
+%
+z = H*x + w;
+
+% Prior distribution (defined by m0 & P0)
+%
+m0 = x0;
+P0 = diag([(2*pi)^2 (2*pi)^2]);
+
+% Apply EKF filter
+%
+m = extended_kalman_filter(z,Ffunc,H,Q,R,m0,P0);
+
+% Plot trajecotry and EKF estimates
+%
+figure;
+plot(t,x); hold on;
+plot(t,m,'--');
+xlabel('Time (s)');
+ylabel('States');
+grid on;
+legend({'x_1 - angle','x_2 - velocity','x_1 estimate','x_2 estimate'})
+% export_fig('../figures/pendulum_trajectory.pdf')
+
+
+%% Compute the posterior Cramer-Rao bound (PCRB)
+%
+
+M = 100;    % Number of Monte Carlo samples
+
+% Initialise variables
+%
+
+Qinv=inv(Q);
+Rinv=inv(R);
+J=zeros(2,2,N);
+J(:,:,1)=inv(P0);
+pcrb=zeros(2,N);
+
+% Initalise all trajectories
+%
+xk=zeros(length(m0),M);
+for i=1:M
+    xk(:,i)=mvnrnd(m0,P0)';
+end
+
[…]

### 2013-02-06 code/example_kf.m

@@ -1,0 +1,139 @@
+%% Simulate linear motion model and KF estimates
+%
+close all
+
+% Define time
+%
+delta = 0.01;  % Sampling period
+t=0:delta:4;
+N=length(t);
+
+% Transition model (linear)
+%
+F = [1, delta; 0 1];
+f = @(x)(F*x);
+Ffunc = @(x)F;
+
+% Observation matrix (only observe position)
+%
+H = [1 0];
+
+% Define noise vectors
+%
+sigma_v = 0.01;  % Process noise
+sigma_w = 0.1;   % Measurement noise
+Q = sigma_v^2*[1 0; 0 1];
+R = sigma_w^2 .* eye(size(H,1));
+
+w = mvnrnd(zeros(size(R,1),1),R,N)';
+v = mvnrnd(zeros(size(Q,1),1),Q,N)';
+
+% Starting state
+%
+x0 = [3; 0];
+
+% Calculate state trajectory
+%
+x = zeros(length(x0),N);
+x(:,1)=x0;
+for i=2:N
+    x(:,i) = f(x(:,i-1)) + v(:,i);
+end
+
+% Define observation vector
+%
+z = H*x + w;
+
+% Prior distribution (defined by m0 & P0)
+%
+m0 = x0;
+P0 = diag([(2*pi)^2 (2*pi)^2]);
+
+% Apply EKF filter
+%
+m = extended_kalman_filter(z,Ffunc,H,Q,R,m0,P0);
+
+% Plot trajecotry and EKF estimates
+%
+figure;
+plot(t,x); hold on;
+plot(t,m,'--');
+xlabel('Time (s)');
+ylabel('States');
+grid on;
+legend({'x_1 - position','x_2 - velocity','x_1 estimate','x_2 estimate'})
+% export_fig('../figures/motion_trajectory.pdf')
+
+
+%% Compute the posterior Cramer-Rao bound (PCRB)
+%
+
+% Initialise variables
+%
+pcrb=zeros(2,N);
+J=zeros(2,2,N);
+J(:,:,1)=inv(P0);
+
+% Compute the PCRB using closed form expressions
+%
+U = F'*inv(Q)*F;
+V = -F'*inv(Q);
+W = inv(Q) + H'*inv(R)*H;
+
+for k=2:N
+    
+    % Recursively compute the Fisher information matrix
+    %
+    J(:,:,k) = W - V'*inv(J(:,:,k-1) + U)*V;
+
+    % Compute the PCRB at the curret time
[…]

### 2013-02-06 code/extended_kalman_filter.m

@@ -1,0 +1,49 @@
+% EXTENDED_KALMAN_FILTER Implements an extended Kalman filter (or regular 
+% Kalman filter).
+% 
+% Inputs: z - measurements
+%         Ffunc - transition matrix function (a function that takes the
+%                 current state and returns the Jacobian). For a regular
+%                 Kalman fiter use @(x)F, where F is the transition matrix
+%         H - the observation matrix
+%         Q - process covariance
+%         R - measurement covariance
+%         m0 - mean of prior distribution
+%         P0 - covariance of prior distribution
+%
+% Outputs: m - the posterior mean (the estimated state)
+%          P - the posterior covariance
+%
+% Kelvin Layton
+% Jan 2013
+%
+function [m, P] = extended_kalman_filter(z,Ffunc,H,Q,R,m0,P0)
+
+    % Initialise variables
+    %
+    N=length(z);
+    m=zeros([length(m0),N]);
+    P=zeros([size(P0),N]);
+
+    % Run Kalman filter over the given data
+    %
+    m(:,1)=m0;
+    P(:,:,1)=P0;
+    for i=2:N
+        
+        % Get linearisation for EKF (note when Ffunc=@(x)F it returns F)
+        %
+        Fhat=Ffunc(m(:,i-1));
+        
+        % Prediction step
+        %
+        m(:,i) = Fhat*m(:,i-1);
+        P(:,:,i) = Fhat*P(:,:,i-1)*Fhat' + Q;
+
+        % Update step
+        %
+        K = P(:,:,i)*H'*inv((H*P(:,:,i)*H' + R));
+        m(:,i) = m(:,i) + K*(z(:,i)-H*m(:,i));
+        P(:,:,i) = (eye(length(m0))-K*H)*P(:,:,i);
+    end
+end

### 2013-02-06 code/model_pendulum.m

@@ -1,0 +1,24 @@
+% MODEL_PENDULUM This function implements the state space representation of
+% the nonlinear pendulum equations.
+% 
+% Inputs: x - the current state
+%         delta - the time step
+%         mode - 'transition' returns the new state,
+%                'jacobian' returns the Jacobian at the current state.
+%
+% Kelvin Layton
+% Jan 2013
+%
+function [out] = model_pendulum(x, delta, mode)
+
+    if mode(1)=='t';
+        % Nonlinear transition model
+        out = [x(1) + x(2)*delta; ...
+                x(2) - 10*sin(x(1))*delta];
+
+    else
+        % Jacobian
+        out = [1, delta; ...
+                -10*cos(x(1))*delta, 1];
+    end
+end

### 2013-02-06 code/JRint_6thOrder.m
Included Dean's 6th order model

@@ -1,0 +1,41 @@
+function x = JRint_6thOrder(x,parameters)
+
+dt = parameters{1};         % time step
+mu = parameters{2};        % this is just the mean when this function is used in estimation
+
+% get current parameters
+e_0 = parameters{3};        % maximal firing parameter
+v_0 = parameters{4};        % firing threshold
+r = parameters{5};              % sigmoid slope
+
+A = parameters{6};          % excitatory gain
+a = parameters{7};          % excitatory time constant
+B = parameters{8};         % inhibitory gain
+b = parameters{9};         % inhibitory time constant
+
+C1 = parameters{10};        % connectivity parameter - excitatory feedback PSP
+C2 = parameters{11};        % connectivity parameter - excitatory feedback firing
+C3 = parameters{12};        % connectivity parameter - inhibitory feedback PSP
+C4 = parameters{13};        % connectivity parameter - excitatory feedback firing
+
+% get current states
+v_f = x(1);
+z_f = x(2);
+v_p1 = x(3);
+z_p1 = x(4);
+v_p2 = x(5);
+z_p2 = x(6);
+
+v_p = v_p1 - v_p2;                   % the minus is where the inhibition is introduced
+
+f_p = 2*e_0 / (1 + exp(r*(v_0 - v_p)));           % pyramidal firing
+f_e = 2*e_0 / (1 + exp(r*(v_0 - C1*v_f)));       % excitatory feedback firing
+f_i = 2*e_0 / (1 + exp(r*(v_0 - C3*v_f)));       % inhibitory feedback firing
+
+% state update equation - propagate states
+x(1) = v_f + dt*z_f;
+x(2) = z_f + dt*(A*a*f_p - 2*a*z_f - a^2*v_f);
+x(3) = v_p1 + dt*z_p1;
+x(4) = z_p1 + dt*(A*a*(mu + C2*f_e) - 2*a*z_p1 - a^2*v_p1);
+x(5) = v_p2 + dt*z_p2;
+x(6) = z_p2 + dt*(B*b*C4*f_i - 2*b*z_p2 - b^2*v_p2);

### 2013-02-06 code/SetParameters_6thOrder.m

@@ -1,0 +1,52 @@
+
+
+% set parameters for the 6th order JR model
+
+% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+%% this set is for alpha rhythm
+% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+mu = 220;        % this is just the mean when this function is used in estimation
+sigma = 5.74;
+
+
+% get current parameters
+e_0 = 2.5;        % maximal firing parameter
+v_0 = 6;        % firing threshold
+r = 0.56;              % sigmoid slope
+
+A = 3.25;          % excitatory gain
+a = 100;          % excitatory time constant
+B = 22;         % inhibitory gain
+b = 50;         % inhibitory time constant
+
+% connectivity parameters from J and R model
+C = 135; % 270;        % 135 for alpha activity
+C1 = C;        % connectivity parameter - excitatory feedback PSP
+C2 = 0.8*C;        % connectivity parameter - excitatory feedback firing
+C3 = 0.25*C;        % connectivity parameter - inhibitory feedback PSP
+C4 = 0.25*C;        % connectivity parameter - excitatory feedback firing
+
+
+% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+% %% this set is for seizure
+% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+mu = 90;        % this is just the mean when this function is used in estimation
+sigma = 1.74;
+
+
+% get current parameters
+e_0 = 2.5;        % maximal firing parameter
+v_0 = 6;        % firing threshold
+r = 0.56;              % sigmoid slope
+
+A = 8.25;          % excitatory gain
+a = 100;          % excitatory time constant
+B = 22;         % inhibitory gain
+b = 50;         % inhibitory time constant
+
+% connectivity parameters from J and R model
+C = 135; % 270;        % 135 for alpha activity
+C1 = C;        % connectivity parameter - excitatory feedback PSP
+C2 = 0.8*C;        % connectivity parameter - excitatory feedback firing
+C3 = 0.25*C;        % connectivity parameter - inhibitory feedback PSP
+C4 = 0.25*C;        % connectivity parameter - excitatory feedback firing

### 2013-02-06 code/main_6thOrder.m

@@ -1,0 +1,56 @@
+
+% this code repricates the structure of the Voss / Schiff FN simulation /
+% estimation
+
+clear all
+close all
+clc
+
+tic
+
+NStates = 6;                               % in the non-augmented JR model
+
+N = 100000;                                % number of samples
+dT = 0.001;                               % sampling time step (global)
+dt = 1*dT;                                 % integration time step
+nn = fix(dT/dt);                         % (used in for loop for forward modelling) the integration time step can be small that the sampling (fix round towards zero)
+
+t = 0:dt:(N-1)*dt;
+
+% preallocate for speed
+x0 = zeros(NStates,N);              % true state trajectory
+
+% Initial conditions
+x0(:,1) = zeros(NStates,1);                   % initial state
+
+% intial true parameters value
+SetParameters_6thOrder;
+
+parameters = {dt,...
+    mu,...
+    e_0,...
+    v_0,...
+    r,...
+    A,...         % excitatory gain
+    a,...          % excitatory time constant
+    B,...        % inhibitory gain
+    b,...        % inhibitory time constant
+    C1,...       % connectivity parameter - excitatory feedback PSP
+    C2,...       % connectivity parameter - excitatory feedback firing
+    C3,...       % connectivity parameter - inhibitory feedback PSP
+    C4};       % connectivity parameter - excitatory feedback firing
+
+% define input
+e = sqrt(dt)*A*a*sigma*randn(N,1);
+
+% Euler-Maruyama integration
+for n=1:N-1
+    x0(:,n+1) = JRint_6thOrder(x0(:,n), parameters)  + [0; 0; 0; e(n); 0; 0];
+end
+
+C = [0 0 1 0 -1 0];           % observation function
+y = C*x0;
+
+figure
+plot(t,x0')
+plot(t,y);

### 2013-02-07 code/example_ekf.m
Fixed EKF function
Prediction step based on f(x) not linearisation

@@ -11,7 +11,7 @@ N=length(t);
 % Transition model
 %
 f = @(x)model_pendulum(x,delta,'transition');
-Ffunc = @(x)model_pendulum(x,delta,'jacobian');
+F = @(x)model_pendulum(x,delta,'jacobian');
 
 % Observation matrix (only observe angle)
 %
@@ -50,7 +50,7 @@ P0 = diag([(2*pi)^2 (2*pi)^2]);
 
 % Apply EKF filter
 %
-m = extended_kalman_filter(z,Ffunc,H,Q,R,m0,P0);
+m = extended_kalman_filter(z,f,F,H,Q,R,m0,P0);
 
 % Plot trajecotry and EKF estimates
 %
@@ -99,9 +99,9 @@ for k=2:N
 
         % Compute the PCRB terms for the current trajectory realisation
         %
-        F = Ffunc(xk(:,i));
-        U = U + F'*Qinv*F;
-        V = V + -F'*Qinv;
+        Fhat = F(xk(:,i));
+        U = U + Fhat'*Qinv*Fhat;
+        V = V + -Fhat'*Qinv;
     end
     U=U./M;
     V=V./M;
@@ -128,7 +128,7 @@ parfor r=1:num_trials
     %
     v = mvnrnd([0 0]',Q,N)';
     x = zeros(2,N);
-    x(:,1)=mvnrnd(m0,P0)';
+    x(:,1)=x0; %mvnrnd(m0,P0)';
     for i=2:N
         x(:,i) = f(x(:,i-1)) + v(:,i);
     end
@@ -140,7 +140,7 @@ parfor r=1:num_trials
 
     % Apply EKF filter
     %
-    m = extended_kalman_filter(z,Ffunc,H,Q,R,m0,P0);
+    m = extended_kalman_filter(z,f,F,H,Q,R,m0,P0);
 
     % Accumulate the estimation error
     %

### 2013-02-07 code/example_kf.m

@@ -51,7 +51,7 @@ P0 = diag([(2*pi)^2 (2*pi)^2]);
 
 % Apply EKF filter
 %
-m = extended_kalman_filter(z,Ffunc,H,Q,R,m0,P0);
+m = extended_kalman_filter(z,f,Ffunc,H,Q,R,m0,P0);
 
 % Plot trajecotry and EKF estimates
 %
@@ -115,7 +115,7 @@ parfor r=1:num_trials
 
     % Apply EKF filter (really the KF in this case)
     %
-    m = extended_kalman_filter(z,Ffunc,H,Q,R,m0,P0);
+    m = extended_kalman_filter(z,f,Ffunc,H,Q,R,m0,P0);
 
     % Accumulate the estimation error
     %

### 2013-02-07 code/extended_kalman_filter.m

@@ -2,7 +2,9 @@
 % Kalman filter).
 % 
 % Inputs: z - measurements
-%         Ffunc - transition matrix function (a function that takes the
+%         f - transition function. For a regular Kalman fiter use 
+%                 @(x)(F*x), where F is the transition matrix
+%         F - transition matrix function (a function that takes the
 %                 current state and returns the Jacobian). For a regular
 %                 Kalman fiter use @(x)F, where F is the transition matrix
 %         H - the observation matrix
@@ -17,7 +19,7 @@
 % Kelvin Layton
 % Jan 2013
 %
-function [m, P] = extended_kalman_filter(z,Ffunc,H,Q,R,m0,P0)
+function [m, P] = extended_kalman_filter(z,f,F,H,Q,R,m0,P0)
 
     % Initialise variables
     %
@@ -25,19 +27,20 @@ function [m, P] = extended_kalman_filter(z,Ffunc,H,Q,R,m0,P0)
     m=zeros([length(m0),N]);
     P=zeros([size(P0),N]);
 
+        
     % Run Kalman filter over the given data
     %
     m(:,1)=m0;
     P(:,:,1)=P0;
     for i=2:N
         
-        % Get linearisation for EKF (note when Ffunc=@(x)F it returns F)
+        % Get linearisation for EKF (note when F=@(x)Fmat it returns Fmat)
         %
-        Fhat=Ffunc(m(:,i-1));
+        Fhat=F(m(:,i-1));
         
         % Prediction step
         %
-        m(:,i) = Fhat*m(:,i-1);
+        m(:,i) = f(m(:,i-1));
         P(:,:,i) = Fhat*P(:,:,i-1)*Fhat' + Q;
 
         % Update step

### 2013-02-07 code/JRint_6thOrder.m
Included Jacobian, cleaned up code, implement EKF
Added model_JR function to either transition the model or return the
Jacobian, suitable for EKF and PCRLB caclualations

@@ -1,41 +1,0 @@
-function x = JRint_6thOrder(x,parameters)
-
-dt = parameters{1};         % time step
-mu = parameters{2};        % this is just the mean when this function is used in estimation
-
-% get current parameters
-e_0 = parameters{3};        % maximal firing parameter
-v_0 = parameters{4};        % firing threshold
-r = parameters{5};              % sigmoid slope
-
-A = parameters{6};          % excitatory gain
-a = parameters{7};          % excitatory time constant
-B = parameters{8};         % inhibitory gain
-b = parameters{9};         % inhibitory time constant
-
-C1 = parameters{10};        % connectivity parameter - excitatory feedback PSP
-C2 = parameters{11};        % connectivity parameter - excitatory feedback firing
-C3 = parameters{12};        % connectivity parameter - inhibitory feedback PSP
-C4 = parameters{13};        % connectivity parameter - excitatory feedback firing
-
-% get current states
-v_f = x(1);
-z_f = x(2);
-v_p1 = x(3);
-z_p1 = x(4);
-v_p2 = x(5);
-z_p2 = x(6);
-
-v_p = v_p1 - v_p2;                   % the minus is where the inhibition is introduced
-
-f_p = 2*e_0 / (1 + exp(r*(v_0 - v_p)));           % pyramidal firing
-f_e = 2*e_0 / (1 + exp(r*(v_0 - C1*v_f)));       % excitatory feedback firing
-f_i = 2*e_0 / (1 + exp(r*(v_0 - C3*v_f)));       % inhibitory feedback firing
-
-% state update equation - propagate states
-x(1) = v_f + dt*z_f;
-x(2) = z_f + dt*(A*a*f_p - 2*a*z_f - a^2*v_f);
-x(3) = v_p1 + dt*z_p1;
-x(4) = z_p1 + dt*(A*a*(mu + C2*f_e) - 2*a*z_p1 - a^2*v_p1);
-x(5) = v_p2 + dt*z_p2;
-x(6) = z_p2 + dt*(B*b*C4*f_i - 2*b*z_p2 - b^2*v_p2);

### 2013-02-07 code/SetParametersJR.m

@@ -1,0 +1,60 @@
+% Set parameters for the 6th order JR model
+%
+function params = SetParametersJR(mode)
+
+if nargin<1
+    mode = 'alpha';
+end
+
+switch mode
+    case 'alpha'
+        
+        % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+        % this set is for alpha rhythm
+        % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+        params.mu = 220;        % this is just the mean when this function is used in estimation
+        params.sigma = 5.74;
+        
+        
+        % get current parameters
+        params.e_0 = 2.5;        % maximal firing parameter
+        params.v_0 = 6;        % firing threshold
+        params.r = 0.56;              % sigmoid slope
+        
+        params.A = 3.25;          % excitatory gain
+        params.a = 100;          % excitatory time constant
+        params.B = 22;         % inhibitory gain
+        params.b = 50;         % inhibitory time constant
+        
+        % connectivity parameters from J and R model
+        C = 135; % 270;        % 135 for alpha activity
+        params.C1 = C;        % connectivity parameter - excitatory feedback PSP
+        params.C2 = 0.8*C;        % connectivity parameter - excitatory feedback firing
+        params.C3 = 0.25*C;        % connectivity parameter - inhibitory feedback PSP
+        params.C4 = 0.25*C;        % connectivity parameter - excitatory feedback firing
+        
+    case 'seizure'
+        % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+        % this set is for seizure
+        % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+        params.mu = 90;        % this is just the mean when this function is used in estimation
+        params.sigma = 1.74;
+        
+        
+        % get current parameters
+        params.e_0 = 2.5;        % maximal firing parameter
+        params.v_0 = 6;        % firing threshold
+        params.r = 0.56;              % sigmoid slope
+        
+        params.A = 8.25;          % excitatory gain
+        params.a = 100;          % excitatory time constant
+        params.B = 22;         % inhibitory gain
+        params.b = 50;         % inhibitory time constant
+        
+        % connectivity parameters from J and R model
+        C = 135; % 270;        % 135 for alpha activity
[…]

### 2013-02-07 code/SetParameters_6thOrder.m

@@ -1,52 +1,0 @@
-
-
-% set parameters for the 6th order JR model
-
-% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-%% this set is for alpha rhythm
-% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-mu = 220;        % this is just the mean when this function is used in estimation
-sigma = 5.74;
-
-
-% get current parameters
-e_0 = 2.5;        % maximal firing parameter
-v_0 = 6;        % firing threshold
-r = 0.56;              % sigmoid slope
-
-A = 3.25;          % excitatory gain
-a = 100;          % excitatory time constant
-B = 22;         % inhibitory gain
-b = 50;         % inhibitory time constant
-
-% connectivity parameters from J and R model
-C = 135; % 270;        % 135 for alpha activity
-C1 = C;        % connectivity parameter - excitatory feedback PSP
-C2 = 0.8*C;        % connectivity parameter - excitatory feedback firing
-C3 = 0.25*C;        % connectivity parameter - inhibitory feedback PSP
-C4 = 0.25*C;        % connectivity parameter - excitatory feedback firing
-
-
-% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-% %% this set is for seizure
-% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-mu = 90;        % this is just the mean when this function is used in estimation
-sigma = 1.74;
-
-
-% get current parameters
-e_0 = 2.5;        % maximal firing parameter
-v_0 = 6;        % firing threshold
-r = 0.56;              % sigmoid slope
-
-A = 8.25;          % excitatory gain
-a = 100;          % excitatory time constant
-B = 22;         % inhibitory gain
-b = 50;         % inhibitory time constant
-
-% connectivity parameters from J and R model
-C = 135; % 270;        % 135 for alpha activity
-C1 = C;        % connectivity parameter - excitatory feedback PSP
-C2 = 0.8*C;        % connectivity parameter - excitatory feedback firing
-C3 = 0.25*C;        % connectivity parameter - inhibitory feedback PSP
-C4 = 0.25*C;        % connectivity parameter - excitatory feedback firing

### 2013-02-07 code/main_6thOrder.m

@@ -1,56 +1,0 @@
-
-% this code repricates the structure of the Voss / Schiff FN simulation /
-% estimation
-
-clear all
-close all
-clc
-
-tic
-
-NStates = 6;                               % in the non-augmented JR model
-
-N = 100000;                                % number of samples
-dT = 0.001;                               % sampling time step (global)
-dt = 1*dT;                                 % integration time step
-nn = fix(dT/dt);                         % (used in for loop for forward modelling) the integration time step can be small that the sampling (fix round towards zero)
-
-t = 0:dt:(N-1)*dt;
-
-% preallocate for speed
-x0 = zeros(NStates,N);              % true state trajectory
-
-% Initial conditions
-x0(:,1) = zeros(NStates,1);                   % initial state
-
-% intial true parameters value
-SetParameters_6thOrder;
-
-parameters = {dt,...
-    mu,...
-    e_0,...
-    v_0,...
-    r,...
-    A,...         % excitatory gain
-    a,...          % excitatory time constant
-    B,...        % inhibitory gain
-    b,...        % inhibitory time constant
-    C1,...       % connectivity parameter - excitatory feedback PSP
-    C2,...       % connectivity parameter - excitatory feedback firing
-    C3,...       % connectivity parameter - inhibitory feedback PSP
-    C4};       % connectivity parameter - excitatory feedback firing
-
-% define input
-e = sqrt(dt)*A*a*sigma*randn(N,1);
-
-% Euler-Maruyama integration
-for n=1:N-1
-    x0(:,n+1) = JRint_6thOrder(x0(:,n), parameters)  + [0; 0; 0; e(n); 0; 0];
-end
-
-C = [0 0 1 0 -1 0];           % observation function
-y = C*x0;
-
-figure
-plot(t,x0')
-plot(t,y);

### 2013-02-07 code/main_JR.m

@@ -1,0 +1,78 @@
+% This code repricates the structure of the Voss / Schiff FN simulation /
+% estimation
+%
+
+clear all
+close all
+clc
+
+
+N = 1000;             	% number of samples
+dT = 0.001;          	% sampling time step (global)
+dt = 1*dT;            	% integration time step
+nn = fix(dT/dt);      	% (used in for loop for forward modelling) the integration time step can be small that the sampling (fix round towards zero)
+
+t = 0:dt:(N-1)*dt;
+
+% Intial true parameter values
+%
+parameters = SetParametersJR('seizure');
+parameters.dt = dt;
+A=parameters.A;
+a=parameters.a;
+sigma=parameters.sigma;
+
+% Initialise random number generator for repeatability
+%
+rng(0);
+
+% Transition model
+%
+NStates = 6;                           
+f = @(x)model_JR(x,'transition',parameters);
+F = @(x)model_JR(x,'jacobian',parameters);
+
+% Initialise trajectory state
+%
+x0 = zeros(NStates,1);                   % initial state
+x = zeros(NStates,N); 
+x(:,1) = x0;
+
+Q = (sqrt(dt)*A*a*sigma)^2*eye(NStates);
+R = 0.01^2*eye(1);
+
+% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+% Generate trajectory
+% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+% Define input
+%
+e = sqrt(dt)*A*a*sigma*randn(N,1);
+
+% Euler-Maruyama integration
+%
+for n=1:N-1
+    x(:,n+1) = f(x(:,n))  + [0; 0; 0; e(n); 0; 0];
+end
+
+H = [0 0 1 0 -1 0];           % observation function
+y = H*x;
+
+% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+% Run EKF
+% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+% Prior distribution (defined by m0 & P0)
+%
+m0 = x0;
+P0 = F(x0)*Q*F(x0)';
+P0=Q;
+
+% Apply EKF filter
+%
[…]

### 2013-02-07 code/model_JR.m

@@ -1,0 +1,100 @@
+% MODEL_JR This function implements the state space representation of
+% the nonlinear neural mass model equations.
+% 
+% Inputs: x - the current state
+%         params - a structure of parameters
+%         mode - 'transition' returns the new state,
+%                'jacobian' returns the Jacobian at the current state.
+%
+% Jan 2013
+%
+function out = model_JR(x,mode,params)
+
+dt = params.dt;         % time step
+mu = params.mu;        % this is just the mean when this function is used in estimation
+
+% Get current parameters
+%
+e_0 = params.e_0;        % maximal firing parameter
+v_0 = params.v_0;        % firing threshold
+r = params.r;              % sigmoid slope
+
+A = params.A;          % excitatory gain
+a = params.a;          % excitatory time constant
+B = params.B;         % inhibitory gain
+b = params.b;         % inhibitory time constant
+
+C1 = params.C1;        % connectivity parameter - excitatory feedback PSP
+C2 = params.C2;        % connectivity parameter - excitatory feedback firing
+C3 = params.C3;        % connectivity parameter - inhibitory feedback PSP
+C4 = params.C4;        % connectivity parameter - excitatory feedback firing
+
+
+% Linear component of model
+%
+F = [1, dt, 0, 0, 0, 0; ...
+    -a^2*dt, 1-2*a*dt, 0, 0, 0, 0; ...
+    0, 0, 1, dt, 0, 0; ...
+    0, 0, -a^2*dt, 1-2*a*dt, 0, 0; ...
+    0, 0, 0, 0, 1, dt; ...
+    0, 0, 0, 0, -b^2*dt, 1-2*b*dt];
+
+% Sigmoid functions
+%
+f_p = 1 / (1 + exp(r*(v_0 - (x(3)-x(5)))));   % pyramidal firing
+f_e = 1 / (1 + exp(r*(v_0 - C1*x(1))));       % excitatory feedback firing
+f_i = 1 / (1 + exp(r*(v_0 - C3*x(1))));       % inhibitory feedback firing
+
+if mode(1)=='t';
+    
+    % Nonlinear component
+    %
[…]

### 2013-02-07 code/test_jacobian.m

@@ -1,0 +1,50 @@
+
+% This script performs a quick visual test of the calculated jacobian
+% 
+% A single parameter is changed and the output vector is plotted as well as
+% the linear approximation given by the jacobian
+%
+
+dt = 0.001; 
+params = SetParametersJR('alpha');
+params.dt = dt;
+
+NStates = 6;                           
+f = @(x)model_JR(x,'transition',params);
+F = @(x)model_JR(x,'jacobian',params);
+
+rng(0);
+
+p=linspace(-1,1,1000);
+P=length(p);
+
+x=repmat(rand(6,1),1,P);
+y=zeros(NStates,P);
+yhat=zeros(NStates,P);
+
+% Calculate the output vector as the parameter is varied
+%
+perturbIdx=1;
+x(perturbIdx,:)=p;
+for i=1:P
+    y(:,i)=f(x(:,i));
+end
+
+% Calculate the linear approximation using the jacobian about the specified
+% fixed point
+%
+p0=0.06;
+[~,idx] = min(abs(p-p0));
+p0=p(idx);
+x0=x(:,idx);
+for i=1:P
+    yhat(:,i)=f(x0) + F(x0)*(x(:,i)-x0) ;
+end
+
+% Plot elements of vector for different paramter values
+%
+plot(p,y);
+hold on;
+plot(p0,y(:,idx),'o')
+plot(p,yhat,'--')
+xlim(0.2*[-1 1]);

### 2013-02-07 code/test_crlb_recursions.m
Added test of Bergman recursion for PCRLB

@@ -1,0 +1,86 @@
+% This script performs a quick validation of the Bergman recursion as
+% equivalent to the Tichavsky recursion for non-singular Q matrices
+%
+% The advantage of the Bergman recursion is apparent for singular Q
+% matrices
+%
+%
+
+%%
+% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+% Simulate linear motion model and KF estimates
+% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+close all
+
+% Define time
+%
+delta = 0.01;  % Sampling period
+t=0:delta:4;
+N=length(t);
+
+% Transition model (linear)
+%
+F = [1, delta; 0 1];
+f = @(x)(F*x);
+Ffunc = @(x)F;
+
+% Observation matrix (only observe position)
+%
+H = [1 0];
+
+% Define noise vectors
+%
+sigma_v = 0.01;  % Process noise
+sigma_w = 1;   % Measurement noise
+
+Q = sigma_v^2*[1 0; 0 1];   % Non-singular Q - either recursion works
+% Q = sigma_v^2*[1 0; 0 0];   % Singular Q - must use Bergman recursion
+
+R = sigma_w^2 .* eye(size(H,1));
+
+% Prior distribution
+%
+x0 = [3; 0];
+m0 = x0;
+P0 = diag([(pi)^2 (pi)^2]);
+
+%% 
+% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+% Compute the posterior Cramer-Rao bound (PCRB)
+% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+
+% Initialise variables
+%
+pcrb=zeros(2,N);
+pcrb2=zeros(2,N);
+J=zeros(2,2,N);
+P=zeros(2,2,N);
+P(:,:,1)=P0;
+J(:,:,1)=inv(P0);
+
+% Compute the PCRB using closed form expressions
+%
+U = F'*inv(Q)*F;
+V = -F'*inv(Q);
+W = inv(Q) + H'*inv(R)*H;
+
+Rinv = H'*inv(R)*H;
+G = eye(2);
+
+for k=2:N
+    
[…]

### 2013-02-07 code/main_JR.m
Included PCRLB calculation using Monte Carlo approx

@@ -2,10 +2,6 @@
 % estimation
 %
 
-clear all
-close all
-clc
-
 
 N = 1000;             	% number of samples
 dT = 0.001;          	% sampling time step (global)
@@ -16,7 +12,9 @@ t = 0:dt:(N-1)*dt;
 
 % Intial true parameter values
 %
-parameters = SetParametersJR('seizure');
+% mode='alpha';
+mode='seizure';
+parameters = SetParametersJR(mode);
 parameters.dt = dt;
 A=parameters.A;
 a=parameters.a;
@@ -38,8 +36,11 @@ x0 = zeros(NStates,1);                   % initial state
 x = zeros(NStates,N); 
 x(:,1) = x0;
 
-Q = (sqrt(dt)*A*a*sigma)^2*eye(NStates);
-R = 0.01^2*eye(1);
+QnonSingular = (sqrt(dt)*A*a*sigma)^2*eye(NStates);
+Q = zeros(NStates);
+Q(4,4) = (sqrt(dt)*A*a*sigma)^2;
+
+R = 1^2*eye(1);
 
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 % Generate trajectory
@@ -47,16 +48,17 @@ R = 0.01^2*eye(1);
 
 % Define input
 %
-e = sqrt(dt)*A*a*sigma*randn(N,1);
+v = mvnrnd(zeros(NStates,1),Q,N)';
 
 % Euler-Maruyama integration
 %
 for n=1:N-1
-    x(:,n+1) = f(x(:,n))  + [0; 0; 0; e(n); 0; 0];
+    x(:,n+1) = f(x(:,n))  + v(:,n);
 end
 
 H = [0 0 1 0 -1 0];           % observation function
-y = H*x;
+w = mvnrnd(zeros(size(H,1),1),R,N)';
+y = H*x + w;
 
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 % Run EKF
@@ -65,8 +67,8 @@ y = H*x;
 % Prior distribution (defined by m0 & P0)
 %
 m0 = x0;
-P0 = F(x0)*Q*F(x0)';
-P0=Q;
+P0 = F(x0)*QnonSingular*F(x0)';
+
 
 % Apply EKF filter
 %
@@ -76,3 +78,110 @@ figure
[…]

### 2013-02-08 doc/PCRB.tex
Updated CRLB documentation
Included recursion by both Tichavsky and Bergman

@@
 
-\input{../common/latex_includes.tex}
-\usepackage[T1]{fontenc}
-\usepackage{palatino}
-
-\usepackage{siunitx}
-% \sisetup{%
-%   mode = math,
-% %  per-mode = symbol,
-% %  per-mode=repeated-symbol,
-%   per-mode=reciprocal,
-%   detect-all
-% }
-% \sisetup{per-mode=reciprocal}
+\input{latex_includes.tex}
 
 \title{PCRB}
-\author{Kelvin Layton}
+
 \date{\today}
@@
 The model takes the form
+\begin{subequations}
 \begin{align}
@@
 \end{align}
-where $\vec v_k$ and $\vec w_k$ are white, uncorrelated, zero mean Gaussian noise vectors with covariances of $\mat Q$ and $\mat R$, respectively.
+\label{eqn:model_general}
+\end{subequations}
+where $\vec v_k$ and $\vec w_k$ are white, uncorrelated, zero mean Gaussian noise vectors with covariances of $\mat Q$ and $\mat R$, respectively. For this model the transition density is defined as
+\begin{align}
+	p(\vec x_k|\vec x_{k-1}) &= \mathcal N(\vec f(\vec x_{k-1})),\mat Q) \\
+	&= (2\pi)^{-d/2} |\mat Q|^{-1/2} \exp\left(-\frac{1}{2}(\vec x_k - \vec f(\vec x_{k-1}))^T \mat Q^{-1} (\vec x_k - \vec f(\vec x_{k-1})) \right)
+\end{align}
+and the likelihood is
+\begin{align}
+	p(\vec y_k|\vec x_k) &= \mathcal N(\mat H\vec x_k,\mat R) \\
+	&= (2\pi)^{-n/2} |\mat R|^{-1/2} \exp\left(-\frac{1}{2}(\vec y_k - \mat H\vec x_k)^T \mat R^{-1} (\vec y_k - \mat H\vec x_k) \right)
+\end{align}
 
-\subsection{Pendulum example}
+\section{PCRB}
+
+The PCRB is computed from the Fisher information matrix computed on the joint density $p(\vec y,\vec x)$,
+\begin{align}
+	\mat J &= \mathbb E\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right]
+\end{align}
+where the expectation is taken over $\vec y$ and $\vec x$. The PCRB bounds the covariance of an unbiased estimator, $\mat P$,
+\begin{align}
+	\mat P &\ge \mat J^{-1}
+\end{align}
+
+Recursive formulas are based on the logarithm of the likelihood and transition densities, which for the model above are
+\begin{align}
+	\log p(\vec x_k|\vec x_{k-1}) &= C_1 -\frac{1}{2}(\vec x_k - \vec f(\vec x_{k-1}))^T \mat Q^{-1} (\vec x_k - \vec f(\vec x_{k-1})) \\
+	\log p(\vec y_k|\vec x_k) &= C_2 -\frac{1}{2}(\vec y_k - \mat H\vec x_k)^T \mat R^{-1} (\vec y_k - \mat H\vec x_k)
+\end{align}
+
+\subsection{Fisher information recursion (non-singular $Q$)} 
+A recursive formula for the FIM, $\mat J_k$, was developed in \cite{Tichavsky1998}. The recursion proceeds as
+\begin{align}
+	\mat J_k(\vec x_k) &= \mat W_k - \mat V_k^T\left[ \mat J_{k-1}(\vec x_{k-1}) + \mat U_k\right]^{-1}\mat V_k 
+\end{align}
+where
+\begin{align}
+	\mat U_k &= -\mathbb E\left[ \nabla_{\vec x_{k-1}}\nabla_{\vec x_{k-1}}^T \log p(\vec x_k|\vec x_{k-1}) \right] \label{eqn:recursion_U}\\
+	\mat V_k &= -\mathbb E\left[ \nabla_{\vec x_{k-1}}\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) \right] \label{eqn:recursion_V}\\
+	\mat W_k &= -\mathbb E\left[ \nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) \right] - \mathbb E\left[ \nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) \right] \label{eqn:recursion_W}
+\end{align}
+
+The derivatives for the above model are
+\begin{align}
+	\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) &= -(\vec x_k - \vec f(\vec x_{k-1}))^T\mat Q^{-1}  \\
+	\nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) &= -\mat Q^{-1} \\
+	\nabla_{\vec x_{k-1}}\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) &= \mat F(\vec x_{k-1})\mat Q^{-1} \\
+	\nabla_{\vec x_{k-1}}^T \log p(\vec x_k|\vec x_{k-1}) &= (\vec x_k - \vec f(\vec x_{k-1}))^T\mat Q^{-1} \mat F(\vec x_{k-1}) \\
+	\nabla_{\vec x_{k-1}}\nabla_{\vec x_{k-1}}^T \log p(\vec x_k|\vec x_{k-1}) &= -\mat F(\vec x_{k-1})^T\mat Q^{-1} \mat F(\vec x_{k-1}) \\
+	\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) &= (\vec y_k - \mat H\vec x_k)^T\mat R^{-1} \mat H \\
+	\nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) &= -\mat H^T\mat R^{-1} \mat H 
+\end{align}
+Substituting these into \eqref{eqn:recursion_U}--\eqref{eqn:recursion_W} gives,
+\begin{align}
+	 \mat U_k &= \mathbb E\left[ \mat F(\vec x_{k-1})^T\mat Q^{-1} \mat F(\vec x_{k-1}) \right ] \label{eqn:pcrb_u} \\
+	\mat V_k &= \mathbb E\left[ -\mat F(\vec x_{k-1})^T\mat Q^{-1} \right ] \label{eqn:pcrb_v} \\
+	\mat W_k &= \mat H^T\mat R^{-1}\mat H + \mat Q^{-1}
+\end{align}
+
+When noise is only added to a subset of the states each transition, $\mat Q$ is singular and $\mat Q^{-1}$ does not exist. Another method is required.
+
+\subsection{Cramer-Rao bound recursion (singular $Q$)}
+
+An alternative formulation is to recursively compute the bound at each step.
+That is, compute $\mat P_k = \mat J_k^{-1}$ directly rather than $\mat J_k$.
+The formulas suggested in \cite{Bergman2001} are applicable to wider class of models (e.g.
+multiplicative noise), denoted by a general transition function $\tilde{\vec f}(\vec x_k,\vec v_k)$.
+The recursion is
+\begin{align}
+	\mat P_k &= \bar{\mat F}_{k-1}\left(\mat P_{k-1}^{-1} + \bar{\mat R}_{k-1}^{-1}\right)^{-1}\bar{\mat F}_{k-1} + \bar{\mat G}_{k-1}\mat Q_k \bar{\mat G}_{k-1}^{-1}
+\end{align}
+where
+\begin{align}
+	\bar{\mat F}_k &= \mathbb E \left[ \nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
+	\bar{\mat R}_k^{-1} &= -\mathbb E\left[ \nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) \right] \\
+	\bar{\mat G}_k &= \mathbb E \left[ \nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
+	\mat Q_k^{-1} &= -\mathbb E \left[ \nabla_{\vec v_k}\nabla_{\vec v_k}^T \log p(\vec v_k) \right]
+\end{align}
+and $\mat P_0^{-1}$ is the inverse of the prior covariance.
+The model in \eqref{eqn:model_general} is represented by $\tilde{\vec f}(\vec x_k,\vec v_k) = \vec f(\vec x_k) + \vec v_k$.
+This simplifies the derivatives to
+\begin{align}
+	\nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat F_k(\vec x_k) \\
+	\nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat I_{d\times d} \\
+	\nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) &= -\mat H^T\mat R^{-1} \mat H 
+\end{align}
+and the corresponding matrices to	
+\begin{align}
+	\bar{\mat F}_k &= \mathbb E\left[ \mat F_k(\vec x_k)\right] \label{eqn:pcrb_term_F}\\
+	\bar{\mat R}_k^{-1} &= -\mat H^T\mat R^{-1} \mat H \\
+	\bar{\mat G}_k &= \mat I_{d\times d} 
+\end{align}
+Notice that $\mat Q^{-1}$ is not directly required in the recursion, so this method is applicable to neural field models where process noise is only added to a single state.
+
+\subsection{Monte Carlo approximations}
+
+The expectations in \eqref{eqn:pcrb_u} and \eqref{eqn:pcrb_v} or \eqref{eqn:pcrb_term_F} can be approximated by numerical integration using many trajectory realisations. That is,
+\begin{align}
+	\mat U_k &\approx \frac{1}{M}\sum_{i=1}^M \mat F(\vec x_{k-1}^{(i)})^T\mat Q^{-1} \mat F(\vec x_{k-1}^{(i)}) \\
+	\mat V_k &\approx \frac{1}{M}\sum_{i=1}^M \mat F(\vec x_{k-1}^{(i)})^T\mat Q^{-1}  
+\end{align}
+or alternatively,
+\begin{align}
+	\bar{\mat F}_{k-1} &\approx \frac{1}{M}\sum_{i=1}^M \mat F(\vec x_{k-1}^{(i)})
+\end{align}
+
+\section{Pendulum example}
 
@@
 
-\section{PCRB}
-
-The PCRB is computed from the Fisher information matrix computed on the joint density $p(\vec y,\vec x)$,
-\begin{align}
-	\mat J &= \mathbb E\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right]
-\end{align}
-where the expectation is taken over $\vec y$ and $\vec x$. A recursive formula was developed in \cite{Tichavsky1998} and is based on the transition density,
-\begin{align}
-	p(\vec x_k|\vec x_{k-1}) &= \mathcal N(\vec f(\vec x_{k-1})),\mat Q) \\
-	&= (2\pi)^{-d/2} |\mat Q|^{-1/2} \exp\left(-\frac{1}{2}(\vec x_k - \vec f(\vec x_{k-1}))^T \mat Q^{-1} (\vec x_k - \vec f(\vec x_{k-1})) \right)
-\end{align}
-The PCRB is calculated using the logarithm of the transition density,
-\begin{align}
-	\log p(\vec x_k|\vec x_{k-1}) = C -\frac{1}{2}(\vec x_k - \vec f(\vec x_{k-1}))^T \mat Q^{-1} (\vec x_k - \vec f(\vec x_{k-1})) 
-\end{align}
-The derivatives are
-\begin{align}
-	\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) &= -(\vec x_k - \vec f(\vec x_{k-1}))^T\mat Q^{-1}  \\
-	\nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) &= -\mat Q^{-1} \\
-	\nabla_{\vec x_{k-1}}\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) &= \mat F(\vec x_{k-1})\mat Q^{-1} \\
-	\nabla_{\vec x_{k-1}}^T \log p(\vec x_k|\vec x_{k-1}) &= (\vec x_k - \vec f(\vec x_{k-1}))^T\mat Q^{-1} \mat F(\vec x_{k-1}) \\
-	\nabla_{\vec x_{k-1}}\nabla_{\vec x_{k-1}}^T \log p(\vec x_k|\vec x_{k-1}) &= -\mat F(\vec x_{k-1})^T\mat Q^{-1} \mat F(\vec x_{k-1}) 
-\end{align}
-Now the terms in the recursive formula for the Fisher information matrix are
-\begin{align}
-	 \mat U_k &= \mathbb E\left[ \mat F(\vec x_{k-1})^T\mat Q^{-1} \mat F(\vec x_{k-1}) \right ] \label{eqn:pcrb_u} \\
-	\mat V_k &= \mathbb E\left[ -\mat F(\vec x_{k-1})^T\mat Q^{-1} \right ] \label{eqn:pcrb_v} \\
-	\mat W_k &= \mat H^T\mat R^{-1}\mat H + \mat Q^{-1}
-\end{align}
-
-\subsection{Monte Carlo approximation}
-
-The expectations in \eqref{eqn:pcrb_u} and \eqref{eqn:pcrb_v} can be approximated by numerical integration using many trajectory realisations. That is,
-\begin{align}
-	\mat U_k &\approx \frac{1}{M}\sum_{i=1}^M \mat F(\vec x_{k-1}^{(i)})^T\mat Q^{-1} \mat F(\vec x_{k-1}^{(i)}) \\
-	\mat V_k &\approx \frac{1}{M}\sum_{i=1}^M \mat F(\vec x_{k-1}^{(i)})^T\mat Q^{-1} 
-\end{align}
-
-\subsection{Pendulum simulations}
+\subsection{Simulations}
 
@@
 	\centering
-	\includegraphics[width=0.75\textwidth]{figures/pendulum_trajectory}
+	\includegraphics[width=0.75\textwidth]{../figures/pendulum_trajectory}
 	\caption{Trajectory of a pendulum and EKF estimates}
@@
 	\centering
-	\includegraphics[width=0.75\textwidth]{figures/pendulum_pcrb}
-	\caption{MSE of EKF estimates (averaged over 1000 realisations) compared to the PCRB.
-The suboptimal EKF is dependent on the accuracy of the linearisation at different angles, an optimal filter would achieve the CRB at all times}
+	\includegraphics[width=0.75\textwidth]{../figures/pendulum_pcrb}
+	\caption{MSE of EKF estimates (averaged over 1000 realisations) compared to the PCRB. The suboptimal EKF is slightly above the PCRB.}
 \end{figure}
 
-% \small
-% \bibliographystyle{ieeetr}
-% \bibliography{../library}
+\small
+\bibliographystyle{ieeetr}
+\bibliography{references}
+% \bibliography{../../library}
 

### 2013-02-18 code/test_crlb_recursions.m
Add other CRLB recursions

@@ -6,18 +6,19 @@
 %
 %
 
+
 %%
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 % Simulate linear motion model and KF estimates
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
-close all
+% close all
 
 % Define time
 %
 delta = 0.01;  % Sampling period
-t=0:delta:4;
-N=length(t);
+N=30;
+t=(0:N-1)*delta
 
 % Transition model (linear)
 %
@@ -31,7 +32,7 @@ H = [1 0];
 
 % Define noise vectors
 %
-sigma_v = 0.01;  % Process noise
+sigma_v = 1;  % Process noise
 sigma_w = 1;   % Measurement noise
 
 Q = sigma_v^2*[1 0; 0 1];   % Non-singular Q - either recursion works
@@ -41,9 +42,10 @@ R = sigma_w^2 .* eye(size(H,1));
 
 % Prior distribution
 %
-x0 = [3; 0];
+x0 = [3; 3];
 m0 = x0;
-P0 = diag([(pi)^2 (pi)^2]);
+P0 = diag(100*[(pi)^2 (pi)^2]);
+
 
 %% 
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
@@ -56,9 +58,24 @@ P0 = diag([(pi)^2 (pi)^2]);
 pcrb=zeros(2,N);
 pcrb2=zeros(2,N);
 J=zeros(2,2,N);
+J2=zeros(2,2,N);
 P=zeros(2,2,N);
+P2=zeros(2,2,N);
+P3=zeros(2,2,N);
+Pkf=zeros(2,2,N);
 P(:,:,1)=P0;
+P2(:,:,1)=P0;
+P3(:,:,1)=P0;
+Pkf(:,:,1)=P0;
 J(:,:,1)=inv(P0);
+J2(:,:,1)=inv(P0);
+
+pcrb(:,1) = diag(inv(J(:,:,1)));
+pcrb2(:,1) = diag(inv(J2(:,:,1)));
+pcrb3(:,1) = diag(P(:,:,1));
+pcrb4(:,1) = diag(Pkf(:,:,1));
+pcrb5(:,1) = diag(P2(:,:,1));
+pcrb6(:,1) = diag(P3(:,:,1));
 
[…]

### 2013-02-18 code/utils/test_expectation_gaussian.m
Expectation of error function derivative

@@ -1,0 +1,18 @@
+M=1000000;
+
+x0=2;
+P0=4;
+
+v0=1;
+Q0=1;
+
+% Generate samples and propogate them through a Gaussian function. 
+% (Gaussian is the derivative of the error function sigmoid)
+%
+x = mvnrnd(x0,P0,M);
+
+f = 1./sqrt(2*pi*Q0).* exp(-(x-v0).^2./(2*Q0));
+
+mean_sample = mean(f)
+
+mean_theoretical = 1./sqrt(2*pi*(Q0+P0)) .* exp(-(x0-v0).^2./(2*(Q0+P0)))

### 2013-02-20 code/test_crlb_recursions.m
Fixed Bergman iteration
Updated CRLB calcualation to compute *posterior* bound not just a bound
of prior

@@ -6,23 +6,22 @@
 %
 %
 
-
 %%
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 % Simulate linear motion model and KF estimates
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
-% close all
+close all
 
 % Define time
 %
-delta = 0.01;  % Sampling period
-N=30;
-t=(0:N-1)*delta
+delta = 0.1;  % Sampling period
+t=0:delta:4;
+N=length(t);
 
 % Transition model (linear)
 %
-F = [1, delta; 0 1];
+F = [1.5, delta; 0 1.5];
 f = @(x)(F*x);
 Ffunc = @(x)F;
 
@@ -42,10 +41,9 @@ R = sigma_w^2 .* eye(size(H,1));
 
 % Prior distribution
 %
-x0 = [3; 3];
+x0 = [3; 0];
 m0 = x0;
 P0 = diag(100*[(pi)^2 (pi)^2]);
-
 
 %% 
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
@@ -58,24 +56,12 @@ P0 = diag(100*[(pi)^2 (pi)^2]);
 pcrb=zeros(2,N);
 pcrb2=zeros(2,N);
 J=zeros(2,2,N);
-J2=zeros(2,2,N);
-P=zeros(2,2,N);
-P2=zeros(2,2,N);
-P3=zeros(2,2,N);
-Pkf=zeros(2,2,N);
-P(:,:,1)=P0;
-P2(:,:,1)=P0;
-P3(:,:,1)=P0;
-Pkf(:,:,1)=P0;
-J(:,:,1)=inv(P0);
-J2(:,:,1)=inv(P0);
+Pprior=zeros(2,2,N);
+Pprior(:,:,1)=P0;
 
-pcrb(:,1) = diag(inv(J(:,:,1)));
-pcrb2(:,1) = diag(inv(J2(:,:,1)));
-pcrb3(:,1) = diag(P(:,:,1));
-pcrb4(:,1) = diag(Pkf(:,:,1));
-pcrb5(:,1) = diag(P2(:,:,1));
-pcrb6(:,1) = diag(P3(:,:,1));
+Pprior2=zeros(2,2,N);
+Pprior2(:,:,1)=P0;
[…]

### 2013-02-21 code/main_JR.m
Fixed recursion in main JR script

@@ -12,8 +12,8 @@ t = 0:dt:(N-1)*dt;
 
 % Intial true parameter values
 %
-% mode='alpha';
-mode='seizure';
+mode='alpha';
+% mode='seizure';
 parameters = SetParametersJR(mode);
 parameters.dt = dt;
 A=parameters.A;
@@ -36,7 +36,7 @@ x0 = zeros(NStates,1);                   % initial state
 x = zeros(NStates,N); 
 x(:,1) = x0;
 
-QnonSingular = (sqrt(dt)*A*a*sigma)^2*eye(NStates);
+
 Q = zeros(NStates);
 Q(4,4) = (sqrt(dt)*A*a*sigma)^2;
 
@@ -67,7 +67,7 @@ y = H*x + w;
 % Prior distribution (defined by m0 & P0)
 %
 m0 = x0;
-P0 = F(x0)*QnonSingular*F(x0)';
+P0 = 100.^2*eye(NStates);
 
 
 % Apply EKF filter
@@ -75,8 +75,8 @@ P0 = F(x0)*QnonSingular*F(x0)';
 m = extended_kalman_filter(y,f,F,H,Q,R,m0,P0);
 
 figure
-plot(t,x([1 3 5],:)'); hold on;
-plot(t,m([1 3 5],:)','--');
+plot(t,x([3 5],:)'); hold on;
+plot(t,m([3 5],:)','--');
 
 
 
@@ -96,10 +96,7 @@ pcrb=zeros(NStates,N);
 
 % Initalise all trajectories
 %
-xk=zeros(NStates,M);
-for i=1:M
-    xk(:,i)=mvnrnd(m0,P0)';
-end
+xk=mvnrnd(m0,P0,M)';
 
 % Closed form terms
 %
@@ -109,7 +106,7 @@ Rinv = H'*inv(R)*H;
 % Compute the PCRB using a Monte Carlo approximation
 %
 for k=2:N
-    Fhat = zeros(NStates,NStates);
+    Fhat = zeros(NStates);
     
     v = mvnrnd(zeros(NStates,1),Q,M)';
     parfor i=1:M
@@ -126,13 +123,12 @@ for k=2:N
         
     % Recursively compute the Fisher information matrix
     %
-    P(:,:,k) = Fhat*inv(inv(P(:,:,k-1)) + Rinv)*Fhat' + Q;
[…]

### 2013-02-21 code/model_pendulum.m
Tweaked pendulum simulations to generate oscillations in the CRLB

@@ -14,11 +14,11 @@ function [out] = model_pendulum(x, delta, mode)
     if mode(1)=='t';
         % Nonlinear transition model
         out = [x(1) + x(2)*delta; ...
-                x(2) - 10*sin(x(1))*delta];
+                x(2) - 100*sin(x(1))*delta];
 
     else
         % Jacobian
         out = [1, delta; ...
-                -10*cos(x(1))*delta, 1];
+                -100*cos(x(1))*delta, 1];
     end
 end

### 2013-02-21 doc/PCRB.tex
Fix minus sign typo
Fixed Bergman iteration
Updated CRLB calcualation to compute *posterior* bound not just a bound
of prior
Added discussion of sigmoid width
Also removed latex_includes.tex and fixed up PCRB document

@@
 
-\input{latex_includes.tex}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{graphicx}
+\usepackage{color}
+\usepackage{subfig}
+\usepackage{float} % For [H] placement
+
+\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
+
+\usepackage{booktabs}
+
+\usepackage[bookmarks=true,ocgcolorlinks=true,plainpages=false, breaklinks=true, bookmarksopen=true, bookmarksnumbered=true]{hyperref}
+%\hypersetup{bookmarks=false}  %hide the bookmarks bar
+%\hypersetup{bookmarksopen=false}  % expand tree of bookmarks or just show first level
+\hypersetup{linkcolor=blue, citecolor=magenta,urlcolor=blue} % electronic
+\hypersetup{colorlinks=true}
+
+\usepackage[T1]{fontenc}
+\usepackage{palatino}
+
+\allowdisplaybreaks
+
+\renewcommand{\vec}[1]{\ensuremath{{\boldsymbol #1}}}
+\newcommand{\mat}[1]{\ensuremath{\boldsymbol{#1}}}
 
@@
 \begin{subequations}
+	\label{eqn:model_general}
 \begin{align}
-	\vec x_k &= \vec f(\vec x_{k-1}) + \vec v_k \\
+	\vec x_k &= \vec f_k(\vec x_{k-1}) + \vec v_k \\
 	\vec y_k &= \mat H\vec x_{k-1} + \vec w_k
 \end{align}
-\label{eqn:model_general}
 \end{subequations}
@@
 
-The PCRB is computed from the Fisher information matrix computed on the joint density $p(\vec y,\vec x)$,
+The posterior Cram\'er-Rao bound (PCRB) is computed from the Fisher information matrix computed on the joint density $p(\vec y,\vec x)$,
 \begin{align}
@@
 
-When noise is only added to a subset of the states each transition, $\mat Q$ is singular and $\mat Q^{-1}$ does not exist. Another method is required.
+This method is not applicable when noise is only added to a subset of the states each transition. In this case, $\mat Q$ is singular and $\mat Q^{-1}$ does not exist. Another method is required.
 
@@
 
-An alternative formulation is to recursively compute the bound at each step.
-That is, compute $\mat P_k = \mat J_k^{-1}$ directly rather than $\mat J_k$.
+An alternative formulation is to recursively compute the bound instead of the matrix, i.e., $\mat P_k = \mat J_k^{-1}$ directly rather than $\mat J_k$.
 The formulas suggested in \cite{Bergman2001} are applicable to wider class of models (e.g.
@@
 The recursion is
+\begin{subequations}
 \begin{align}
-	\mat P_k &= \bar{\mat F}_{k-1}\left(\mat P_{k-1}^{-1} + \bar{\mat R}_{k-1}^{-1}\right)^{-1}\bar{\mat F}_{k-1} + \bar{\mat G}_{k-1}\mat Q_k \bar{\mat G}_{k-1}^{-1}
+	\mat P_{k|k-1} &= \bar{\mat F}_k\mat P_{k-1|k-1}\bar{\mat F}_k + \bar{\mat G}_k\mat Q_k \bar{\mat G}_k^{-1} \\
+		\mat P_{k|k} &= \left(\mat P_{k|k-1}^{-1} + \bar{\mat R}_{k-1}^{-1}\right)^{-1}
 \end{align}
+\end{subequations}
 where
@@
 	\bar{\mat F}_k &= \mathbb E\left[ \mat F_k(\vec x_k)\right] \label{eqn:pcrb_term_F}\\
-	\bar{\mat R}_k^{-1} &= -\mat H^T\mat R^{-1} \mat H \\
+	\bar{\mat R}_k^{-1} &= \mat H^T\mat R^{-1} \mat H \\
 	\bar{\mat G}_k &= \mat I_{d\times d} 
@@
 \end{align}
+
+\subsection{Analytical evaluation of expectations}
+
+For the JR model the expectation in \eqref{eqn:pcrb_term_F} above is
+\begin{align}
+	\bar{\mat F}_k &= \mathbb E\left[ \mat F_k(\vec x_k)\right] \\
+	&= \mat F_k^{\rm (lin)} + \mathbb E\left[ \mat G_k(\vec x_k)\right] 
+\end{align}
+where $\mat G_k(\vec x_k)$ contains the nonlinear component of the Jacobian. These nonlinear terms are derivatives of the error function and have the form,
+\begin{align}
+	\exp(-(x-v_0)^2/\zeta^2)
+\end{align}
+The expectation of such terms can only be evaluated when we know the prior mean and covariance at all times.
 

### 2013-02-28 code/compute_pcrb_J.m
Created function for CRB calculation

@@ -1,0 +1,81 @@
+% COMPUTE_PCRB_J Compute the Posterior CRB using Tichavsky iteration
+% 
+% Inputs: t - vector of time points
+%         f - transition function. For a regular Kalman fiter use 
+%                 @(x)(F*x), where F is the transition matrix
+%         F - transition matrix function (a function that takes the
+%                 current state and returns the Jacobian). For a regular
+%                 Kalman fiter use @(x)F, where F is the transition matrix
+%         H - observation matrix function. For linear measurement 
+%                 use @(x)H, where H is observation matrix
+%         Q - process covariance
+%         R - measurement covariance
+%         m0 - mean of prior distribution
+%         P0 - covariance of prior distribution
+%         M - number of Monte Carlo samples
+%
+% Outputs: pcrb - the posterior CRB
+%
+% Kelvin Layton
+% Feb 2013
+%
+function pcrb = compute_pcrb_J(t,f,F,H,Q,R,m0,P0,M)
+
+N = length(t);
+Nstates=length(m0);
+
+% Initialise variables
+%
+
+J=zeros(Nstates,Nstates,N);
+J(:,:,1)=inv(P0);
+pcrb=zeros(Nstates,N);
+
+% Initalise all trajectories
+%
+xk=mvnrnd(m0,P0,M)';
+
+
+Rinv=inv(R);
+Qinv=inv(Q);
+
+
+% Compute the PCRB using a Monte Carlo approximation
+%
+for k=2:N
+    U = zeros(Nstates);
+    V = zeros(Nstates);
+    W = zeros(Nstates);
+    
+    v = mvnrnd(zeros(Nstates,1),Q,M)';
+    parfor i=1:M
+        % Linearise transition function
+        %
+        Fhat = F(xk(:,i));
+        
+        % Sample the next time point for the current trajectory realisation
+        %
+        xk(:,i) = f(xk(:,i)) + v(:,i);
+
+        % Linearise measurment function
+        %
+        Hhat = H(xk(:,i));
+        
+        U = U + Fhat'*Qinv*Fhat;
+        V = V - Fhat'*Qinv;
+        W = W + Qinv + Hhat'*Rinv*Hhat;
+
+    end
+    U=U./M;
+    V=V./M;
+    W=W./M;
+
+    % Recursively compute the Fisher information matrix
+    %
+    J(:,:,k) = W - V'*inv(J(:,:,k-1) + U)*V;
+    
[…]

### 2013-02-28 code/example_ekf.m

@@ -5,7 +5,7 @@ close all
 % Define time
 %
 delta = 0.001;  % Sampling period
-N=4000;
+N=400;
 t=(0:N-1)*delta;
 
 % Transition model
@@ -17,12 +17,10 @@ F = @(x)model_pendulum(x,delta,'jacobian');
 %
 H = [1 0];
 
-% Define noise vectors
+% Define noise covariances
 %
-sigma_v = 0.001;  % Process noise
-sigma_w = 0.1;   % Measurement noise
-Q = sigma_v^2*[1 0; 0 1];
-R = sigma_w^2 .* eye(size(H,1));
+Q = (0.01)^2*[1 0; 0 1];         % Process noise
+R = (0.1)^2 .* eye(size(H,1));    % Measurement noise
 
 w = mvnrnd(zeros(size(R,1),1),R,N)';
 v = mvnrnd(zeros(size(Q,1),1),Q,N)';
@@ -69,63 +67,12 @@ legend({'x_1 - angle','x_2 - velocity','x_1 estimate','x_2 estimate'})
 
 M = 100;    % Number of Monte Carlo samples
 
-% Initialise variables
-%
-
-Qinv=inv(Q);
-Rinv=inv(R);
-J=zeros(2,2,N);
-J(:,:,1)=inv(P0);
-pcrb=zeros(2,N);
-fnorm=zeros(N,1);
-
-% Initalise all trajectories
-%
-% xk=mvnrnd(m0,P0,M)';
-xk=repmat(x0,1,M);
-
-% Compute the PCRB using a Monte Carlo approximation
-%
-for k=2:N
-    U = zeros(2,2);
-    V = zeros(2,2);
-    
-    v = mvnrnd([0 0]',Q,M)';
-    Fnorm=0;
-    parfor i=1:M
-        % Sample the next time point for the current trajectory realisation
-        %
-        xk(:,i) = f(xk(:,i)) + v(:,i);
-
-        % Compute the PCRB terms for the current trajectory realisation
-        %
-        Fhat = F(xk(:,i));
-        U = U + Fhat'*Qinv*Fhat;
-        V = V + -Fhat'*Qinv;
-        
-        Fnorm = Fnorm + norm(Fhat);
-    end
-    U=U./M;
-    V=V./M;
-    W=Qinv + H'*Rinv*H;
-    
[…]

### 2013-03-01 doc/Sigmoid_Functions.tex

--- 2013-02-06 doc/PCRB.tex
@@
 
-\input{../common/latex_includes.tex}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{graphicx}
+\usepackage{color}
+\usepackage{subfig}
+\usepackage{float} % For [H] placement
+
+\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
+
+\usepackage{booktabs}
+
+\usepackage[bookmarks=true,ocgcolorlinks=true,plainpages=false, breaklinks=true, bookmarksopen=true, bookmarksnumbered=true]{hyperref}
+%\hypersetup{bookmarks=false}  %hide the bookmarks bar
+%\hypersetup{bookmarksopen=false}  % expand tree of bookmarks or just show first level
+\hypersetup{linkcolor=blue, citecolor=magenta,urlcolor=blue} % electronic
+\hypersetup{colorlinks=true}
+
 \usepackage[T1]{fontenc}
@@
 
-\usepackage{siunitx}
-% \sisetup{%
-%   mode = math,
-% %  per-mode = symbol,
-% %  per-mode=repeated-symbol,
-%   per-mode=reciprocal,
-%   detect-all
-% }
-% \sisetup{per-mode=reciprocal}
+\allowdisplaybreaks
+
+\renewcommand{\vec}[1]{\ensuremath{{\boldsymbol #1}}}
+\newcommand{\mat}[1]{\ensuremath{\boldsymbol{#1}}}
+
+\DeclareMathOperator*{\erf}{erf}
 
 \title{PCRB}
-\author{Kelvin Layton}
+
 \date{\today}
@@
 
-\section{Model}
+\section{Sigmoid functions}
 
-The model takes the form
+The sigmoid can be specified using the logistic function
 \begin{align}
-	\vec x_k &= \vec f(\vec x_{k-1}) + \vec v_k \\
-	\vec y_k &= \mat H\vec x_{k-1} + \vec w_k
+	g(x) = \frac{1}{1 + \exp(\varsigma(v_0 - x))}
 \end{align}
-where $\vec v_k$ and $\vec w_k$ are white, uncorrelated, zero mean Gaussian noise vectors with covariances of $\mat Q$ and $\mat R$, respectively.
-
-\subsection{Pendulum example}
-
-Let $\vec x_k = [a_k\quad b_k]^T$ where $a_k$ is the pendulum angle and $b_k$ is the angular velocity,
+which has the derivative
 \begin{align}
-	\vec f(\vec x_{k}) &= \left[\begin{array}{c}
-								a_k + b_k\Delta \\ -b_k - g\sin(a_k)\Delta
-								\end{array}\right]
-\end{align}
-The linearisation used for the EKF is
-\begin{align}
-	\mat F(\vec x_{k}) &= \left[\begin{array}{cc}
-					1 & \Delta \\
-					-g\cos(a_k) & 1 \\
-				\end{array}\right]
-\end{align}
-The observation matrix is $H=[1\quad 0]$, so that only the first state element is measured.
-
-\section{PCRB}
-
-The PCRB is computed from the Fisher information matrix computed on the joint density $p(\vec y,\vec x)$,
-\begin{align}
-	\mat J &= \mathbb E\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right]
-\end{align}
-where the expectation is taken over $\vec y$ and $\vec x$. A recursive formula was developed in \cite{Tichavsky1998} and is based on the transition density,
-\begin{align}
-	p(\vec x_k|\vec x_{k-1}) &= \mathcal N(\vec f(\vec x_{k-1})),\mat Q) \\
-	&= (2\pi)^{-d/2} |\mat Q|^{-1/2} \exp\left(-\frac{1}{2}(\vec x_k - \vec f(\vec x_{k-1}))^T \mat Q^{-1} (\vec x_k - \vec f(\vec x_{k-1})) \right)
-\end{align}
-The PCRB is calculated using the logarithm of the transition density,
-\begin{align}
-	\log p(\vec x_k|\vec x_{k-1}) = C -\frac{1}{2}(\vec x_k - \vec f(\vec x_{k-1}))^T \mat Q^{-1} (\vec x_k - \vec f(\vec x_{k-1})) 
-\end{align}
-The derivatives are
-\begin{align}
-	\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) &= -(\vec x_k - \vec f(\vec x_{k-1}))^T\mat Q^{-1}  \\
-	\nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) &= -\mat Q^{-1} \\
-	\nabla_{\vec x_{k-1}}\nabla_{\vec x_k}^T \log p(\vec x_k|\vec x_{k-1}) &= \mat F(\vec x_{k-1})\mat Q^{-1} \\
-	\nabla_{\vec x_{k-1}}^T \log p(\vec x_k|\vec x_{k-1}) &= (\vec x_k - \vec f(\vec x_{k-1}))^T\mat Q^{-1} \mat F(\vec x_{k-1}) \\
-	\nabla_{\vec x_{k-1}}\nabla_{\vec x_{k-1}}^T \log p(\vec x_k|\vec x_{k-1}) &= -\mat F(\vec x_{k-1})^T\mat Q^{-1} \mat F(\vec x_{k-1}) 
-\end{align}
-Now the terms in the recursive formula for the Fisher information matrix are
-\begin{align}
-	 \mat U_k &= \mathbb E\left[ \mat F(\vec x_{k-1})^T\mat Q^{-1} \mat F(\vec x_{k-1}) \right ] \label{eqn:pcrb_u} \\
-	\mat V_k &= \mathbb E\left[ -\mat F(\vec x_{k-1})^T\mat Q^{-1} \right ] \label{eqn:pcrb_v} \\
-	\mat W_k &= \mat H^T\mat R^{-1}\mat H + \mat Q^{-1}
+	g'(x) = \varsigma g(x)(1 - g(x))
 \end{align}
 
-\subsection{Monte Carlo approximation}
-
-The expectations in \eqref{eqn:pcrb_u} and \eqref{eqn:pcrb_v} can be approximated by numerical integration using many trajectory realisations. That is,
+Alternatively, using the error function
 \begin{align}
-	\mat U_k &\approx \frac{1}{M}\sum_{i=1}^M \mat F(\vec x_{k-1}^{(i)})^T\mat Q^{-1} \mat F(\vec x_{k-1}^{(i)}) \\
-	\mat V_k &\approx \frac{1}{M}\sum_{i=1}^M \mat F(\vec x_{k-1}^{(i)})^T\mat Q^{-1} 
+	g(x) = \frac{1}{2}\left( 1 + \erf\left( \frac{x - v_0}{\xi\sqrt{2}}\right)\right)
+\end{align}
+which has the derivative
+\begin{align}
+	g'(x) = \frac{1}{\xi\sqrt{2\pi}} \exp\left(-\left(\frac{x - v_0}{\xi\sqrt{2}}\right)^2 \right) 
 \end{align}
 
-\subsection{Pendulum simulations}
+\section{Relationship between widths}
 
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.75\textwidth]{figures/pendulum_trajectory}
-	\caption{Trajectory of a pendulum and EKF estimates}
-\end{figure}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.75\textwidth]{figures/pendulum_pcrb}
-	\caption{MSE of EKF estimates (averaged over 1000 realisations) compared to the PCRB.
-The suboptimal EKF is dependent on the accuracy of the linearisation at different angles, an optimal filter would achieve the CRB at all times}
-\end{figure}
+Through numerical simulation we can deduce that setting the error function width to
+\begin{align}
+	\xi = \frac{1.699}{\varsigma}
+\end{align}
+gives a sigmoid with approximately the same shape as a logistic function with width $\varsigma$.
+This relationship seems to hold across a wide range of widths.
+It is likely this number could be derived analytically.
 
-% \small
-% \bibliographystyle{ieeetr}
-% \bibliography{../library}
+% \bibliography{../../library}
 

### 2013-03-01 manuscript/manuscript.tex

--- 2013-03-01 doc/Sigmoid_Functions.tex
@@
 
-\DeclareMathOperator*{\erf}{erf}
-
-\title{PCRB}
-
+\title{Statistical performance analysis of neural mass models?}
 \date{\today}
@@
 
-\section{Sigmoid functions}
+\section{Introduction}
 
-The sigmoid can be specified using the logistic function
-\begin{align}
-	g(x) = \frac{1}{1 + \exp(\varsigma(v_0 - x))}
-\end{align}
-which has the derivative
-\begin{align}
-	g'(x) = \varsigma g(x)(1 - g(x))
-\end{align}
+History of models and estimation attempts
 
-Alternatively, using the error function
-\begin{align}
-	g(x) = \frac{1}{2}\left( 1 + \erf\left( \frac{x - v_0}{\xi\sqrt{2}}\right)\right)
-\end{align}
-which has the derivative
-\begin{align}
-	g'(x) = \frac{1}{\xi\sqrt{2\pi}} \exp\left(-\left(\frac{x - v_0}{\xi\sqrt{2}}\right)^2 \right) 
-\end{align}
+Usefulness of bound: benchmark algorithms, experimental design, insight about inherent uncertainty -- which do we address
 
-\section{Relationship between widths}
+\section{Background}
 
-Through numerical simulation we can deduce that setting the error function width to
-\begin{align}
-	\xi = \frac{1.699}{\varsigma}
-\end{align}
-gives a sigmoid with approximately the same shape as a logistic function with width $\varsigma$.
-This relationship seems to hold across a wide range of widths.
-It is likely this number could be derived analytically.
+Present 3 models?
 
-% \bibliography{../../library}
+\section{Theory / Performance Bounds}
+
+Introduce recursive computation of bound and Monte Carlo approximation
+
+\section{Methods?}
+
+\section{Results}
+
+\subsection{Effect of increasing model complexity -- bound should increase}
+
+e.g. Bound for model A > model B > model C
+
+\subsection{Effect of parameters changing -- bound should change}
+
+e.g. seizure easier to estimate than alpha rhythms
+
+\subsection{Effect of measurement noise? -- pretty boring}
+
+e.g. scalp EEG vs iEEG?
+
+\subsection{Augmented models (joint estimation of parameters and states)?}
+
+\section{Discussion}
+
+extensions to other applications e.g. experimental design
+
+contextualize work w.r.t. existing literature?
+
+\section{Conclusion}
+
+% \small
+% \bibliographystyle{ieeetr}
+% \bibliography{references}
 

### 2013-04-11 doc/PCRB.tex
Latest intermediate

@@
 \begin{align}
-	\vec x_k &= \vec f_k(\vec x_{k-1}) + \vec v_k \\
-	\vec y_k &= \mat H\vec x_{k-1} + \vec w_k
+	\vec x_k &= \vec f_{k-1}(\vec x_{k-1}) + \vec v_k \\
+	\vec y_k &= \mat H\vec x_k + \vec w_k
 \end{align}
@@
 
+\section{Classical CRB}
+
+The Fisher information matrix is
+\begin{align}
+	\mat J_{\rm F}(\vec x) &= \mathbb E_{\vec y| \vec x}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y|\vec x) \right]
+\end{align}
+Notice only the likelihood is used since no prior statistics exist for the unknown parameters/state $\vec x$.
+
 \section{PCRB}
@@
 \begin{align}
-	\mat J &= \mathbb E\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right]
+	\mat J &= \mathbb E_{x,y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right] \\
+	&= \mathbb E_{x,y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y|\vec x) \right] + \mathbb E_{x}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec x) \right]
 \end{align}
-where the expectation is taken over $\vec y$ and $\vec x$. The PCRB bounds the covariance of an unbiased estimator, $\mat P$,
+where the expectation is taken over $\vec y$ and $\vec x$ in the first term and $\vec x$ for the second term. The PCRB bounds the covariance of an unbiased estimator, $\mat P$,
 \begin{align}
@@
 \end{align}
+Therefore the final recursion is
+\begin{align}
+	\mat J_{k|k} = \left( \mat Q + \bar{\mat F}_{k-1} \mat J_{k-1}^{-1} \bar{\mat F}_{k-1}^T\right)^{-1} + \mat H_k^T \mat R^{-1} \mat H_k
+\end{align}
+
 Notice that $\mat Q^{-1}$ is not directly required in the recursion, so this method is applicable to neural field models where process noise is only added to a single state.
@@
 
+\section{Hybrid CRB}
+
+In this formulation the unknown vector, $\vec x$ is composed of random unknown elements, $\vec z$, and deterministic unknown elements, $\vec \theta$. The hybrid Fisher information matrix is
+\begin{align}
+	\mat J &= \mathbb E_{y,z|\theta}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec z|\vec\theta) \right] \\
+	&= \mathbb E_{y,z|\theta}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y|\vec z,\vec\theta) \right] + \mathbb E_{z|\theta}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec z|\vec\theta) \right] \\
+	&= \mathbb E_{z|\theta}\left[ \mat J_{\rm F}(\vec x,\vec\theta) \right] + \mathbb E_{z|\theta}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec z|\vec\theta) \right]
+\end{align}
+where $\mat J_{\rm F}$ is the classical Fisher information matrix for deterministic parameters.
+Although we may place prior statistics on the unknown parameters, $\vec\theta$, for estimation purposes, that framework would not allow us to analyse the estimation uncertainty of particular parameter combinations.
+
+\subsection{Summary}
+
+\begin{table}[H]
+\caption{Different bound formulations}
+\label{tbl:bound_types}
+\centering
+\begin{tabular}{p{0.2\textwidth}p{0.4\textwidth}p{0.4\textwidth}}
+  \toprule
+  Bound & Details & Output  \\
+  \midrule
+Bayesian CRB & Features \begin{itemize}
+     \item Prior on $z$ 
+	\item Prior on $\theta$
+	\item Singular Q
+  \end{itemize} & Bound on $z$ and $\theta$ for different $R$, Number of samples, models, and uncertainty in unknown parameters (some notion of robustness to parameter estimates?)  \\
+Hybrid CRB & Features \begin{itemize}
+     \item Prior on $z$ 
+	\item Deterministic $\theta$
+	\item Singular Q
+  \end{itemize} & Bound on $z$ and $\theta$ for specific $\theta$  \\
+  \bottomrule
+\end{tabular}
+\end{table}
+
+\section{Hybrid example}
+
+\todo Need to derive recursion for hybrid state vector
+
+\section{Simple hybrid example}
+
+Consider the simple model $y_i = z t_i + \theta + w_i$ where $w_i$ is noise
+\begin{align}
+	\vec y = \left[\begin{array}{cc}
+	 				 1 & t_1 \\ \vdots & \vdots \\ 1 & t_n 
+					\end{array}\right] \left[\begin{array}{c}
+					 				 \theta \\ z
+									\end{array}\right] 	+ \vec w
+\end{align}
+Here $z$ is random parameter with prior $p_0 = \mathcal N(z_0,P_0)$, $\theta$ is deterministic unknown parameter, $w_i$ is random noise with distribution $\mathcal N(0,R)$.
+
+The information matrix is
+\begin{align}
+	J = J_{\rm D} + J_{\rm P}
+\end{align}
+where
+\begin{align}
+	J_{\rm P} = \left[\begin{array}{cc}
+	 				 0 & 0 \\ 0 & P_0^{-1} 
+					\end{array}\right]
+\end{align}
+The likelihood is
+\begin{align}
+	\log p(\vec y|\vec x) &= c_1 -\frac{1}{2}(\vec y - \mat H\vec x)^T \mat R^{-1} (\vec y - \mat H\vec x) \\
+-\nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y|\vec x) &= \mat H^T\mat R^{-1} \mat H \\
+J_{\rm D} &= \mat H^T\mat R^{-1}\mat H
+\end{align}
+
 \section{Pendulum example}
 
-Let $\vec x_k = [a_k\quad b_k]^T$ where $a_k$ is the pendulum angle and $b_k$ is the angular velocity,
+Let $\vec x_k = [a_k\quad b_k\quad g]^T$ where $a_k$ is the pendulum angle and $b_k$ is the angular velocity and $g$ is an unknown deterministic parameter.
 \begin{align}
 	\vec f(\vec x_{k}) &= \left[\begin{array}{c}
-								a_k + b_k\Delta \\ -b_k - g\sin(a_k)\Delta
+								a_k + b_k\Delta \\ b_k - g\sin(a_k)\Delta \\ g
 								\end{array}\right]
@@
 \begin{align}
-	\mat F(\vec x_{k}) &= \left[\begin{array}{cc}
-					1 & \Delta \\
-					-g\cos(a_k) & 1 \\
+	\mat F(\vec x_{k}) &= \left[\begin{array}{ccc}
+					1 & \Delta & 0\\
+					-g\cos(a_k)\Delta & 1 & -\sin(a_k)\Delta\\
+					0 & 0 & 1
 				\end{array}\right]
 \end{align}
-The observation matrix is $H=[1\quad 0]$, so that only the first state element is measured.
+The observation matrix is $H=[1\quad 0\quad 0]$, so that only the first state element is measured.
 

### 2013-04-11 code/compute_pcrb_conditional.m

@@ -19,21 +19,26 @@
 % Kelvin Layton
 % Feb 2013
 %
-function pcrb = compute_pcrb_P(t,f,F,H,Q,R,m0,P0,M)
+function pcrb = compute_pcrb_conditional(t,f,F,H,Q,R,m0,P0,theta,M)
 
 N = length(t);
 Nstates=length(m0);
+Nparams=length(theta);
+Nunknown=Nstates+Nparams;
 
 % Initialise variables
 %
+P0all = blkdiag(P0,inf*ones(Nparams,Nparams));
+P=zeros(Nunknown,Nunknown,N);
+P(:,:,1)=P0all;
 
-P=zeros(Nstates,Nstates,N);
-P(:,:,1)=P0;
-pcrb=zeros(Nstates,N);
+pcrb=zeros(Nunknown,N);
+
 
 % Initalise all trajectories
 %
 xk=mvnrnd(m0,P0,M)';
+xk=[xk; bsxfun(@times,theta,ones(1,M))];
 
 Rinv=inv(R);
 
@@ -42,10 +47,10 @@ Rinv=inv(R);
 % Compute the PCRB using a Monte Carlo approximation
 %
 for k=2:N
-    Fhat = zeros(Nstates);
-    Rinvhat = zeros(Nstates);
+    Fhat = zeros(Nunknown);
+    Rinvhat = zeros(Nunknown);
     
-    v = mvnrnd(zeros(Nstates,1),Q,M)';
+    v = mvnrnd(zeros(Nunknown,1),Q,M)';
     parfor i=1:M
         % Sample the next time point for the current trajectory realisation
         %
@@ -62,9 +67,6 @@ for k=2:N
     Fhat=Fhat./M;
     Rinvhat=Rinvhat./M;
         
-%     Fnorm(k)=norm(Fhat);
-%     Feig(k) = min(eig(Fhat));
-    
     % Recursively compute the Fisher information matrix
     %
     P(:,:,k) = inv( inv(Fhat*P(:,:,k-1)*Fhat' + Q) + Rinvhat);

### 2013-04-11 code/example_ekf_augmented.m

@@ -5,98 +5,112 @@ close all
 % Define time
 %
 delta = 0.001;  % Sampling period
-N=400;
+N=40;
 t=(0:N-1)*delta;
 
 % Transition model
 %
 f = @(x)model_pendulum(x,delta,'transition');
 F = @(x)model_pendulum(x,delta,'jacobian');
+fa = @(x)model_pendulum_augmented(x,delta,'transition');
+Fa = @(x)model_pendulum_augmented(x,delta,'jacobian');
 
 % Observation matrix (only observe angle)
 %
 H = [1 0];
+Ha = [1 0 0];
 
 % Define noise covariances
 %
-Q = (0.01)^2*[1 0; 0 1];         % Process noise
-R = (0.1)^2 .* eye(size(H,1));    % Measurement noise
+Q = (0.01)^2*eye(2);
+Qa = blkdiag(Q,0);         % Process noise
+R = (0.1)^2 .* eye(size(Ha,1));    % Measurement noise 
 
 w = mvnrnd(zeros(size(R,1),1),R,N)';
-v = mvnrnd(zeros(size(Q,1),1),Q,N)';
+v = mvnrnd(zeros(size(Qa,1),1),Qa,N)';
+
+% Deterministic parameter
+%
+g = 100;
 
 % Starting state
 %
-x0 = [pi/2; 0];
+x0 = [pi/4; 0; g];
 
 % Calculate state trajectory
 %
 x = zeros(length(x0),N);
 x(:,1)=x0;
 for i=2:N
-    x(:,i) = f(x(:,i-1)) + v(:,i);
+    x(:,i) = fa(x(:,i-1)) + v(:,i);
 end
 
 % Define observation vector
 %
-z = H*x + w;
+z = Ha*x + w;
 
 % Prior distribution (defined by m0 & P0)
 %
-m0 = x0;
-P0 = diag(0.001*[(2*pi)^2 (2*pi)^2]);
+m0 = [pi/2; 0];
+m0a = [m0; g];
+P0 = diag([(0.01*pi)^2 (0.01*pi)^2]);
+P0param = 1000^2;
+P0a = blkdiag(P0,P0param);
 
 % Apply EKF filter
 %
-m = extended_kalman_filter(z,f,F,H,Q,R,m0,P0);
+m = extended_kalman_filter(z,fa,Fa,Ha,Qa,R,m0a,P0a);
+m2 = extended_kalman_filter(z,f,F,H,Q,R,m0,P0);
 
[…]

### 2013-04-11 code/example_kf_augmented.m

@@ -5,24 +5,24 @@ close all
 % Define time
 %
 delta = 0.01;  % Sampling period
-t=0:delta:4;
+t=(0:delta:4)';
 N=length(t);
 
 % Transition model (linear)
 %
-F = [1, delta; 0 1];
+F = [1, delta, 1; 0 1 0; 0 0 1];
 f = @(x)(F*x);
 Ffunc = @(x)F;
 
 % Observation matrix (only observe position)
 %
-H = [1 0];
+H = [1 0 0];
 
 % Define noise vectors
 %
 sigma_v = 0.01;  % Process noise
 sigma_w = 0.1;   % Measurement noise
-Q = sigma_v^2*[1 0; 0 1];
+Q = sigma_v^2*eye(3);
 R = sigma_w^2 .* eye(size(H,1));
 
 w = mvnrnd(zeros(size(R,1),1),R,N)';
@@ -30,7 +30,7 @@ v = mvnrnd(zeros(size(Q,1),1),Q,N)';
 
 % Starting state
 %
-x0 = [3; 0];
+x0 = [3; 0; 0.5];
 
 % Calculate state trajectory
 %
@@ -47,7 +47,7 @@ z = H*x + w;
 % Prior distribution (defined by m0 & P0)
 %
 m0 = x0;
-P0 = diag([(2*pi)^2 (2*pi)^2]);
+P0 = diag([(2*pi)^2 (2*pi)^2 (2*pi)^2]);
 
 % Apply EKF filter
 %
@@ -64,14 +64,14 @@ grid on;
 legend({'x_1 - position','x_2 - velocity','x_1 estimate','x_2 estimate'})
 % export_fig('../figures/motion_trajectory.pdf')
 
-
+return
 %% Compute the posterior Cramer-Rao bound (PCRB)
 %
 
 % Initialise variables
 %
-pcrb=zeros(2,N);
-J=zeros(2,2,N);
+pcrb=zeros(3,N);
+J=zeros(3,3,N);
 J(:,:,1)=inv(P0);
 
 % Compute the PCRB using closed form expressions
@@ -99,22 +99,22 @@ D21 = -F*inv(Q);
 D12 = D21';
 D22 = inv(Q) + H'*inv(R)*H;
 
-Jss = dare(inv(D11)*D12,eye(2),D22-D21*inv(D11)*D12,D11);
+Jss = dare(inv(D11)*D12,eye(3),D22-D21*inv(D11)*D12,D11);
 
[…]

### 2013-04-11 code/model_pendulum_augmented.m

@@ -1,0 +1,26 @@
+% MODEL_PENDULUM This function implements the state space representation of
+% the nonlinear pendulum equations.
+% 
+% Inputs: x - the current state
+%         delta - the time step
+%         mode - 'transition' returns the new state,
+%                'jacobian' returns the Jacobian at the current state.
+%
+% Kelvin Layton
+% Jan 2013
+%
+function [out] = model_pendulum_augmented(x, delta, mode)
+
+    if mode(1)=='t';
+        % Nonlinear transition model
+        out = [x(1) + x(2)*delta; ...
+                x(2) - x(3)*sin(x(1))*delta; ...
+                x(3)];
+
+    else
+        % Jacobian
+        out = [1, delta, 0; ...
+                -x(3)*cos(x(1))*delta, 1, -sin(x(1))*delta; ...
+                0, 0, 1];
+    end
+end

### 2013-04-11 manuscript/manuscript.tex
added figures and a decription of the models

@@
 
+\paragraph{Neural Mass Model}
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics{./figures/pdf/SingleMass.pdf}
+  \end{center}
+  \caption{\emph{\textbf{Neural Mass Model.} The somas of the neural mass are represented by the triangle.
+The synaptic connections are represented by the circle.
+The input mean firing rate, $g(v_m(t))$, is convolved with post-synaptic response kernel (PSRK), shown in the inset on the left, to give the mean membrane potential, $v_n(t)$.
+This is transformed by the sigmoid, shown in the inset on the right, to give the output mean firing rate, $g(v_n(t))$.}}
+  \label{fig:SingleNeuralMass}
+\end{figure}
+Neural mass models map from a mean firing rate of a population of pre-synaptic neurons to a mean membrane potential of a post-synaptic population, which in turn determines that population's output firing rate.
+A graphical description of a neural mass model can be seen in Figure~\ref{fig:SingleNeuralMass}.
+
+To define a standard neural mass model, we begin by defining the post-synaptic potential of population $n$, $v_n(t)$, as a result of an input firing rate from population $m$, $g(v_m(t))$, as the convolution
+\begin{equation}\label{eq:conv_eq}
+    v_n(t) = \frac{\alpha_{mn}}{\tau_{mn}}\int_{-\infty}^t  h_{mn}(t-t')g(v_m(t')) \,\mathrm{d}t',
+\end{equation}
+where $\alpha_{mn}$ is the gain for the post-synaptic response kernel (PSRK), denoted by $h_{mn}(t)$, from neural population $m$ to $n$, and $\tau_{mn}$ is the membrane time constant.
+A common form of the PSRK is
+\begin{equation}
+    h_{mn}(t) = \eta(t)t\exp\left(-\frac{t}{\tau_{mn}}\right),
+\end{equation}
+where $\eta(t)$ is the Heaviside step function. An example of a PSRK can be seen in the left inset of Figure~\ref{fig:SingleNeuralMass}.
+
+The firing rate of each population is related to the mean membrane potential by a sigmoidal activation function; the logistic function sigmoid is typically used: 
+\begin{align}
+    g\left(v_n(t)\right) =& \frac{1}{1+\exp{\left(\varsigma_n\left(v_{0n} - v_n(t)\right)\right)}}.
+    \label{eq:sigmoid}
+\end{align}
+The quantity $\varsigma_n$ describes the slope of the sigmoid (approximating the variance of firing thresholds within the populations) and $v_{0n}$ describes the mean firing threshold.
+An example of the sigmoidal activation function can be seen in the right inset of Figure~\ref{fig:SingleNeuralMass}.
+
+The convolution in Equation~\ref{eq:conv_eq} can also be written as the ordinary differential equation (ODE)
+\begin{equation}\label{eq:2ndOrder}
+    \mathrm{D}v_n(t) = \frac{\mathrm{d}^2 v_n(t)}{\mathrm{d}t^2} + \frac{2}{\tau_{mn}}\frac{\mathrm{d} v_n(t)}{\mathrm{d}t} + \frac{1}{\tau_{mn}^2} v_n(t) = \frac{\alpha_{mn}}{\tau_{mn}} g(v_m(t)),
+\end{equation}
+where $\mathrm{D}$ is a linear differential operator. Equation~\ref{eq:2ndOrder} can be written as two coupled first-order ODEs by 
+\begin{equation} \label{eq:2ndOrderNMM}
+    \frac{\mathrm{d} v_n(t)}{\mathrm{d}t} = z_n(t),\,\,\,\,\,    \frac{\mathrm{d}z_n(t)}{\mathrm{d}t} = \frac{\alpha_{mn}}{\tau_{mn}} g(v_m(t)) - \frac{2}{\tau_{mn}}z_n(t) - \frac{1}{\tau_{mn}^2} v_n(t),
+\end{equation}
+where $z_n(t)$ is a dummy variable. This forms the basis of a state-space model in a canonical format.
+
+The parameters of the model, $\tau$, $\alpha$, $v_0$ and $\varsigma$, can be set so the neural mass has characteristics of specific neural populations, such as pyramidal neurons, spiny stellate cells and fast and slow inhibitory interneurons (GABA$_\mathrm{a}$ and GABA$_\mathrm{b}$).
+The neural populations can then be connected to represent the circuitry of a cortical column and further to form networks of columns.
+Various kinds of neural mass models have been developed \cite{Silva1974,Jansen1995,Wendling2002,David2003}, which are depicted in Figure~\ref{fig:NMMs}.
+Each synaptic connection in these networks can be described by the 2$^{\mathrm{nd}}$-order system of Equation~\ref{eq:2ndOrderNMM}, where the resultant dimensions of the networks of neural mass models in Figures~\ref{fig:NMMs} a, b and c are 6, 12 and 18, respectively.
+\begin{figure}[ht]
+	\centering
+		\includegraphics[scale=1]{./figures/pdf/NeuralMassesHoriz.pdf}
+	\caption{\emph{\textbf{Models of Cortical Columns.} Each model has a pyramidal neural population, with various forms of feedback and inputs (Cx.
+= Cortex, Th.
+= Thalamus).
+The EEG is taken as the mean membrane potential of the pyramidal population, $v_p(t)$, with additive measurement noise.
+a) Minimal model of a column with inhibitory feedback \cite{Silva1974}.
+b) An extension with excitatory feedback and afferent input from thalamus \cite{Jansen1995,David2003}.
+c) Column with inhibition occurring across two times scales, enabling modelling of higher frequency activity observed in seizures~\cite{Wendling2002}.}}
+	\label{fig:NMMs}
+\end{figure}
+The parameters of the neural masses not only define the population type, but also the behaviour exhibited by the model.
+For example, certain parameter combinations result in a model of a cortical column that will generate alpha wave type activity (normal activity) and, for another set of parameters, we create a model that will exhibit epileptic behaviour \cite{Wendling2002}.
+Therefore, we consider a family of models, which we define generally as
+\begin{align}\label{eq:NeuralMassModel}
+    \dot{\mathbf{x}}(t) =& f_{\theta}\left(\mathbf{x}(t),\mathbf{u}(t)\right)\\
+    y(t) =& \mathbf{C}\mathbf{x}(t) + e(t),
+\end{align}
+where $\mathbf{x}(t) \in \mathbb{R}^{n_x}$ is a state vector representing the postsynaptic membrane potentials generated by each population synapse and their time derivatives, $n_x$ is the number of states and $\mathbf{u}(t)$ represents the system input, which may be from afferent connections or other brain regions, exogenous input, or account for model inaccuracies.
+The function $f_{\theta}(\cdot)$ describes the dynamics, where $\theta \in \mathbb{R}^{n_{\theta}}$ determines the model type and the behaviour it exhibits.
+The iEEG is denoted by $y(t)$, $\mathbf{C}$ is the observation matrix, and $e(t)$ is the observation noise.
+
 Present 3 models?

### 2013-04-11 code/test_model_erf.m

@@ -68,10 +68,9 @@ y = H*x + w;
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
 % Prior distribution (defined by m0 & P0)
-%
-m0 = x0;
-P0 = F(x0)*QnonSingular*F(x0)';
 
+m0 = x0;
+P0 = F(x0)*QnonSingular*F(x0)';             
 
 % Apply EKF filter
 %

### 2013-04-11 manuscript/manuscript.tex
add relation to general model to J and R

@@
 
+The model focused on in the following
+estimation sections is the formulation by Jansen and Rit \cite{Jansen-Rit-95}.
+Given space limitations we refer the reader to the article by Jansen
+and Rit\cite{Jansen-Rit-95} where the original state-space equations
+are presented. The parameters that were described in this section are related to the Jansen and Rit paper by
+\begin{align}
+    A =& \frac{\alpha_{pe}}{2e_0c_1} = \frac{\alpha_{pi}}{2e_0c_3} = \frac{\alpha_{ep}}{2e_0c_2} = \frac{\alpha_{xp}}{2e_0} \\
+    B =& \frac{\alpha_{ip}}{2e_0c_4} \\
+    a =& \frac{1}{\tau_{pe}} = \frac{1}{\tau_{pi}} = \frac{1}{\tau_{ep}} = \frac{1}{\tau_{xp}} \\
+    b =& \frac{1}{\tau_{ip}},
+\end{align}
+where $e_0$ is a parameter that scales the maximum firing rate, $A$ and $B$ are synaptic gains for excitation and inhibition respectively, and $a$ and $b$ are the reciprocals of the synaptic time constants for excitation and inhibition, respectively, and the subscripts $p$, $e$, $x$, and $i$ denote pyramidal ($p$), excitatory interneuron (spiny stellate) ($e$), external ($x$) or inhibitory interneuron ($i$) populations, respectively.
+By making the assumption that all excitatory synapses share the same time constants and by defining the connectivity constants, $c_1$, $c_2$, $c_3$ and $c_4$, the network of neural masses on the left side of Figure~\ref{fig:JansenAndRit} can be reduced to the system depicted on the right hand side.
 Present 3 models?

### 2013-04-18 manuscript/manuscript.tex
added bib
commit

@@
 where $e_0$ is a parameter that scales the maximum firing rate, $A$ and $B$ are synaptic gains for excitation and inhibition respectively, and $a$ and $b$ are the reciprocals of the synaptic time constants for excitation and inhibition, respectively, and the subscripts $p$, $e$, $x$, and $i$ denote pyramidal ($p$), excitatory interneuron (spiny stellate) ($e$), external ($x$) or inhibitory interneuron ($i$) populations, respectively.
-By making the assumption that all excitatory synapses share the same time constants and by defining the connectivity constants, $c_1$, $c_2$, $c_3$ and $c_4$, the network of neural masses on the left side of Figure~\ref{fig:JansenAndRit} can be reduced to the system depicted on the right hand side.
+By making the assumption that all excitatory synapses share the same time constants and by defining the connectivity constants, $c_1$, $c_2$, $c_3$ and $c_4$, the network of neural masses is the JR NMM.
+
 Present 3 models?
@@
 
-% \small
-% \bibliographystyle{ieeetr}
-% \bibliography{references}
+\small
+\bibliographystyle{ieeetr}
+\bibliography{References}
 

### 2013-05-09 code/example_ekf_augmented.m
unstable augmented

@@ -54,7 +54,7 @@ z = Ha*x + w;
 m0 = [pi/2; 0];
 m0a = [m0; g];
 P0 = diag([(0.01*pi)^2 (0.01*pi)^2]);
-P0param = 1000^2;
+P0param = 10^2;
 P0a = blkdiag(P0,P0param);
 
 % Apply EKF filter
@@ -74,7 +74,7 @@ grid on;
 legend({'x_1 - angle','x_2 - velocity','x_1 estimate','x_2 estimate'},'Location','Best')
 % export_fig('../figures/pendulum_trajectory.pdf')
 
-
+return
 %% Compute the posterior Cramer-Rao bound (PCRB)
 %
 
@@ -82,7 +82,7 @@ M = 10000;    % Number of Monte Carlo samples
 
 pcrb = compute_pcrb_P(t,f,F,@(x)H,Q,R,m0,P0,M);
 pcrb_a = compute_pcrb_P(t,fa,Fa,@(x)Ha,Qa,R,m0a,P0a,M);
-% pcrb2 = compute_pcrb_conditional(t,fa,Fa,@(x)Ha,Qa,R,m0(1:2),P0(1:2,1:2),g,M);
+pcrb2 = compute_pcrb_conditional(t,fa,Fa,@(x)Ha,Qa,R,m0(1:2),P0(1:2,1:2),g,M);
 
 return
 
@@ -127,6 +127,7 @@ figure
 % semilogy(t,(mse),'x-')
 semilogy(t,(pcrb),'o-'); hold on;
 semilogy(t,(pcrb_a),'s-');
+semilogy(t,(pcrb2),'d-');
 grid on;
 legend({'MSE','PCRB'})
 xlabel('Time (s)');

### 2013-05-09 code/example_kf_augmented.m

@@ -5,7 +5,7 @@ close all
 % Define time
 %
 delta = 0.01;  % Sampling period
-t=(0:delta:4)';
+t=(0:delta:0.1)';
 N=length(t);
 
 % Transition model (linear)
@@ -94,6 +94,7 @@ end
 %% Solution to Ricatti equation
 %
 
+if 0
 D11 = F'*inv(Q)*F;
 D21 = -F*inv(Q);
 D12 = D21';
@@ -102,7 +103,7 @@ D22 = inv(Q) + H'*inv(R)*H;
 Jss = dare(inv(D11)*D12,eye(3),D22-D21*inv(D11)*D12,D11);
 
 pcrb_ss = diag(inv(Jss))
-
+end
 
 %% Compute the MSE of the extended Kalman filter 
 %

### 2013-06-13 code/utils/find_erf_width.m
minor change

@@ -9,6 +9,10 @@ sig1 = 1 ./ (1 + exp(rSig*(v0Sig - x)));
 rErf = lsqnonlin(@cost_func,5);
 
 
+figure
+plot(x,sig1)
+hold
+plot(x,0.5 .* ( 1 + erf((x - v0Erf)./(sqrt(2)*rErf))),'r')
 
 function diff = cost_func(r)
     sig2 = 0.5 .* ( 1 + erf((x - v0Erf)./(sqrt(2)*r)) );

### 2013-06-13 manuscript/manuscript.tex
a
minor change
added plot for example data

@@
 
+Not claiming anything about which model is the 'best', but simply what level of complexity is possible - how increasing complexity affects uncertainty.
 \section{Background}
@@
 \end{figure}
+
+\begin{figure}[ht]
+	\centering
+		\includegraphics[scale=1]{./figures/pdf/Example_Data.pdf}
+		\caption{\textbf{Example data generated using the neural mass models.}}
+	\label{fig:NMMs}
+\end{figure}
+
 The parameters of the neural masses not only define the population type, but also the behaviour exhibited by the model.

### 2013-06-13 code/main_JR.m

@@ -2,7 +2,7 @@
 % estimation
 
 
-N = 1000;             	% number of samples
+N = 2000;             	% number of samples
 dT = 0.001;          	% sampling time step (global)
 dt = 1*dT;            	% integration time step
 nn = fix(dT/dt);      	% (used in for loop for forward modelling) the integration time step can be small that the sampling (fix round towards zero)
@@ -11,7 +11,7 @@ t = 0:dt:(N-1)*dt;
 
 % Intial true parameter values
 mode='alpha';
-mode='seizure';
+% mode='seizure';
 parameters = SetParametersJR(mode);
 parameters.dt = dt;
 A=parameters.A;
@@ -37,7 +37,6 @@ x(:,1) = x0;
 
 Q = zeros(NStates);
 Q(4,4) = (sqrt(dt)*A*a*sigma)^2;
-
 R = 1^2*eye(1);
 
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
@@ -58,7 +57,33 @@ H = [0 0 1 0 -1 0];           % observation function
 w = mvnrnd(zeros(size(H,1),1),R,N)';
 y = H*x + w;
 figure,plot(t,-H*x)
-% 
+
+
+% save(['JR_' mode '.mat'],'y','t') 
+
+
+%%
+% One column = 17.35 max
+% two column = 8.3 cm max
+FS = 8;             % fontsize - point
+font = 'arial';
+fig_width = 5.7;  % cm
+fig_height = 2;  % cm
+fig_filename = '/Users/dean/Projects/CRLBNeuro/manuscript/figures/pdf/JR_seizure.pdf';
+figure('units','centimeters','papersize',[fig_height fig_width],'filename',fig_filename,'position',[2 2 fig_width fig_height]);
+
+LW = 1;         % linewidth
+plot(t,-y,'k','linewidth',LW)
+axis tight
+axis off
+set(gca,'fontsize',FS,'fontname',font)
+text_x = -0.25;
+text_y = max(-y);
+text(text_x,text_y,'D','fontsize',FS,'fontname',font)
+% xlabel('Time (s)','fontsize',FS,'fontname',font)
+% ylabel('Amplitude','fontsize',FS,'fontname',font)
+
+%% 
[…]

### 2013-06-13 code/main_Lopes.m

@@ -4,7 +4,7 @@ clc
 close all
 clear
 
-N = 2000;             	% number of samples
+N = 5000;             	% number of samples
 dT = 0.001;          	% sampling time step (global)
 dt = 1*dT;            	% integration time step
 nn = fix(dT/dt);      	% (used in for loop for forward modelling) the integration time step can be small that the sampling (fix round towards zero)
@@ -59,6 +59,7 @@ R = 1^2*eye(1);
 w = mvnrnd(zeros(size(H,1),1),R,N)';
 y = H*x + w;
 
+% save(['Lopes_' mode '.mat'],'y','t') 
 % figure
 % plot(H*x)
 

### 2013-06-13 code/main_Wendling.m

@@ -4,7 +4,7 @@ clc
 close all
 clear
 
-N = 1000;             	% number of samples
+N = 2000;             	% number of samples
 dT = 0.001;          	% sampling time step (global)
 dt = 1*dT;            	% integration time step
 nn = fix(dT/dt);      	% (used in for loop for forward modelling) the integration time step can be small that the sampling (fix round towards zero)
@@ -13,8 +13,8 @@ t = 0:dt:(N-1)*dt;
 
 % Intial true parameter values
 % mode = 'background';
-% mode = 'alpha';       % this is more like alpha!!!
-mode = 'spikes';
+mode = 'alpha';       % this is more like alpha!!!
+% mode = 'spikes';
 parameters = SetParametersWendling(mode);
 parameters.dt = dt;
 A = parameters.A;
@@ -41,7 +41,7 @@ x(:,1) = x0;
 
 % Define input
 Q = zeros(NStates);
-Q(4,4) = (sqrt(dt)*A*a*sigma)^2;
+Q(4,4) = 0.4225*(sqrt(dt)*A*a*sigma)^2;         % note! the scaling makes the noise the same as the other models
 v = mvnrnd(zeros(NStates,1),Q,N)';
 
 % Euler-Maruyama integration
@@ -55,8 +55,9 @@ R = 1^2*eye(1);
 w = mvnrnd(zeros(size(H,1),1),R,N)';
 y = H*x + w;
 
-% figure
-% plot(t,-H*x)
+% save(['Wendling_' mode '.mat'],'y','t') 
+figure
+plot(t,-H*x)
 
 %% Run EKF for this model
 

### 2013-06-13 code/plot_simulated_data.m

@@ -1,0 +1,125 @@
+% One column = 17.35 max
+% two column = 8.3 cm max
+close all
+clc
+clear
+
+FS = 8;             % fontsize - point
+font = 'arial';
+fig_width = 17;  % cm
+fig_height = 5;  % cm
+
+LW = 1;         % linewidth
+
+min_y = -18;
+max_y = 27;
+
+fig_filename = '/Users/dean/Projects/CRLBNeuro/manuscript/figures/pdf/JR_seizure.pdf';
+figure('units','centimeters','papersize',[fig_height fig_width],'filename',fig_filename,'position',[2 2 fig_width fig_height]);
+
+
+%
+load Lopes_alpha.mat
+y = detrend(y);
+max_l_a = max(y(3001:end))
+min_l_a = min(y(3001:end))
+
+ax(1) = subplot(231);
+
+plot(t(3001:end)-3,y(3001:end),'k','linewidth',LW)
+
+ylim(ax(1),[min_y,max_y])
+xlim(ax(1),[0 2])
+
+axis off
+set(gca,'fontsize',FS,'fontname',font)
+text_x = -0.25;
+text_y = max(y(3001:end));
+text(text_x,text_y,'A','fontsize',FS,'fontname',font)
+
+%
+load JR_alpha.mat
+y = detrend(y);
+max_j_a = max(y)
+min_j_a = min(y)
+
+ax(2) = subplot(232);
+
+plot(t,y,'k','linewidth',LW)
+
+ylim(ax(2),[min_y,max_y])
+xlim(ax(2),[0 2])
+
+axis off
+set(gca,'fontsize',FS,'fontname',font)
+text_x = -0.25;
+text_y = max(y);
+text(text_x,text_y,'B','fontsize',FS,'fontname',font)
+
+load JR_seizure.mat
+y = detrend(y);
+max_j_s = max(y)
+min_j_s = min(y)
+
+ax(3) = subplot(235);
+
+plot(t,y,'k','linewidth',LW)
+
+ylim(ax(3),[min_y,max_y])
+xlim(ax(3),[0 2])
+
+axis off
+set(gca,'fontsize',FS,'fontname',font)
+text_x = -0.25;
+text_y = max(-y);
+text(text_x,text_y,'D','fontsize',FS,'fontname',font)
+
+%
+load Wendling_alpha.mat
+y = detrend(y);
+max_w_a = max(y)
+min_w_a = min(y)
+
[…]

### 2013-06-13 manuscript/manuscript.tex
Draft of theory section

@@
 \section{Background}
+\label{sec:background}
 
@@
 \begin{align}\label{eq:NeuralMassModel}
-    \dot{\mathbf{x}}(t) =& f_{\theta}\left(\mathbf{x}(t),\mathbf{u}(t)\right)\\
-    y(t) =& \mathbf{C}\mathbf{x}(t) + e(t),
+    \dot{\vec{x}}(t) =& \vec f_{\theta}\left(\vec{x}(t),\vec{u}(t)\right)\\
+    y(t) =& \mat{C}\vec{x}(t) + e(t),
 \end{align}
-where $\mathbf{x}(t) \in \mathbb{R}^{n_x}$ is a state vector representing the postsynaptic membrane potentials generated by each population synapse and their time derivatives, $n_x$ is the number of states and $\mathbf{u}(t)$ represents the system input, which may be from afferent connections or other brain regions, exogenous input, or account for model inaccuracies.
-The function $f_{\theta}(\cdot)$ describes the dynamics, where $\theta \in \mathbb{R}^{n_{\theta}}$ determines the model type and the behaviour it exhibits.
-The iEEG is denoted by $y(t)$, $\mathbf{C}$ is the observation matrix, and $e(t)$ is the observation noise.
+where $\vec{x}(t) \in \mathbb{R}^{n_x}$ is a state vector representing the postsynaptic membrane potentials generated by each population synapse and their time derivatives, $n_x$ is the number of states and $\vec{u}(t)$ represents the system input, which may be from afferent connections or other brain regions, exogenous input, or account for model inaccuracies.
+The function $\vec f_{\theta}(\cdot)$ describes the dynamics, where $\theta \in \mathbb{R}^{n_{\theta}}$ determines the model type and the behaviour it exhibits.
+The iEEG is denoted by $y(t)$, $\vec{C}$ is the observation matrix, and $e(t)$ is the observation noise.
 
@@
 
-Present 3 models?
+\section{Theory}
 
-\section{Theory / Performance Bounds}
+The Bayesian Cramer-Rao bound (BCRB) can be used to examine the inherent uncertainty in a neural mass model and also assess the performance of different estimation algorithms.
 
-Introduce recursive computation of bound and Monte Carlo approximation
+INCLUDE TIME DEPENDENCE STRAIGHT AWAY
 
-\section{Methods?}
+Central to the BCRB is the Bayes information matrix defined as
+\begin{align}
+	\mat J_k &= \mathbb E_{\vec x,\vec y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right] 
+\end{align}
+The expectation is over the state, $\vec x$ and the measurements, $\vec y$ with respect to the joint distribution $p(\vec y,\vec x)$.
 
-\section{Results}
+The BCRB is based on the Bayes information matrix, $J_{\rm B}$, which contains a data term and a prior term \cite{sdf},
+\begin{align}
+	J_{\rm B} = J_{\rm D} + J_{\rm P}
+\end{align}
+where 
+\begin{align}
+	J_{\rm D} &= \mathbb E_{\vec x,\vec y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log \ell(\vec y|\vec x) \right] \\
+	J_{\rm P} &= \mathbb E_{\vec x}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec x) \right]
+\end{align}
+The data term, $J_{\rm D}$, depends on the likelihood, $\ell$, and the prior term, $J_{\rm P}$, depends on the prior distribution, $p$.
+The BCRB is ideally suited to filtering applications since the prior distribution captures the information up to the current time.
+
+The BCRB gives a lower bound on the mean squared error
+\begin{align}
+	\mathsf{MSE}_k = \mathbb E_{\vec x_k,\vec y} [(\hat{\vec x}_k(\vec y) - \vec x_k)(\hat{\vec x}_k(\vec y) - \vec x_k)^T] \ge \mat J_{k}^{-1}
+\end{align}
+where $\hat{\vec x}_k$ is the estimated state at time $k$ given the measurements from time $1,\ldots,k$.
+
+\subsection{Recursive Bayesian CRB}
+
+It is necessary to develop a recursive formula for $\mat J_k$ for filtering applications to prevent the above computations becoming unwieldy.
+The popular recursion in \cite{Tichavsky1998} is not applicable to the neural field models described in Section~\ref{sec:background} since noise only enters the system through a single state and the resultant covariance matrix, $\mat Q$, is singular.
+The recursion in \cite{Bergman2001} is applicable to this class of models and can be summarised as
+\begin{align}
+	\mat J_k = \left( \mat Q + \bar{\mat F}_{k-1} \mat J_{k-1}^{-1} \bar{\mat F}_{k-1}^T\right)^{-1} + \mat C^T \mat R^{-1} \mat C
+\end{align}
+where
+\begin{align}
+		\bar{\mat F}_k &= \mathbb E_{\vec x_k} \left[ \nabla_{\vec x_k}^T \vec f(\vec x_k)\right] 
+		\label{eqn:F_matrix_expectation}
+\end{align}
+See Appendix~\ref{sec:derivation_recursion} for details.
+The expectation over the state $\vec x_k$ in \eqref{eqn:F_matrix_expectation} can rarely be computed in closed form so Monte Carlo techniques are employed.
+A large number of state realisations are generated at each time point and the expectation is approximated by an average over the realisations.
+
+\section{Simulations}
 
@@
 
+\appendix
+\section{Derivation of Recursive Bayesian CRB}
+\label{sec:derivation_recursion}
+
+The recursion in \cite{Bergman2001} is
+\begin{subequations}
+\begin{align}
+	\mat P_{k|k-1} &= \bar{\mat F}_k\mat P_{k-1|k-1}\bar{\mat F}_k + \bar{\mat G}_k\mat Q_k \bar{\mat G}_k^{-1} \\
+		\mat P_{k|k} &= \left(\mat P_{k|k-1}^{-1} + \bar{\mat R}_{k-1}^{-1}\right)^{-1}
+\end{align}
+\end{subequations}
+where
+\begin{align}
+	\bar{\mat F}_k &= \mathbb E \left[ \nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
+	\bar{\mat R}_k^{-1} &= -\mathbb E\left[ \nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) \right] \\
+	\bar{\mat G}_k &= \mathbb E \left[ \nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
+	\mat Q_k^{-1} &= -\mathbb E \left[ \nabla_{\vec v_k}\nabla_{\vec v_k}^T \log p(\vec v_k) \right]
+\end{align}
+and $\mat P_0^{-1}$ is the inverse of the prior covariance.
+The model in \eqref{eqn:model_general} is represented by $\tilde{\vec f}(\vec x_k,\vec v_k) = \vec f(\vec x_k) + \vec v_k$.
+This simplifies the derivatives to
+\begin{align}
+	\nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat F_k(\vec x_k) \\
+	\nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat I_{d\times d} \\
+	\nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) &= -\mat H^T\mat R^{-1} \mat H 
+\end{align}
+and the corresponding matrices to	
+\begin{align}
+	\bar{\mat F}_k &= \mathbb E\left[ \mat F_k(\vec x_k)\right] \label{eqn:pcrb_term_F}\\
+	\bar{\mat R}_k^{-1} &= \mat H^T\mat R^{-1} \mat H \\
+	\bar{\mat G}_k &= \mat I_{d\times d} 
+\end{align}
+Therefore the final recursion is
+\begin{align}
+	\mat J_{k|k} = \left( \mat Q + \bar{\mat F}_{k-1} \mat J_{k-1}^{-1} \bar{\mat F}_{k-1}^T\right)^{-1} + \mat H_k^T \mat R^{-1} \mat H_k
+\end{align}
+
 \small

### 2013-06-13 code/plot_simulated_data.m
updated figure

@@ -5,11 +5,13 @@ clc
 clear
 
 FS = 8;             % fontsize - point
+FS_Panel_Label = 12;
+
 font = 'arial';
 fig_width = 17;  % cm
 fig_height = 5;  % cm
 
-LW = 1;         % linewidth
+LW = 0.5;         % linewidth
 
 min_y = -18;
 max_y = 27;
@@ -109,7 +111,7 @@ axis off
 set(gca,'fontsize',FS,'fontname',font)
 text_x = -0.25;
 text_y = max(y);
-text(text_x,text_y,'E','fontsize',FS,'fontname',font)
+text(text_x,text_y,'E','fontsize',FS_Panel_Label,'fontname',font)
 
 % this is the legenend for time and amplitude
 ax(6) = subplot(234);

### 2013-06-16 manuscript/manuscript.tex
Updated theory section

@@
 The Bayesian Cramer-Rao bound (BCRB) can be used to examine the inherent uncertainty in a neural mass model and also assess the performance of different estimation algorithms.
+The bound was formulated in \cite{VanTrees1968} for the estimation of random parameters, making it well suited to the filtering problem described above, where the state is considered a random variable.
+The BCRB is a lower bound on the mean squared error of an estimator $\hat{\vec x}(\vec y)$ given by
+\begin{align}
+	\mathsf{MSE} = \mathbb E_{\vec x,\vec y} [(\hat{\vec x}(\vec y) - \vec x)(\hat{\vec x}(\vec y) - \vec x)^T] \ge \mat J^{-1}
+	\label{eqn:mse_bound}
+\end{align}
+where $\mat J$ is the Bayes information matrix defined as
+\begin{align}
+	\mat J &= -\mathbb E_{\vec x,\vec y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right] 
+	\label{eqn:bayes_matrix}
+\end{align}
+where $\nabla_{\vec x} = [\partial/\partial x_1,\ldots,\partial/\partial x_n]^T$ and the expectation is taken over the state $\vec x$ and the measurements $\vec y$.
 
-INCLUDE TIME DEPENDENCE STRAIGHT AWAY
+\subsection{Recursive Bayesian CRB}
 
-Central to the BCRB is the Bayes information matrix defined as
+To calculate a bound for filtering applications we are required to calculate $\mat J$ defined by \eqref{eqn:bayes_matrix} at each time, $k$.
+Fortunately, the structure of the joint density $p(\vec y,\vec x)$ allows for a recursive method to calculate $\mat J_k$ from $\mat J_{k-1}$ as demonstrated in \cite{Tichavsky1998}.
+However, this recursion is not directly applicable to the neural field models described in Section~\ref{sec:background} since noise only enters the system through a single state and the resultant covariance matrix, $\mat Q$, is singular.
+An alternative recursion proposed in \cite{Bergman2001} is suitable to this class of models and can be summarised as
 \begin{align}
-	\mat J_k &= \mathbb E_{\vec x,\vec y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right] 
-\end{align}
-The expectation is over the state, $\vec x$ and the measurements, $\vec y$ with respect to the joint distribution $p(\vec y,\vec x)$.
-
-The BCRB is based on the Bayes information matrix, $J_{\rm B}$, which contains a data term and a prior term \cite{sdf},
-\begin{align}
-	J_{\rm B} = J_{\rm D} + J_{\rm P}
+	\mat J_k = \left( \mat Q + {\mat F}_{k-1} \mat J_{k-1}^{-1} {\mat F}_{k-1}^T\right)^{-1} + \mat C^T \mat R^{-1} \mat C
 \end{align}
@@
 \begin{align}
-	J_{\rm D} &= \mathbb E_{\vec x,\vec y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log \ell(\vec y|\vec x) \right] \\
-	J_{\rm P} &= \mathbb E_{\vec x}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec x) \right]
-\end{align}
-The data term, $J_{\rm D}$, depends on the likelihood, $\ell$, and the prior term, $J_{\rm P}$, depends on the prior distribution, $p$.
-The BCRB is ideally suited to filtering applications since the prior distribution captures the information up to the current time.
-
-The BCRB gives a lower bound on the mean squared error
-\begin{align}
-	\mathsf{MSE}_k = \mathbb E_{\vec x_k,\vec y} [(\hat{\vec x}_k(\vec y) - \vec x_k)(\hat{\vec x}_k(\vec y) - \vec x_k)^T] \ge \mat J_{k}^{-1}
-\end{align}
-where $\hat{\vec x}_k$ is the estimated state at time $k$ given the measurements from time $1,\ldots,k$.
-
-\subsection{Recursive Bayesian CRB}
-
-It is necessary to develop a recursive formula for $\mat J_k$ for filtering applications to prevent the above computations becoming unwieldy.
-The popular recursion in \cite{Tichavsky1998} is not applicable to the neural field models described in Section~\ref{sec:background} since noise only enters the system through a single state and the resultant covariance matrix, $\mat Q$, is singular.
-The recursion in \cite{Bergman2001} is applicable to this class of models and can be summarised as
-\begin{align}
-	\mat J_k = \left( \mat Q + \bar{\mat F}_{k-1} \mat J_{k-1}^{-1} \bar{\mat F}_{k-1}^T\right)^{-1} + \mat C^T \mat R^{-1} \mat C
-\end{align}
-where
-\begin{align}
-		\bar{\mat F}_k &= \mathbb E_{\vec x_k} \left[ \nabla_{\vec x_k}^T \vec f(\vec x_k)\right] 
+		{\mat F}_k &= \mathbb E_{\vec x_k} \left[ \nabla_{\vec x_k}^T \vec f(\vec x_k)\right] 
 		\label{eqn:F_matrix_expectation}
@@
 See Appendix~\ref{sec:derivation_recursion} for details.
-The expectation over the state $\vec x_k$ in \eqref{eqn:F_matrix_expectation} can rarely be computed in closed form so Monte Carlo techniques are employed.
+The expectation over the state $\vec x_k$ in \eqref{eqn:F_matrix_expectation} can rarely be computed in closed form for nonlinear transition functions so Monte Carlo techniques must be employed.
 A large number of state realisations are generated at each time point and the expectation is approximated by an average over the realisations.
@@
 \appendix
-\section{Derivation of Recursive Bayesian CRB}
-\label{sec:derivation_recursion}
+\section{Derivation of Recursive Bayesian CRB}\label{sec:derivation_recursion}
 

### 2013-06-18 manuscript/manuscript.tex
Draft simulation section

@@
 However, this recursion is not directly applicable to the neural field models described in Section~\ref{sec:background} since noise only enters the system through a single state and the resultant covariance matrix, $\mat Q$, is singular.
-An alternative recursion proposed in \cite{Bergman2001} is suitable to this class of models and can be summarised as
+An alternative recursion proposed in \cite{Bergman2001} is suitable to the models considered here and can be summarised as
 \begin{align}
@@
 
+The bound defined above can be used in several ways.
+Firstly, different models can be compared to test the feasibility of estimating the unknown states from the measurements available.
+Secondly, different parameter sets can be tested within a single model to assess how the performance of an estimator is expected to change for different types of activity.
+Finally, the efficiency of a given estimation algorithm (e.g.
+EKF, particle filter) can be assessed against the lower bound.
+This indicated whether developing progressively advanced estimation algorithms is warranted, a question that has not been considered in the neural estimation literature thus far.
+
+This paper assumes the parameters of a given model are constant and known.
+However, it is important to note that the results and conclusions presented in this paper are directly applicable to the case of slowly varying unknown parameters.
+This statement can be quantified by first considering a state vector that is augmented with constant unknown parameters \cite{sdf}.
+In this case, the information about the parameters accumulates each time step and the performance bound asymptotically approaches the case for constant known parameters.
+As an extension, if the parameters are varying slowly with respect to the sampling rate the bound will remain sufficiently close to the case of known parameters.
+
 \section{Simulations}
 
-\subsection{Effect of increasing model complexity -- bound should increase}
+\subsection{Across models - model feasibility}
 
-e.g. Bound for model A > model B > model C
+We consider the feasibility of the three models described in Section~\ref{sec:background}.
+Models A and B contain 6 unknown states and model C has 10 unknown states.
+An important question to ask as the models become increasingly complex is: how well can we estimate the states from the available measurements?
+To investigate this, we compute the BCRB for each model in an operating mode analogous to alpha rhythms.
+The specific parameters are listed in Appendix~\ref{sec:parameter_table}.
 
-\subsection{Effect of parameters changing -- bound should change}
+\subsection{Within model - seizure, alpha}
 
-e.g. seizure easier to estimate than alpha rhythms
+It has been well demonstrated that models of type B and C can generate vastly different oscillations for different parameters \cite{sdf}.
+Bifurcation theory provides a theoretical tool to analyse this feature and has been reported in several places \cite{sdf}.
+Good qualitative agreement between EEG data and the models has been observed and consequently different parameters sets have been assigned to different types of brain activity such as alpha rhythms or epileptic seizures \cite{sfd}.
+
+Here, we investigate ones ability to estimate the states of a model for different types of activity.
+We calculate the BCRB for model B for two parameter sets representing alpha rhythms and epileptic seizures.
+
+Fig.~\ref{sdf} illustrates that the seizure type behaviour is easier to estimate than background alpha rhythms...<text here>.
+This is perhaps not surprising since, intuitively, seizure oscillations appear more structured and have larger amplitudes as illustrated in Fig.~\ref{fig:NMMs}.
+
+\subsection{Benchmark algorithms}
+
+e.g. EKF doesn't reach bound for model X, thus better algorithms are required.
 
@@
 
-\subsection{Augmented models (joint estimation of parameters and states)?}
-
 \section{Discussion}
@@
 extensions to other applications e.g. experimental design
+
+algorithm performance
 
@@
 
+\section{List of notation and parameters}
+\label{sec:parameter_table}
+
+Table here
+
 \small

### 2013-06-18 code/PlotEPSP.m
updated figures

@@ -1,0 +1,104 @@
+clc
+close all
+clear
+
+t=linspace(0,0.1,1000);
+A = 3.25;       % mV
+a = 100;        % s^{-1}
+h_e=A*a*t.*exp(-a*t);
+
+fig_height = 3;
+fig_width = 3;
+FS_label = 8;
+FS_ticks = 8;
+
+set(0,'defaultAxesFontName', 'timesnewroman')
+set(0,'defaultTextFontName', 'timesnewroman')
+set(0,'defaultUicontrolFontName', 'timesnewroman')
+set(0,'defaultUitableFontName', 'timesnewroman')
+set(0,'defaultUipanelFontName', 'timesnewroman')
+set(0,'defaultuicontrolunits','centimeters') ;
+
+%%
+
+figure('units','centimeters',...
+    'color','white',...
+    'position',[0 0 fig_width fig_height],...
+    'papersize',[fig_width fig_height],...
+    'PaperPositionMode','auto',...
+    'renderer','painters')
+
+ax1 = axes;
+Left =1.5;
+Bottom = 0.9;
+Width = 2;
+Height = 2;
+set(ax1,'units','centimeters','position',[Left Bottom Width Height]);
+
+plot(ax1,[1/a 1/a],[0 max(h_e)],'-.r'),hold on
+plot(ax1,[t(1) t(end)],[max(h_e) max(h_e)],'--')
+plot(ax1,t,h_e,'k')
+% axis tight
+% plot(t,(A/a)*ones(1,1000),'k')
+
+box off
+set(gca,'fontsize',FS_ticks,'xticklabel',{},'yticklabel',{})
+leg = legend('$\tau$','$\propto \alpha$');
+set(leg,'units','centimeters',...
+    'interpreter','latex',...
+    'location','southeast',...
+    'box','off',...
+    'fontsize',FS_label,...
+    'fontname','timesnewroman',...
+    'position',[2.2 1.5 1.25 1]);
+axis tight
+ylim([0 max(h_e)+0.1])
+xlabel('Time (s)','fontname','arial','fontsize',FS_label)
+ylabel('PSRK (mV)','fontsize',FS_label,'fontname','arial')
+
+
+
+%%
+
+e_0 = 1;          % s^{-1}
+v_0 = 6;              % mV
+r = 0.56;           % mV^{-1}
+
+min_v = 15;
+max_v = 20;
+v = linspace(v_0-min_v,v_0+20,1000);
+
[…]

### 2013-06-18 code/plot_simulated_data.m
added correct figures

@@ -37,7 +37,7 @@ axis off
 set(gca,'fontsize',FS,'fontname',font)
 text_x = -0.25;
 text_y = max(y(3001:end));
-text(text_x,text_y,'A','fontsize',FS,'fontname',font)
+text(text_x,text_y,'A','fontsize',FS_Panel_Label,'fontname',font)
 
 %
 load JR_alpha.mat
@@ -56,7 +56,7 @@ axis off
 set(gca,'fontsize',FS,'fontname',font)
 text_x = -0.25;
 text_y = max(y);
-text(text_x,text_y,'B','fontsize',FS,'fontname',font)
+text(text_x,text_y,'B','fontsize',FS_Panel_Label,'fontname',font)
 
 load JR_seizure.mat
 y = detrend(y);
@@ -74,7 +74,7 @@ axis off
 set(gca,'fontsize',FS,'fontname',font)
 text_x = -0.25;
 text_y = max(-y);
-text(text_x,text_y,'D','fontsize',FS,'fontname',font)
+text(text_x,text_y,'D','fontsize',FS_Panel_Label,'fontname',font)
 
 %
 load Wendling_alpha.mat
@@ -93,7 +93,7 @@ axis off
 set(gca,'fontsize',FS,'fontname',font)
 text_x = -0.25;
 text_y = max(y);
-text(text_x,text_y,'C','fontsize',FS,'fontname',font)
+text(text_x,text_y,'C','fontsize',FS_Panel_Label,'fontname',font)
 
 load Wendling_spikes.mat
 y = detrend(y);

### 2013-06-30 manuscript/manuscript.tex
updated figures
added correct figures
edits to intro

@@
 
+\newcommand{\unit}[1]{\textcolor{green}{#1}}
+\newcommand{\brian}[1]{\textcolor{blue}{#1}}
+
 \title{Statistical performance analysis of neural mass models?}
@@
 
-History of models and estimation attempts
+Mesoscopic mathematical models of the brain are becoming increasingly popular as tools for generating hypotheses on the physical principals that govern neurodynamics.
+Forward models have led to theory on the emergence of the brains rhythms, the mechanisms of anesthetic agents, the brains response to sensory stimuli, short-term memory formation, visual hallucinations and epileptic seizures.
+Most recently, advances engineering methods and computing resources have expanded the utility of these models as integral elements of frameworks to estimate of system and parametric states that can not be directed measured using electrophysiology.
+An established framework to reliably estimate physical properties underlying brain dynamics has the potential to revolutionise how we observe, diagnose, treat and cure neurological disorders.
+For example, in epilepsy the dynamics of hidden system and parametric states are thought to be highly patient-specific and unknown.
+Thus the ability to estimate these quantities using the framework will enable a greater understanding of individuals’ pathologies, leading to more tailored therapies and better patient outcomes.
+Furthermore, subject-specific models will enable the application of control theory to the neurosciences, which is typically limited to man-made systems.
+
+To date a large body of work has demonstrated validity of these models in-principal, through the ability to reproduce experimental data.
+The next stage is understanding the performance of estimation algorithms.
+Once models have been fit to subjects a final step in assessing validity will be to examine the within subject predictive power.In order for such a framework to be established as a clinical platform further work is required for establishing validity and feasibility for the mesoscopic models.
+This paper introduces engineering tools to neuroscience community for formally accessing feasibility.
+We also postulate how these tools can be used to increase performance of estimation methods.
+
+Influence on sampling rate and estimation?
+
+ of this framework History of models and estimation attempts
 
@@
   \begin{center}
-    \includegraphics{./figures/pdf/SingleMass.pdf}
+    \includegraphics{./figures/pdf/SingleMass_plos.pdf}
   \end{center}
@@
 	\centering
-		\includegraphics[scale=1]{./figures/pdf/NeuralMassesHoriz.pdf}
+		\includegraphics[scale=1]{./figures/pdf/NeuralMassesHoriz_plos.pdf}
 	\caption{\emph{\textbf{Models of Cortical Columns.} Each model has a pyramidal neural population, with various forms of feedback and inputs (Cx.

### 2013-07-17 code/compare_models.m
Working on comparing models

@@ -1,6 +1,6 @@
 
 N = 400;             	% number of samples
-dT = 0.001;          	% sampling time step (global)
+dT = 0.0001;          	% sampling time step (global)
 dt = 1*dT;            	% integration time step
 nn = fix(dT/dt);      	% (used in for loop for forward modelling) the integration time step can be small that the sampling (fix round towards zero)
 
@@ -28,16 +28,16 @@ a = paramsL.a;
 sigma = paramsL.sigma;
 
 Q_NM = zeros(4);
-Q_NM(2,2) = (sqrt(dt)*A*a*sigma)^2;
+Q_NM(2,2) = (10*sqrt(dt)*A*a*sigma)^2;
 
 Q_L = zeros(6);
-Q_L(2,2) = (sqrt(dt)*A*a*sigma)^2;
+Q_L(2,2) = (10*sqrt(dt)*A*a*sigma)^2;
 
 Q_JR = zeros(6);
-Q_JR(4,4) = (sqrt(dt)*A*a*sigma)^2;
+Q_JR(4,4) = (10*sqrt(dt)*A*a*sigma)^2;
 
 Q_W = zeros(10);
-Q_W(4,4) = (sqrt(dt)*A*a*sigma)^2;
+Q_W(4,4) = (10*sqrt(dt)*A*a*sigma)^2;
 
 H_NM = [1 0 0 0];
 H_L = [1 0 -1 0 0 0];
@@ -63,22 +63,22 @@ F_W = @(x)model_Wendling(x,'jacobian',paramsW);
 % Priors
 
 m0_NM = zeros(4,1);
-P0_NM = 100.^2*eye(4);
+P0_NM = 200.^2*eye(4);
 
 m0_LJR = zeros(6,1);
-P0_LJR = 100.^2*eye(6);
+P0_LJR = 200.^2*eye(6);
 
 m0_W = zeros(10,1);
-P0_W = 100.^2*eye(10);
+P0_W = 200.^2*eye(10);
 
 %% 
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 % Compute the posterior Cramer-Rao bound (PCRB)
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
[…]

### 2013-07-18 manuscript/manuscript.tex
minor edit
Working on comparing models
Merged changes to mansuscript
Code and figure for EKF benchmarking

@@
 \section{Introduction}
+\label{sec:introduction}
 
@@
 
-e.g. EKF doesn't reach bound for model X, thus better algorithms are required.
+EKF doesn't reach bound for any nonlinear model, thus better algorithms are required.
 
-\subsection{Effect of measurement noise? -- pretty boring}
-
-e.g. scalp EEG vs iEEG?
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics{./figures/pdf/CRBbar}
+  \end{center}
+  \caption{The mean squared error of the EKF against the BCRB for three models}
+  \label{fig:CrbBar}
+\end{figure}
 
@@
 contextualize work w.r.t. existing literature?
+
+we say nothing about how algorithms will work with uncertainty in the actual model -- thus can be considered a ``best case'' situation, where we know the true model exactly.
 

### 2013-07-18 code/generateEKFresults.m

@@ -1,0 +1,138 @@
+% This code repricates the structure of the Voss / Schiff FN simulation /
+% estimation
+clc
+close all
+clear
+
+N = 100;             	% number of samples
+dT = 0.001;          	% sampling time step (global)
+dt = 1*dT;            	% integration time step
+nn = fix(dT/dt);      	% (used in for loop for forward modelling) the integration time step can be small that the sampling (fix round towards zero)
+
+t = 0:dt:(N-1)*dt;
+
+
+% Initialise random number generator for repeatability
+rng(0);
+
+M = 100;    % Mumber of Monte Carlo trials for MSE calculation
+
+% Intial true parameter values
+% mode = 'background';
+mode = 'alpha';       
+
+NStatesModels = [6 6 10 3];  
+
+% Model parameters
+params{1} = SetParametersJR(mode);
+params{2} = SetParametersLopes(mode);
+params{3} = SetParametersWendling(mode);
+params{4} = params{1}; % Only need dt
+
+% Transition functions
+models{1} = @model_JR;
+models{2} = @model_Lopes;
+models{3} = @model_Wendling;
+models{4} = @model_Linear;
+
+% Measurement fnctions
+Hmodels{1} = [0 0 1 0 -1 0];           % JR
+Hmodels{2} = [1 0 -1 0 0 0];           % Lopes
+Hmodels{3} = [0 0 1 0 -1 0 -1 0 0 0];  % Wendling
+Hmodels{4} = [1 1 1];
+
+% Indices to plot (only potentials, not derivatives)
+displayStates{1} = [1 3 5];
+displayStates{2} = [1 3 5];
+displayStates{3} = [1 3 5 7 9];
+displayStates{4} = [1 2 3];
+noiseInd = [4 2 4 2];     % NEED TO CHECK THESE! (and adjust Wendling noise)
+noiseScale = [1 1 0.4225 1];
+
+mseModels = cell(length(models),1);
+crbModels = cell(length(models),1);
+
+for iModel = 1:length(models)
+    
+parameters = params{iModel};
+parameters.dt = dt;
+A = parameters.A;
+a = parameters.a;
+sigma = parameters.sigma;
[…]

### 2013-07-18 code/model_Linear.m

@@ -1,0 +1,20 @@
+% MODEL_Linear This function implements the state space representation of
+% a simple linear equation
+% 
+% Inputs: x - the current state
+%         params - a structure of parameters
+%         mode - 'transition' returns the new state,
+%                'jacobian' returns the Jacobian at the current state.
+%
+% Jan 2013
+%
+function out = model_Linear(x,mode,params)
+
+F = [1, params.dt 0; 0 1 0; 0 params.dt 1];
+% F = [1, params.dt ; 0 1];
+
+if mode(1)=='t';
+    out = F*x;
+else
+    out = F;
+end

### 2013-07-18 code/plotEKFresults.m

@@ -1,0 +1,52 @@
+
+FS = 8;             % fontsize - point
+FS_Panel_Label = 12;
+
+font = 'arial';
+fig_width = 17;  % cm
+fig_height = 7;  % cm
+
+LW = 0.5;         % linewidth
+
+fig_filename = '/Users/klayton/Dropbox/CRLBNeuro/manuscript/figures/pdf/CRBbar.pdf';
+figure('units','centimeters','papersize',[fig_height fig_width],'filename',fig_filename,'position',[2 2 fig_width fig_height]);
+
+load crb_ekf_results.mat
+
+%%
+mseBar = [];
+crbBar = [];
+for iModel=1:length(mseModels)
+    mseBar = [mseBar; mseModels{iModel}; nan];
+    crbBar = [crbBar; crbModels{iModel}; nan];
+end
+
+
+b2=bar(mseBar(:)','BaseValue',1e-6,'BarWidth',0.15); hold on;
+b1=bar(crbBar(:)','BaseValue',1e-6); hold on;
+set(gca,'YScale','log','YGrid','on')
+set(b2,'EdgeColor','none')
+
+
+ch1 = get(b1,'children');
+ch2 = get(b2,'children');
+cind = [1;1;1;1;2;2;2;2;3;3;3;3;3;3;4;4;4;4];
+set(ch1,'FaceVertexCData',cind)
+set(ch2,'FaceVertexCData',cind)
+cmap = [0.7556 0.8196 0.4769; 0.6590 0.4332 0.2696; 0.2245 0.3898 0.5888];
+cmap = rand(4,3);
+colormap(cmap);
+
+for i=1:length(mseBar)
+l1=plot(i,mseBar(i),'o');
+set(l1,'MarkerSize',7,'MarkerEdgeColor','k','MarkerFaceColor',cmap(cind(i),:));
+end
+
+laby=ylabel('MSE');
+set(gca,'XTick',[2 6 11 16],'XTickLabel',{'JR','Lopes','Wendling','Linear Sanity Check'})
+% set(gca,'YTick',[1e-6 1e-4 1e-2 1e0 1e2])
+ylim([1e-6 1e3])
+
+set([gca laby],'fontsize',FS,'fontname',font)
+
+export_fig(fig_filename)

### 2013-07-30 manuscript/manuscript.tex
edit to method fisher info

@@
 Forward models have led to theory on the emergence of the brains rhythms, the mechanisms of anesthetic agents, the brains response to sensory stimuli, short-term memory formation, visual hallucinations and epileptic seizures.
-Most recently, advances engineering methods and computing resources have expanded the utility of these models as integral elements of frameworks to estimate of system and parametric states that can not be directed measured using electrophysiology.
+Most recently, advances engineering methods and computing resources have expanded the utility of these models as integral elements of frameworks to estimate of system and parametric states that can not be directly measured using electrophysiology.
 An established framework to reliably estimate physical properties underlying brain dynamics has the potential to revolutionise how we observe, diagnose, treat and cure neurological disorders.
@@
 
-The Bayesian Cramer-Rao bound (BCRB) can be used to examine the inherent uncertainty in a neural mass model and also assess the performance of different estimation algorithms.
-The bound was formulated in \cite{VanTrees1968} for the estimation of random parameters, making it well suited to the filtering problem described above, where the state is considered a random variable.
-The BCRB is a lower bound on the mean squared error of an estimator $\hat{\vec x}(\vec y)$ given by
+The Bayesian Cramer-Rao bound (BCRB) can be used to examine the inherent uncertainty in a neural mass model and assess the expected performance of different estimation algorithms.
+The bound was first formulated in \cite{VanTrees1968} for estimation of random parameters, making it well suited to the filtering problem described above when the states (membrane potentials and derivatives) are considered uncertain or random variable.
+The BCRB is a lower bound on the mean squared error of an estimator, $\hat{\vec x}(\vec y)$, given by
 \begin{align}
@@
 where $\nabla_{\vec x} = [\partial/\partial x_1,\ldots,\partial/\partial x_n]^T$ and the expectation is taken over the state $\vec x$ and the measurements $\vec y$.
+In Bayesian estimation schemes (including Kalman filtering) $\hat{\vec x}(\vec y) = \mathbb{E}_{\vec x}\left[p\left(x|y\right)\right]$.
+
+The provide some intuition into this Matrix, we can write $p(\vec y,\vec x)$ as the product of the likelihood of the states given the observations and the prior information of the states
+\begin{equation}
+	\log(p(\vec y,\vec x)) = \log\left(\mathcal{L}(\vec x|\vec y)\right) + \log\left(p(\vec x)\right)
+\end{equation}
 

### 2013-08-08 code/generateAlphaSeizureResults.m
Within model comparisons
Included code and figure for alpha vs seizure. Also changed order of
EKF results to match paper

@@ -17,37 +17,35 @@ rng(0);
 
 M = 100;    % Mumber of Monte Carlo trials for MSE calculation
 
-% Intial true parameter values
-% mode = 'background';
-mode = 'alpha';       
 
-NStatesModels = [6 6 10 3];  
+NStatesModels = [6 10 6 10];  
 
 % Model parameters
-params{1} = SetParametersLopes(mode);
-params{2} = SetParametersJR(mode);
-params{3} = SetParametersWendling(mode);
-params{4} = params{1}; % Only need dt
+params{1} = SetParametersJR('alpha');
+params{2} = SetParametersWendling('alpha');
+params{3} = SetParametersJR('seizure');
+params{4} = SetParametersWendling('spikes');
 
 % Transition functions
-models{1} = @model_Lopes;
-models{2} = @model_JR;
-models{3} = @model_Wendling;
-models{4} = @model_Linear;
+models{1} = @model_JR;
+models{2} = @model_Wendling;
+models{3} = @model_JR;
+models{4} = @model_Wendling;
 
 % Measurement fnctions
-Hmodels{1} = [1 0 -1 0 0 0];           % Lopes
-Hmodels{2} = [0 0 1 0 -1 0];           % JR
-Hmodels{3} = [0 0 1 0 -1 0 -1 0 0 0];  % Wendling
-Hmodels{4} = [1 1 1];
+Hmodels{1} = [0 0 1 0 -1 0];           % JR
+Hmodels{2} = [0 0 1 0 -1 0 -1 0 0 0];  % Wendling
+Hmodels{3} = [0 0 1 0 -1 0];           % JR
+Hmodels{4} = [0 0 1 0 -1 0 -1 0 0 0];  % Wendling
 
 % Indices to plot (only potentials, not derivatives)
 displayStates{1} = [1 3 5];
-displayStates{2} = [1 3 5];
-displayStates{3} = [1 3 5 7 9];
-displayStates{4} = [1 2 3];
-noiseInd = [2 4 4 2];     % NEED TO CHECK THESE! (and adjust Wendling noise)
-noiseScale = [1 1 0.4225 1];
+displayStates{2} = [1 3 5 7 9];
+displayStates{3} = [1 3 5];
[…]

### 2013-08-08 code/plotAlphaSeizureResults.m

@@ -1,0 +1,46 @@
+
+
+FS = 8;             % fontsize - point
+FS_Panel_Label = 12;
+
+font = 'arial';
+fig_width = 17;  % cm
+fig_height = 7;  % cm
+
+LW = 0.5;         % linewidth
+
+fig_filename = '/Users/kelvin/Dropbox/CRLBNeuro/manuscript/figures/pdf/AlphaSeizureBar.pdf';
+figure('units','centimeters','papersize',[fig_height fig_width],'filename',fig_filename,'position',[2 2 fig_width fig_height]);
+
+load crb_seizure_results.mat
+
+%%
+crbBar = [];
+for iModel=[1 3 2 4]
+    crbBar = [crbBar; sum(sqrt(crbModels{iModel})); nan];
+end
+
+b1=bar(crbBar(:)','BaseValue',1e-6); hold on;
+% set(gca,'YScale','log','YGrid','on')
+set(gca,'YScale','linear','YGrid','on')
+% xlim([0 20])
+
+ch1 = get(b1,'children');
+% cind = [1;1;1;1;2;2;2;2;2;2;3;3;3;3;4;4;4;4;4;4];
+cind = [1 1 2 2 3 3 4 4]';
+set(ch1,'FaceVertexCData',cind)
+cmap = [0.7556 0.8196 0.4769; 0.6590 0.4332 0.2696; 0.2245 0.3898 0.5888; 0.1831 0.9336 0.5528];
+% cmap = rand(4,3);
+colormap(cmap);
+
+laby=ylabel('RMSE (mV)');
+% set(gca,'XTick',[2 6 11 16],'XTickLabel',{'JR alpha','JR seizure','Wendling alpha','Wendling seizure'})
+set(gca,'XTick',[1 3 5 7],'XTickLabel',{'JR alpha','JR seizure','Wendling alpha','Wendling seizure'})
+% ylim([1e-6 1e3])
+ylim([0 10])
+
+set([gca laby],'fontsize',FS,'fontname',font)
+
+return
+%%
+export_fig(fig_filename)

### 2013-08-08 code/plotCRBdemo.m

@@ -1,0 +1,102 @@
+
+FS = 8;             % fontsize - point
+FS_Panel_Label = 12;
+
+font = 'arial';
+fig_width = 20;  % cm
+fig_height = 20;  % cm
+
+LW = 0.5;         % linewidth
+
+fig_filename = '/Users/klayton/Dropbox/CRLBNeuro/manuscript/figures/pdf/CRBbar.pdf';
+figure('units','centimeters','papersize',[fig_height fig_width],'filename',fig_filename,'position',[2 2 fig_width fig_height]);
+
+xValues=linspace(-2,8,100);
+yValues=linspace(-4,6,100);
+[X,Y]=meshgrid(xValues,yValues);
+
+x0 = 3;
+R = 1;
+P0 = 4;
+
+MU = X./100;
+H = 1./100.*ones(size(MU));
+
+
+p_like = normpdf(Y,MU,R);
+p_x = normpdf(X,x0,P0);
+p_xy = p_like.*p_x;
+
+der_p_x = -inv(P0)*ones(size(X));
+der_p_like = -H.^2*inv(R);
+der_p_xy = der_p_x + der_p_like;
+% der_p_xy = der_p_xy.*p_xy;
+
+R = 0.5;
+P0 = 1;
+
+p_like = normpdf(Y,MU,R);
+p_x = normpdf(X,x0,P0);
+p_xy2 = p_like.*p_x;
+
+der_p_x = -inv(P0)*ones(size(X));
+der_p_like = -H.^2*inv(R);
+der_p_xy2 = der_p_x + der_p_like;
+% der_p_xy2 = der_p_xy2.*p_xy2;
+
+axes('Pos',[0 0 0.5 0.5])
+mesh(xValues,yValues,p_xy)
+
+axes('Pos',[0.5 0 0.5 0.5])
+mesh(xValues,yValues,-der_p_xy)
+
+
+axes('Pos',[0 0.5 0.5 0.5])
+mesh(xValues,yValues,p_xy2)
+
+axes('Pos',[0.5 0.5 0.5 0.5])
+mesh(xValues,yValues,-der_p_xy2)
+
+return
+
+%%
+mseBar = [];
+crbBar = [];
+for iModel=1:length(mseModels)
+    mseBar = [mseBar; sqrt(mseModels{iModel}); nan];
+    crbBar = [crbBar; sqrt(crbModels{iModel}); nan];
+end
+
+
[…]

### 2013-08-08 code/plotEKFresults.m

@@ -8,7 +8,7 @@ fig_height = 7;  % cm
 
 LW = 0.5;         % linewidth
 
-fig_filename = '/Users/klayton/Dropbox/CRLBNeuro/manuscript/figures/pdf/CRBbar.pdf';
+fig_filename = '/Users/kelvin/Dropbox/CRLBNeuro/manuscript/figures/pdf/CRBbar.pdf';
 figure('units','centimeters','papersize',[fig_height fig_width],'filename',fig_filename,'position',[2 2 fig_width fig_height]);
 
 load crb_ekf_results.mat
@@ -17,14 +17,15 @@ load crb_ekf_results.mat
 mseBar = [];
 crbBar = [];
 for iModel=1:length(mseModels)
-    mseBar = [mseBar; mseModels{iModel}; nan];
-    crbBar = [crbBar; crbModels{iModel}; nan];
+    mseBar = [mseBar; sqrt(mseModels{iModel}); nan];
+    crbBar = [crbBar; sqrt(crbModels{iModel}); nan];
 end
 
 
 b2=bar(mseBar(:)','BaseValue',1e-6,'BarWidth',0.15); hold on;
 b1=bar(crbBar(:)','BaseValue',1e-6); hold on;
 set(gca,'YScale','log','YGrid','on')
+% set(gca,'YScale','linear','YGrid','on')
 set(b2,'EdgeColor','none')
 
 
@@ -33,8 +34,8 @@ ch2 = get(b2,'children');
 cind = [1;1;1;1;2;2;2;2;3;3;3;3;3;3;4;4;4;4];
 set(ch1,'FaceVertexCData',cind)
 set(ch2,'FaceVertexCData',cind)
-cmap = [0.7556 0.8196 0.4769; 0.6590 0.4332 0.2696; 0.2245 0.3898 0.5888];
-cmap = rand(4,3);
+cmap = [0.7556 0.8196 0.4769; 0.6590 0.4332 0.2696; 0.2245 0.3898 0.5888; 0.1831 0.9336 0.5528];
+% cmap = rand(4,3);
 colormap(cmap);
 
 for i=1:length(mseBar)
@@ -42,11 +43,14 @@ l1=plot(i,mseBar(i),'o');
 set(l1,'MarkerSize',7,'MarkerEdgeColor','k','MarkerFaceColor',cmap(cind(i),:));
 end
 
-laby=ylabel('MSE');
[…]

### 2013-08-09 manuscript/manuscript.tex
Within model comparisons
Included code and figure for alpha vs seizure. Also changed order of
EKF results to match paper
edits to into and method

@@
 
-\title{Statistical performance analysis of neural mass models?}
+\title{Statistical performance analysis of data-driven neural mass models}
 \date{\today}
@@
 Most recently, advances engineering methods and computing resources have expanded the utility of these models as integral elements of frameworks to estimate of system and parametric states that can not be directly measured using electrophysiology.
-An established framework to reliably estimate physical properties underlying brain dynamics has the potential to revolutionise how we observe, diagnose, treat and cure neurological disorders.
+An established framework to reliably estimate physical properties underlying brain dynamics has the potential to revolutionize how we observe, diagnose, treat and cure neurological disorders.
 For example, in epilepsy the dynamics of hidden system and parametric states are thought to be highly patient-specific and unknown.
-Thus the ability to estimate these quantities using the framework will enable a greater understanding of individuals’ pathologies, leading to more tailored therapies and better patient outcomes.
+Thus the ability to estimate these quantities using the framework will enable a greater understanding of individuals pathologies, leading to more tailored therapies and better patient outcomes.
 Furthermore, subject-specific models will enable the application of control theory to the neurosciences, which is typically limited to man-made systems.
@@
 To date a large body of work has demonstrated validity of these models in-principal, through the ability to reproduce experimental data.
-The next stage is understanding the performance of estimation algorithms.
-Once models have been fit to subjects a final step in assessing validity will be to examine the within subject predictive power.In order for such a framework to be established as a clinical platform further work is required for establishing validity and feasibility for the mesoscopic models.
+Since the parameters of the models are expected to vary across individuals and across disease states, the models must be tuned to individuals to establish clinical utility.
+A critical steps in establishing robust model fitting techniques is understanding the performance and feasibility of estimation algorithms.
 This paper introduces engineering tools to neuroscience community for formally accessing feasibility.
@@
 
-Influence on sampling rate and estimation?
+Once models have been fit to subjects a final step in assessing validity will be to examine the within subject predictive power.
+In order for such a framework to be established as a clinical platform further work is required for establishing validity and feasibility for the mesoscopic models.
 
- of this framework History of models and estimation attempts
+% Influence on sampling rate and estimation?
+% of this framework History of models and estimation attempts
+% Usefulness of bound: benchmark algorithms, experimental design, insight about inherent uncertainty -- which do we address
+% Not claiming anything about which model is the 'best', but simply what level of complexity is possible - how increasing complexity affects uncertainty.
 
-Usefulness of bound: benchmark algorithms, experimental design, insight about inherent uncertainty -- which do we address
-
-Not claiming anything about which model is the 'best', but simply what level of complexity is possible - how increasing complexity affects uncertainty.
 \section{Background}
@@
 \section{Theory}
-
 The Bayesian Cramer-Rao bound (BCRB) can be used to examine the inherent uncertainty in a neural mass model and assess the expected performance of different estimation algorithms.
 The bound was first formulated in \cite{VanTrees1968} for estimation of random parameters, making it well suited to the filtering problem described above when the states (membrane potentials and derivatives) are considered uncertain or random variable.
-The BCRB is a lower bound on the mean squared error of an estimator, $\hat{\vec x}(\vec y)$, given by
+The BCRB is a lower bound on the mean squared error (MSE) of an estimator, $\hat{\vec x}(\vec y)$, given by
 \begin{align}
-	\mathsf{MSE} = \mathbb E_{\vec x,\vec y} [(\hat{\vec x}(\vec y) - \vec x)(\hat{\vec x}(\vec y) - \vec x)^T] \ge \mat J^{-1}
+	\mathsf{MSE} = \mathbb E_{\vec x,\vec y} [(\hat{\vec x}(\vec y) - \vec x)(\hat{\vec x}(\vec y) - \vec x)^T] \ge \mat J^{-1},
 	\label{eqn:mse_bound}
@@
 \begin{align}
-	\mat J &= -\mathbb E_{\vec x,\vec y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right] 
+	\mat J &= -\mathbb E_{\vec x,\vec y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right],
 	\label{eqn:bayes_matrix}
 \end{align}
-where $\nabla_{\vec x} = [\partial/\partial x_1,\ldots,\partial/\partial x_n]^T$ and the expectation is taken over the state $\vec x$ and the measurements $\vec y$.
+$\nabla_{\vec x} = [\partial/\partial x_1,\ldots,\partial/\partial x_n]^T$, the expectation is taken over the state vector $\vec x$, and the measurements $\vec y$.
 In Bayesian estimation schemes (including Kalman filtering) $\hat{\vec x}(\vec y) = \mathbb{E}_{\vec x}\left[p\left(x|y\right)\right]$.
 
-The provide some intuition into this Matrix, we can write $p(\vec y,\vec x)$ as the product of the likelihood of the states given the observations and the prior information of the states
-\begin{equation}
-	\log(p(\vec y,\vec x)) = \log\left(\mathcal{L}(\vec x|\vec y)\right) + \log\left(p(\vec x)\right)
-\end{equation}
+To provide some intuition into the Bayes information matrix, we can write $p(\vec y,\vec x)$ as the product of the likelihood of the states given the observations and the prior information of the states
+\begin{align}
+	p(\vec y,\vec x) =& \mathcal{L}(\vec x|\vec y) p(\vec x) \\
+	\log(p(\vec y,\vec x)) =& \underbrace{\log\left(\mathcal{L}(\vec x|\vec y)\right)}_{\mathrm{Fisher Information}} + \log\left(p(\vec x)\right) \\
+	-\mathbb E_{\vec x,\vec y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right] =& 
+\end{align}
 
 \subsection{Recursive Bayesian CRB}
-
 To calculate a bound for filtering applications we are required to calculate $\mat J$ defined by \eqref{eqn:bayes_matrix} at each time, $k$.
@@
 The specific parameters are listed in Appendix~\ref{sec:parameter_table}.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics{./figures/pdf/AlphaSeizureBar}
+  \end{center}
+  \caption{The BCRB for two models in alpha and seizure states}
+  \label{fig:CrbBar}
+\end{figure}
 

### 2013-08-22 code/main_JR.m
minor edit

@@ -11,7 +11,7 @@ t = 0:dt:(N-1)*dt;
 
 % Intial true parameter values
 mode='alpha';
-% mode='seizure';
+mode='seizure';
 parameters = SetParametersJR(mode);
 parameters.dt = dt;
 A=parameters.A;

### 2013-08-22 manuscript/manuscript.tex
Manuscript and MATLAB code

@@
   \caption{The BCRB for two models in alpha and seizure states}
-  \label{fig:CrbBar}
+  \label{fig:AlphaSeizure}
 \end{figure}
@@
 
-Fig.~\ref{sdf} illustrates that the seizure type behaviour is easier to estimate than background alpha rhythms...<text here>.
-This is perhaps not surprising since, intuitively, seizure oscillations appear more structured and have larger amplitudes as illustrated in Fig.~\ref{fig:NMMs}.
+Fig.~\ref{fig:AlphaSeizure} illustrates that the states underlying seizure type behaviour is more difficult to estimate than background alpha rhythms.
+This is expected since the input noise driving the system is amplified by a larger synaptic gain parameter, $A$, for the seizure state, as listed in Table~\ref{tbl:Parameters}.
+This increases the overall uncertainty in the model, reduces the information obtained by the measurements and thus increases the bound.
 
@@
 Table here
+\begin{table}
+
+	\caption{\label{tbl:Parameters}Parameters used to generate different states}
+\end{table}
 

### 2013-08-22 code/generateConvergenceResults.m

@@ -1,0 +1,105 @@
+% This code repricates the structure of the Voss / Schiff FN simulation /
+% estimation
+clc
+close all
+clear
+
+N = 500;             	% number of samples
+dT = 0.001;          	% sampling time step (global)
+dt = 1*dT;            	% integration time step
+nn = fix(dT/dt);      	% (used in for loop for forward modelling) the integration time step can be small that the sampling (fix round towards zero)
+
+t = 0:dt:(N-1)*dt;
+
+
+% Initialise random number generator for repeatability
+rng(0);
+
+M = 100;    % Mumber of Monte Carlo trials for MSE calculation
+
+
+NStatesModels = [6 6];  
+
+% Model parameters
+params{1} = SetParametersLopes('alpha');
+params{2} = SetParametersLopes('alpha');
+
+
+% Transition functions
+models{1} = @model_Lopes;
+models{2} = @model_Lopes;
+
+% Measurement fnctions
+Hmodels{1} = [1 0 -1 0 0 0];           % Lopes
+Hmodels{2} = [1 0 -1 0 0 0];           % Lopes
+
+% Indices to plot (only potentials, not derivatives)
+displayStates{1} = [1 3 5];
+displayStates{2} = [1 3 5];
+
+noiseInd = [2 2 ];     % NEED TO CHECK THESE! (and adjust Wendling noise)
+noiseScale = [1 1 ];
+
+mseModels = cell(length(models),1);
+crbModels = cell(length(models),1);
+
+for iModel = 1:length(models)
+    
+parameters = params{iModel};
+parameters.dt = dt;
+A = parameters.A;
+a = parameters.a;
+sigma = parameters.sigma;
+
+% Transition model
+NStates = NStatesModels(iModel);                           
+f = @(x)models{iModel}(x,'transition',parameters);
+F = @(x)models{iModel}(x,'jacobian',parameters);
+
+H = Hmodels{iModel};
+
+displayInd = displayStates{iModel};
+
+% Initialise trajectory state
+x0 = zeros(NStates,1);                   % initial state
+
+% Define input
+Q = zeros(NStates);
[…]

### 2013-08-22 code/plotConvergenceResults.m

@@ -1,0 +1,46 @@
+
+
+FS = 8;             % fontsize - point
+FS_Panel_Label = 12;
+
+font = 'arial';
+fig_width = 17;  % cm
+fig_height = 7;  % cm
+
+LW = 0.5;         % linewidth
+
+fig_filename = '/Users/kelvin/Dropbox/CRLBNeuro/manuscript/figures/pdf/AlphaSeizureBar.pdf';
+figure('units','centimeters','papersize',[fig_height fig_width],'filename',fig_filename,'position',[2 2 fig_width fig_height]);
+
+load crb_convergence_results.mat
+
+%%
+crbBar = [];
+for iModel=[1 3 2 4]
+    crbBar = [crbBar; sum(sqrt(crbModels{iModel})); nan];
+end
+
+b1=bar(crbBar(:)','BaseValue',1e-6); hold on;
+% set(gca,'YScale','log','YGrid','on')
+set(gca,'YScale','linear','YGrid','on')
+% xlim([0 20])
+
+ch1 = get(b1,'children');
+% cind = [1;1;1;1;2;2;2;2;2;2;3;3;3;3;4;4;4;4;4;4];
+cind = [1 1 2 2 3 3 4 4]';
+set(ch1,'FaceVertexCData',cind)
+cmap = [0.7556 0.8196 0.4769; 0.6590 0.4332 0.2696; 0.2245 0.3898 0.5888; 0.1831 0.9336 0.5528];
+% cmap = rand(4,3);
+colormap(cmap);
+
+laby=ylabel('RMSE (mV)');
+% set(gca,'XTick',[2 6 11 16],'XTickLabel',{'JR alpha','JR seizure','Wendling alpha','Wendling seizure'})
+set(gca,'XTick',[1 3 5 7],'XTickLabel',{'JR alpha','JR seizure','Wendling alpha','Wendling seizure'})
+% ylim([1e-6 1e3])
+ylim([0 10])
+
+set([gca laby],'fontsize',FS,'fontname',font)
+
+return
+%%
+export_fig(fig_filename)

### 2013-08-22 manuscript/manuscript.tex
edits to dicussion

@@
 \section{Discussion}
+\brian{Brief Summary of Paper}
+This paper has introduced a method for evaluating Bayesian estimation algorithms for fitting neural mass models to electrophysiological data.
+The results demonstrate that it is feasible to recover three neural mass models from data, assuming the model is accurate.
 
-extensions to other applications e.g. experimental design
+\brian{Feasibility of Models}
+We have demonstrated how the method can be used to test the feasibility of recovering the states of models from data.
+In other words, the method can be used to see if a model is too complex to be recovered data regardless of the estimation method.
+If the measurement has no /or little information about the states then the bound will be infinitely high.
 
-algorithm performance
+\brian{Is there a better observation to improve estimation? / experimental design}
+The BCRLB can improve experimental design, where the bound can be used to select what is important to measure. For example, 
+sampling rate affects the CRLB, extensions to other applications e.g. experimental design
+adaptive sensing
+
+\brian{Calculating this provides an extra layer of intuition into the estimation process}
+The evolution of the bound over time provides indication of convergence properties
+
+derivatives higher -> more information
+
+\brian{Stimulation / Excitation to maximize information}
+Probing - maybe
+
+\brian{Is there a better estimator?}
+algorithm performance bench marking
+
+\brian{Not model section}
+Best case analysis which assumes the data comes from the model
+we say nothing about how algorithms will work with uncertainty in the actual model -- thus can be considered a ``best case'' situation, where we know the true model exactly.
+Model section still needs to be done by checking how the models can predict the behavior of the system, check need to be checked experimentally.
+
+\brian{This method is the standard in radar and tracking and papers will not get published without it}
+This is the first point of call for benchmarking estimation and tracking algorithms in the radar applications (refs).
+We argue that this should also be the case for neural state estimation and tracking.
+Neuroscience stands to benefit greatly from estimation and control techniques from engineering
+
+\brian{Big picture}
+Find the right level of model complexity and measurements to get to the truth.
+
+\brian{What a subject specific model will provide the scientist and the neurologist}
 
 contextualize work w.r.t. existing literature?
-
-we say nothing about how algorithms will work with uncertainty in the actual model -- thus can be considered a ``best case'' situation, where we know the true model exactly.
 

### 2013-08-29 manuscript/manuscript.tex
Updated theory section

@@
 The Bayesian Cramer-Rao bound (BCRB) can be used to examine the inherent uncertainty in a neural mass model and assess the expected performance of different estimation algorithms.
-The bound was first formulated in \cite{VanTrees1968} for estimation of random parameters, making it well suited to the filtering problem described above when the states (membrane potentials and derivatives) are considered uncertain or random variable.
+The bound was first formulated in \cite{VanTrees1968} for estimation of random parameters, making it well suited to the filtering problem described above when the states (membrane potentials and derivatives) are modelled as random variables.
 The BCRB is a lower bound on the mean squared error (MSE) of an estimator, $\hat{\vec x}(\vec y)$, given by
@@
 \end{align}
-$\nabla_{\vec x} = [\partial/\partial x_1,\ldots,\partial/\partial x_n]^T$, the expectation is taken over the state vector $\vec x$, and the measurements $\vec y$.
-In Bayesian estimation schemes (including Kalman filtering) $\hat{\vec x}(\vec y) = \mathbb{E}_{\vec x}\left[p\left(x|y\right)\right]$.
+$\nabla_{\vec x} = [\partial/\partial x_1,\ldots,\partial/\partial x_n]^T$, and the expectation is taken over both the state vector $\vec x$ and the measurements $\vec y$.
+In many Bayesian estimation schemes (including Kalman filtering) the estimator is the mean of the posterior distribution, $\hat{\vec x}(\vec y) = \mathbb{E}_{\vec x}\left[ p (\vec x|\vec y)\right]$.
 
-To provide some intuition into the Bayes information matrix, we can write $p(\vec y,\vec x)$ as the product of the likelihood of the states given the observations and the prior information of the states
+To provide some intuition into the Bayes information matrix, we can write the joint distribution $p(\vec y,\vec x)$ as the product of the likelihood function and the prior distribution.
+Taking the logarithm gives
 \begin{align}
-	p(\vec y,\vec x) =& \mathcal{L}(\vec x|\vec y) p(\vec x) \\
-	\log(p(\vec y,\vec x)) =& \underbrace{\log\left(\mathcal{L}(\vec x|\vec y)\right)}_{\mathrm{Fisher Information}} + \log\left(p(\vec x)\right) \\
-	-\mathbb E_{\vec x,\vec y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right] =& 
+	\log p(\vec y,\vec x) =& \log\mathcal{L}(\vec x|\vec y) + \log p(\vec x).
 \end{align}
+Using this relation, the Bayes information matrix in \eqref{eqn:bayes_matrix} can be decomposed into a data component, $\mat J_D$, and a prior component, $\mat J_P$, such that $\mat J = \mat J_D + \mat J_P$.
+The two components are defined as
+\begin{align}
+	\mat J_D &= \mathbb E_{\vec x}\left[ \mat J_F(\vec x) \right]  \\
+	\mat J_P &= -\mathbb E_{\vec x}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec x) \right]
+\end{align}
+where $\mat J_F$ is the classical Fisher information matrix for deterministic states,
+\begin{align}
+	\mat J_F(\vec x) &= -\mathbb E_{\vec y| \vec x}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y|\vec x) \right].
+\end{align}
+Some important insights can be gained from this decomposition.
+First, the contribution of the data to the Bayes information matrix is the expectation of the classical Fisher information matrix taken over all possible states.
+Secondly, the contribution from the prior distribution directly adds to the Bayes information matrix, capturing our prior knowledge about of the states.
 
@@
 To calculate a bound for filtering applications we are required to calculate $\mat J$ defined by \eqref{eqn:bayes_matrix} at each time, $k$.
-Fortunately, the structure of the joint density $p(\vec y,\vec x)$ allows for a recursive method to calculate $\mat J_k$ from $\mat J_{k-1}$ as demonstrated in \cite{Tichavsky1998}.
-However, this recursion is not directly applicable to the neural field models described in Section~\ref{sec:background} since noise only enters the system through a single state and the resultant covariance matrix, $\mat Q$, is singular.
+Fortunately, the structure of the joint density $p(\vec y,\vec x)$ allows for a recursive method to calculate $\mat J_k$ from $\mat J_{k-1}$, first demonstrated in \cite{Tichavsky1998}.
+The recursion presented in that work is not directly applicable to the neural field models described in Section~\ref{sec:background} since noise only enters the system through a single state and the resultant covariance matrix, $\mat Q$, is singular.
 An alternative recursion proposed in \cite{Bergman2001} is suitable to the models considered here and can be summarised as
@@
 
-The bound defined above can be used in several ways.
-Firstly, different models can be compared to test the feasibility of estimating the unknown states from the measurements available.
+The bound defined in \eqref{eqn:mse_bound} can be used in several ways.
+First, different models can be compared to test the feasibility of estimating the unknown states from the measurements available.
 Secondly, different parameter sets can be tested within a single model to assess how the performance of an estimator is expected to change for different types of activity.
-Finally, the efficiency of a given estimation algorithm (e.g.
+Thirdly, the efficiency of a given estimation algorithm (e.g.
 EKF, particle filter) can be assessed against the lower bound.

### 2013-08-29 code/generateConvergenceResults.m

@@ -85,7 +85,7 @@ R = 10^2*eye(1);
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 % Compute CRLB
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-crbAll = compute_pcrb_P(t,f,F,Hfun,Q,R,m0,P0,2000);
+crbAll = compute_pcrb_P(t,f,F,Hfun,Q,R,m0,P0,200);
 
 crb = crbAll(:,end);
 

### 2013-09-04 manuscript/manuscript.tex
edits to discussion

@@
 This paper has introduced a method for evaluating Bayesian estimation algorithms for fitting neural mass models to electrophysiological data.
-The results demonstrate that it is feasible to recover three neural mass models from data, assuming the model is accurate.
+By providing bounds on the estimation accuracy, the results demonstrate that it is feasible to recover three neural mass models of varying complexity from data.
+Under the assumption that these models capture the essential dynamics of the phenomena observed in intracranial electroencephalography, the results imply that is possible to tailor computational models to an individual's brain dynamics.
 
 \brian{Feasibility of Models}
-We have demonstrated how the method can be used to test the feasibility of recovering the states of models from data.
-In other words, the method can be used to see if a model is too complex to be recovered data regardless of the estimation method.
-If the measurement has no /or little information about the states then the bound will be infinitely high.
+We have demonstrated how the methods described in this paper can be used to test the feasibility of recovering the states of computational models from data.
+In other words, the method can be used to test if a model is too complex to be recovered data, regardless of the estimation method.
+The Bayesian Cramer-Rao lower bound (BCRLB) essentially measures the amount of information regarding the states that is present in the measurement.
+If the measurement has no information regarding the states of the model, then the bound will be infinitely high.
+
+Neural mass models have evolved from a description of interaction of two very general populations, excitatory and inhibitory neurons, to interactions of specific populations that describe various networks.
+Cortical neural populations include pyramidal, spiny stellate, fast and slow inhibitory interneurons (ref).
+Deep brain populations include globus paladius, substansia niagara, sub-thalamic nuclei and more (ref).
+As models become more biologically realistic the BCRLB will be come more important for subject-specific modelling.
+Given that these models are all approximations then disturbances and modelling errors need to be accounted for.
+This naturally leads to Bayesian inference methods for subject-specific modelling to account for uncertainty.
 
 \brian{Is there a better observation to improve estimation? / experimental design}
-The BCRLB can improve experimental design, where the bound can be used to select what is important to measure. For example, 
+In addition to evaluating estimation feasability and performance, the BCRLB can improve experimental design.
+For example, the bound can be used to select the nodes in a network that are important to measure in order to g.
+For example,
 sampling rate affects the CRLB, extensions to other applications e.g. experimental design

### 2013-09-04 code/SetParametersWendling.m

@@ -1,4 +1,4 @@
-% Set parameters for the 6th order JR model
+% Set parameters for the Wendling
 %
 function params = SetParametersWendling(mode)
 

### 2013-09-04 code/main_Wendling.m

@@ -14,7 +14,7 @@ t = 0:dt:(N-1)*dt;
 % Intial true parameter values
 % mode = 'background';
 mode = 'alpha';       % this is more like alpha!!!
-% mode = 'spikes';
+mode = 'spikes';
 parameters = SetParametersWendling(mode);
 parameters.dt = dt;
 A = parameters.A;

### 2013-09-19 manuscript/manuscript.tex
Included plot of CRB vs time

@@
 
+\unit{Is the following paragraph needed? -}
 The model focused on in the following
@@
 
-To provide some intuition into the Bayes information matrix, we can write the joint distribution $p(\vec y,\vec x)$ as the product of the likelihood function and the prior distribution.
+To provide some intuition into the Bayes information matrix, we can write the joint distribution $p(\vec y,\vec x)$ as the product of the likelihood function, $\mathcal L$, and the prior distribution, $p_0$.
 Taking the logarithm gives
 \begin{align}
-	\log p(\vec y,\vec x) =& \log\mathcal{L}(\vec x|\vec y) + \log p(\vec x).
+	\log p(\vec y,\vec x) =& \log\mathcal{L}(\vec x|\vec y) + \log p_0(\vec x).
 \end{align}
@@
 	\mat J_D &= \mathbb E_{\vec x}\left[ \mat J_F(\vec x) \right]  \\
-	\mat J_P &= -\mathbb E_{\vec x}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec x) \right]
+	\mat J_P &= -\mathbb E_{\vec x}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p_0(\vec x) \right]
 \end{align}
@@
 First, the contribution of the data to the Bayes information matrix is the expectation of the classical Fisher information matrix taken over all possible states.
-Secondly, the contribution from the prior distribution directly adds to the Bayes information matrix, capturing our prior knowledge about of the states.
+Secondly, the contribution from the prior distribution increases the Bayes information matrix, capturing our prior knowledge about of the states.
 

### 2013-09-19 code/generateConvergenceResults.m

@@ -4,7 +4,7 @@ clc
 close all
 clear
 
-N = 500;             	% number of samples
+N = 100;             	% number of samples
 dT = 0.001;          	% sampling time step (global)
 dt = 1*dT;            	% integration time step
 nn = fix(dT/dt);      	% (used in for loop for forward modelling) the integration time step can be small that the sampling (fix round towards zero)
@@ -18,30 +18,38 @@ rng(0);
 M = 100;    % Mumber of Monte Carlo trials for MSE calculation
 
 
-NStatesModels = [6 6];  
+NStatesModels = [6 6 6];  
 
 % Model parameters
 params{1} = SetParametersLopes('alpha');
 params{2} = SetParametersLopes('alpha');
+params{3} = SetParametersLopes('alpha');
 
 
 % Transition functions
 models{1} = @model_Lopes;
 models{2} = @model_Lopes;
+models{3} = @model_Lopes;
 
 % Measurement fnctions
 Hmodels{1} = [1 0 -1 0 0 0];           % Lopes
 Hmodels{2} = [1 0 -1 0 0 0];           % Lopes
+Hmodels{3} = [1 0 -1 0 0 0];           % Lopes
 
 % Indices to plot (only potentials, not derivatives)
-displayStates{1} = [1 3 5];
-displayStates{2} = [1 3 5];
+displayStates{1} = [1 ];
+displayStates{2} = [1 ];
+displayStates{3} = [1 ];
 
-noiseInd = [2 2 ];     % NEED TO CHECK THESE! (and adjust Wendling noise)
-noiseScale = [1 1 ];
+noiseInd = [2 2 2];     % NEED TO CHECK THESE! (and adjust Wendling noise)
+noiseScale = [1 1 1];
 
-mseModels = cell(length(models),1);
-crbModels = cell(length(models),1);
+crbModels = zeros(6,N,length(models));
+
+priorSigma = [10 100 100];
+measureSigma = [10 10 100];
+
+plotStyles={'-','--',':'};
 
 for iModel = 1:length(models)
     
[…]

### 2013-09-19 code/plotConvergenceResults.m

@@ -9,35 +9,28 @@ fig_height = 7;  % cm
 
 LW = 0.5;         % linewidth
 
-fig_filename = '/Users/kelvin/Dropbox/CRLBNeuro/manuscript/figures/pdf/AlphaSeizureBar.pdf';
+fig_filename = '/Users/klayton/Dropbox/CRLBNeuro/manuscript/figures/pdf/DynamicCRB.pdf';
 figure('units','centimeters','papersize',[fig_height fig_width],'filename',fig_filename,'position',[2 2 fig_width fig_height]);
 
 load crb_convergence_results.mat
 
 %%
-crbBar = [];
-for iModel=[1 3 2 4]
-    crbBar = [crbBar; sum(sqrt(crbModels{iModel})); nan];
-end
+displayInd = 1;
 
-b1=bar(crbBar(:)','BaseValue',1e-6); hold on;
+crbData = squeeze(sqrt(crbModels(displayInd,2:end,:)));
+l1=plot(t(2:end),crbData,'-');
+
 % set(gca,'YScale','log','YGrid','on')
-set(gca,'YScale','linear','YGrid','on')
+% set(gca,'YScale','linear','YGrid','on')
 % xlim([0 20])
+grid on;
 
-ch1 = get(b1,'children');
-% cind = [1;1;1;1;2;2;2;2;2;2;3;3;3;3;4;4;4;4;4;4];
-cind = [1 1 2 2 3 3 4 4]';
-set(ch1,'FaceVertexCData',cind)
-cmap = [0.7556 0.8196 0.4769; 0.6590 0.4332 0.2696; 0.2245 0.3898 0.5888; 0.1831 0.9336 0.5528];
-% cmap = rand(4,3);
-colormap(cmap);
 
 laby=ylabel('RMSE (mV)');
 % set(gca,'XTick',[2 6 11 16],'XTickLabel',{'JR alpha','JR seizure','Wendling alpha','Wendling seizure'})
-set(gca,'XTick',[1 3 5 7],'XTickLabel',{'JR alpha','JR seizure','Wendling alpha','Wendling seizure'})
+% set(gca,'XTick',[1 3 5 7],'XTickLabel',{'JR alpha','JR seizure','Wendling alpha','Wendling seizure'})
 % ylim([1e-6 1e3])
-ylim([0 10])
+% ylim([0 10])
 
 set([gca laby],'fontsize',FS,'fontname',font)
 

### 2013-09-19 manuscript/manuscript.tex
Updated dynamic crb plot

@@
 
+\subsection{Time}
+
+Fig.~\ref{fig:DynamicCRB} displays the evolution of the Bayesian CRB for the first state of the model in \cite{Silva1974}.
+The solid line is for a relatively small prior covariance and low measurement noise; the dashed line is for a larger prior covariance and the dashed dotted line is for a large prior covariance and high measurement noise.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics{./figures/pdf/DynamicCRB}
+  \end{center}
+  \caption{The evolution of the BCRB for of the first state of model A. The ... }
+  \label{fig:DynamicCRB}
+\end{figure}
+
 \subsection{Across models - model feasibility}

### 2013-09-19 code/generateConvergenceResults.m

@@ -4,7 +4,7 @@ clc
 close all
 clear
 
-N = 100;             	% number of samples
+N = 110;             	% number of samples
 dT = 0.001;          	% sampling time step (global)
 dt = 1*dT;            	% integration time step
 nn = fix(dT/dt);      	% (used in for loop for forward modelling) the integration time step can be small that the sampling (fix round towards zero)
@@ -46,7 +46,7 @@ noiseScale = [1 1 1];
 
 crbModels = zeros(6,N,length(models));
 
-priorSigma = [10 100 100];
+priorSigma = [10 20 20];
 measureSigma = [10 10 100];
 
 plotStyles={'-','--',':'};
@@ -86,7 +86,7 @@ R = measureSigma(iModel)^2*eye(1);
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 % Compute CRLB
 % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-crbAll = compute_pcrb_P(t,f,F,Hfun,Q,R,m0,P0,100);
+crbAll = compute_pcrb_P(t,f,F,Hfun,Q,R,m0,P0,1000);
 
 
 crbModels(:,:,iModel) = crbAll;

### 2013-09-19 code/plotConvergenceResults.m

@@ -18,22 +18,28 @@ load crb_convergence_results.mat
 displayInd = 1;
 
 crbData = squeeze(sqrt(crbModels(displayInd,2:end,:)));
-l1=plot(t(2:end),crbData,'-');
+lHandle=plot(1e3*t(2:end),crbData);
 
-% set(gca,'YScale','log','YGrid','on')
-% set(gca,'YScale','linear','YGrid','on')
-% xlim([0 20])
-grid on;
+set(lHandle,'LineWidth',2)
+set(lHandle(1),'Color',[60 117 184]./255,'LineStyle','-');
+set(lHandle(2),'Color',[239 133 0]./255,'LineStyle','--');
+set(lHandle(3),'Color',[198 54 34]./255,'LineStyle','-.');
+set(gca,'Position',[0.1 0.17 0.8 0.8],'TickDir','out');
+xlim([0 80]);
+% grid on;
+box off
 
-
+labx=xlabel('Time (ms)');
 laby=ylabel('RMSE (mV)');
-% set(gca,'XTick',[2 6 11 16],'XTickLabel',{'JR alpha','JR seizure','Wendling alpha','Wendling seizure'})
-% set(gca,'XTick',[1 3 5 7],'XTickLabel',{'JR alpha','JR seizure','Wendling alpha','Wendling seizure'})
-% ylim([1e-6 1e3])
-% ylim([0 10])
+legText={};
+for i=1:length(priorSigma)
+legText = cat(1,legText,sprintf('$P_0=%d, R=%d$',priorSigma(i)^2,measureSigma(i)^2));
+end
+leg=legend(legText);
 
-set([gca laby],'fontsize',FS,'fontname',font)
-
+set([gca laby labx leg],'fontsize',FS,'fontname',font)
+set(leg,'interpreter','latex','Position',get(leg,'Position').*[0.95 1 1.1 1.1]);
+legend('boxoff')
 return
 %%
 export_fig(fig_filename)

### 2014-12-18 manuscript/manuscript.tex
edit to manuscript
edits to paper

@@
 \begin{document}
+\maketitle
 
@@
 Forward models have led to theory on the emergence of the brains rhythms, the mechanisms of anesthetic agents, the brains response to sensory stimuli, short-term memory formation, visual hallucinations and epileptic seizures.
-Most recently, advances engineering methods and computing resources have expanded the utility of these models as integral elements of frameworks to estimate of system and parametric states that can not be directly measured using electrophysiology.
+Most recently, advances engineering and statistical inference methods and computing resources have expanded the utility of these models as integral elements of frameworks to estimate system states that can not be directly measured using electrophysiology.
 An established framework to reliably estimate physical properties underlying brain dynamics has the potential to revolutionize how we observe, diagnose, treat and cure neurological disorders.
@@
 
-To date a large body of work has demonstrated validity of these models in-principal, through the ability to reproduce experimental data.
+To date a large body of work has attempted to demonstrate the validity of these models in-principal, through the ability to qualitatively reproduce experimental data.
 Since the parameters of the models are expected to vary across individuals and across disease states, the models must be tuned to individuals to establish clinical utility.
@@
 This paper introduces engineering tools to neuroscience community for formally accessing feasibility.
-We also postulate how these tools can be used to increase performance of estimation methods.
+We also postulate how these tools can be used to increase performance of estimation methods that are important for patient-specific modeling.
+Once models have been fit to subjects, a final step in assessing validity will be to examine the within subject predictive power.
+In order for such a framework to be established as a clinical platform further work is required for establishing validity and feasibility for the mesoscopic models.
 
-Once models have been fit to subjects a final step in assessing validity will be to examine the within subject predictive power.
-In order for such a framework to be established as a clinical platform further work is required for establishing validity and feasibility for the mesoscopic models.
+The literature describing mesoscopic neural models is dominated by neural field models and neural mass models. Models of this type
 
@@
 Neural mass models map from a mean firing rate of a population of pre-synaptic neurons to a mean membrane potential of a post-synaptic population, which in turn determines that population's output firing rate.
-A graphical description of a neural mass model can be seen in Figure~\ref{fig:SingleNeuralMass}.
+A graphical description of a single population neural mass model can be seen in Figure~\ref{fig:SingleNeuralMass}, showing the key elements including the post-synaptic response kernel and the sigmoidal activation function.
 
@@
 \end{align}
-The quantity $\varsigma_n$ describes the slope of the sigmoid (approximating the variance of firing thresholds within the populations) and $v_{0n}$ describes the mean firing threshold.
+The quantity $\varsigma_n$ describes the slope of the sigmoid (approximating the variance of firing thresholds within the populations) and $v_{0n}$ describes the mean firing threshold relative to a mean rest potential.
 An example of the sigmoidal activation function can be seen in the right inset of Figure~\ref{fig:SingleNeuralMass}.
@@
 EKF, particle filter) can be assessed against the lower bound.
-This indicated whether developing progressively advanced estimation algorithms is warranted, a question that has not been considered in the neural estimation literature thus far.
+The results from this third test will provide an indication on whether developing progressively advanced estimation algorithms is warranted, a question that has not been considered in the neural estimation literature thus far.
 

### 2014-12-18 manuscript/manuscript.tex
edits to intro and discussion

@@
 
-\title{Statistical performance analysis of data-driven neural mass models}
+\title{Statistical performance analysis of data-driven mesoscopic models}
 \date{\today}
@@
 
+\brian{What the paper is about / contribution}
+This paper introduces an existing engineering method to the neuroscience community for evaluating the feasibility of constructing subject-specific stochastic  computational models from measured data.
+Although the methods and models discussed in this paper are can be found in the engineering and computational neuroscience literature, it is important to link the fields together in order to advance our knowledge in systems neuroscience.
+
+\brian{Introduce the importance of the models, forward}
 Mesoscopic mathematical models of the brain are becoming increasingly popular as tools for generating hypotheses on the physical principals that govern neurodynamics.
-Forward models have led to theory on the emergence of the brains rhythms, the mechanisms of anesthetic agents, the brains response to sensory stimuli, short-term memory formation, visual hallucinations and epileptic seizures.
-Most recently, advances engineering and statistical inference methods and computing resources have expanded the utility of these models as integral elements of frameworks to estimate system states that can not be directly measured using electrophysiology.
+Mesoscopic models describe the pooled dynamics of neural populations, analogous to signals recorded from a local field potential.
+Typically the state variables of models at this scale are the average membrane potential or the average firing rates of the neurons across the population.
+Forward models of this kind have led to theory on the emergence of the brains rhythms, the mechanisms of anesthetic agents, the brains response to sensory stimuli, short-term memory formation, visual hallucinations and epileptic seizures.
+
+\brian{Existing lit on models} The literature describing mesoscopic neural models is dominated by neural field models and neural mass models. Models of this type
+
+\brian{Introduce the importance of the models, inverse}
+More recently, advances engineering and statistical inference methods and computing resources have expanded the utility of the mesoscopic models as integral elements of frameworks to estimate system states that can not be directly measured using electrophysiology.
 An established framework to reliably estimate physical properties underlying brain dynamics has the potential to revolutionize how we observe, diagnose, treat and cure neurological disorders.
@@
 
-To date a large body of work has attempted to demonstrate the validity of these models in-principal, through the ability to qualitatively reproduce experimental data.
+\brian{Models are qualitative - need to validate properly}
+To date a large body of work has attempted to demonstrate the validity of mesoscopic models in-principal, through the ability to qualitatively reproduce experimental data.
 Since the parameters of the models are expected to vary across individuals and across disease states, the models must be tuned to individuals to establish clinical utility.
@@
 
-The literature describing mesoscopic neural models is dominated by neural field models and neural mass models. Models of this type
+If model does not work then it is hard to say why.
+For example, the model itself could be too complicated or an estimation method could not be good enough.
+Is an approximate filter good enough etc etc.
 
@@
 \brian{Brief Summary of Paper}
-This paper has introduced a method for evaluating Bayesian estimation algorithms for fitting neural mass models to electrophysiological data.
-By providing bounds on the estimation accuracy, the results demonstrate that it is feasible to recover three neural mass models of varying complexity from data.
+This paper has introduced a method for evaluating Bayesian estimation algorithms for fitting neural models to electrophysiological data.
+The results demonstrate that it is feasible to recover three neural models of varying complexity from data, by providing bounds on the possible estimation accuracy.
 Under the assumption that these models capture the essential dynamics of the phenomena observed in intracranial electroencephalography, the results imply that is possible to tailor computational models to an individual's brain dynamics.
@@
 
-Neural mass models have evolved from a description of interaction of two very general populations, excitatory and inhibitory neurons, to interactions of specific populations that describe various networks.
+Mesoscopic neural models have evolved from a description of interaction of two very general populations, excitatory and inhibitory neurons, to interactions of specific populations that describe various networks.
 Cortical neural populations include pyramidal, spiny stellate, fast and slow inhibitory interneurons (ref).
 Deep brain populations include globus paladius, substansia niagara, sub-thalamic nuclei and more (ref).
-As models become more biologically realistic the BCRLB will be come more important for subject-specific modelling.
-Given that these models are all approximations then disturbances and modelling errors need to be accounted for.
-This naturally leads to Bayesian inference methods for subject-specific modelling to account for uncertainty.
+As models become more biologically realistic the BCRLB will be come more important for subject-specific modeling.
+Given that these models are simplified approximations of the brain, disturbances and modeling errors must be accounted for in the system equations.
+This naturally leads to Bayesian inference methods for subject-specific modeling to account for uncertainty.
 
@@
 we say nothing about how algorithms will work with uncertainty in the actual model -- thus can be considered a ``best case'' situation, where we know the true model exactly.
-Model section still needs to be done by checking how the models can predict the behavior of the system, check need to be checked experimentally.
+Model section still needs to be done by checking how the models can predict the behavior of the system. This need to be checked with test data set.
 

### 2015-01-22 manuscript/manuscript.tex
updated manuscript

@@
 \brian{What the paper is about / contribution}
-This paper introduces an existing engineering method to the neuroscience community for evaluating the feasibility of constructing subject-specific stochastic  computational models from measured data.
-Although the methods and models discussed in this paper are can be found in the engineering and computational neuroscience literature, it is important to link the fields together in order to advance our knowledge in systems neuroscience.
+This paper introduces the Bayesian Cramer Rao lower bound (BCRLB) to the neuroscience community.
+The BCRLB is an engineering method for evaluating the feasibility of constructing stochastic computational models from measured data.
+The lower bound defines the best possible outcome of estimating system states in terms of the levels of certainty.
+Although the methods and models discussed in this paper can be found dispersed in the engineering and computational neuroscience literature, respectively, it is important to link the fields together in order to advance our knowledge in systems neuroscience.
 
@@
 Mesoscopic models describe the pooled dynamics of neural populations, analogous to signals recorded from a local field potential.
-Typically the state variables of models at this scale are the average membrane potential or the average firing rates of the neurons across the population.
-Forward models of this kind have led to theory on the emergence of the brains rhythms, the mechanisms of anesthetic agents, the brains response to sensory stimuli, short-term memory formation, visual hallucinations and epileptic seizures.
-
-\brian{Existing lit on models} The literature describing mesoscopic neural models is dominated by neural field models and neural mass models. Models of this type
+Typically the state variables of models at this scale are the average membrane potential or the average firing rates of the neurons across the population (ref - Beurle, Wilson Cowan, Lopez Da Silva - van Rotterdam).
+More sophisticated models that describe higher-order statistics of neural populations have also been developed (Sompolinsky ?french guys?
+Jonathan Touboul).
+Forward models of this kind have led to theory on the emergence of the brains rhythms (Liley), the mechanisms of anesthetic agents (Liley - Dutch guys), the brains response to sensory stimuli (Jansen and Ritt), short-term memory formation (Coombes), visual hallucinations (Cowan Bressloff, Ermentrout) and epileptic seizures (Freestone, Robinson, Wendling, Grimbert?, Terry).
 
 \brian{Introduce the importance of the models, inverse}
-More recently, advances engineering and statistical inference methods and computing resources have expanded the utility of the mesoscopic models as integral elements of frameworks to estimate system states that can not be directly measured using electrophysiology.
+More recently, advances engineering and statistical inference methods along with computing resources have expanded the utility of the mesoscopic models as integral elements of frameworks to estimate system states that can not be directly measured using electrophysiology (Friston, Freestone, Schiff, Terry).
 An established framework to reliably estimate physical properties underlying brain dynamics has the potential to revolutionize how we observe, diagnose, treat and cure neurological disorders.
-For example, in epilepsy the dynamics of hidden system and parametric states are thought to be highly patient-specific and unknown.
-Thus the ability to estimate these quantities using the framework will enable a greater understanding of individuals pathologies, leading to more tailored therapies and better patient outcomes.
-Furthermore, subject-specific models will enable the application of control theory to the neurosciences, which is typically limited to man-made systems.
+For example, in epilepsy the dynamics of hidden system and parametric states are thought to be highly patient-specific and unknown (Freestone - Book Chapter).
+Thus, the ability to estimate these quantities will enable a greater understanding of individuals pathologies, leading to more tailored therapies and better patient outcomes.
+Furthermore, subject-specific models will enable the application of control theory to the neurosciences, which is typically limited to man-made systems (Schiff Book).
 
-\brian{Models are qualitative - need to validate properly}
-To date a large body of work has attempted to demonstrate the validity of mesoscopic models in-principal, through the ability to qualitatively reproduce experimental data.
-Since the parameters of the models are expected to vary across individuals and across disease states, the models must be tuned to individuals to establish clinical utility.
-A critical steps in establishing robust model fitting techniques is understanding the performance and feasibility of estimation algorithms.
-This paper introduces engineering tools to neuroscience community for formally accessing feasibility.
+\brian{Models are qualitative - need to validate properly - need to estimate parameters to do this}
+To date, a large body of work has attempted to demonstrate the validity of mesoscopic neural models by qualitatively reproducing measured data.
+For example, establishing a match between a generative model's spectral profile with an electrophysiological measurement has typically been used as validation.
+In reality, such a demonstration does little more for neuroscience then generate hypotheses that are currently difficult to test.
+Having said this, many beautiful mathematical results have emerged which capture our imagination, but do not \emph{prove} anything above brain function.
+The reason why this is the case is that the models have inherent uncertainty (unknown inputs) are parameters of these models are difficult (if not impossible) to measure.
+Therefore, parameters are often tuned to get the required output.
+In order to truly validate theory, useful predictions must be made.
+However, testing predictions against measured data has not been standard practice since the parameters of the models are expected to vary across individuals and across disease states, making the testing difficult.
+It is clear that a systematic framework for estimating system variables is required to truly validate mesoscopic neural theory.
+
+A critical step in establishing robust model fitting techniques is understanding the performance and feasibility of estimation algorithms.
+This paper introduces engineering tools to neuroscience community for formally accessing feasibility of estimating system variables.
 We also postulate how these tools can be used to increase performance of estimation methods that are important for patient-specific modeling.
+The methods introduced in this paper can also be used to improve experiments and maximize the information content of measurements.
+This paper does not deal with model selection.
 Once models have been fit to subjects, a final step in assessing validity will be to examine the within subject predictive power.
-In order for such a framework to be established as a clinical platform further work is required for establishing validity and feasibility for the mesoscopic models.
 
-If model does not work then it is hard to say why.
+\brian{Layout}
+The rest of the paper is arranged as follows.
+Section~\ref{sec:background} provides a description of the equations for a neural mass model.
+Then a derivation of the Bayesian Cramer Rao lower bound in Section~\ref{sec:Theory}, followed by a description of a recursion required for applications involving filtering methods for state estimation.
+Next we provide a series of simulation that show how the bound is effected by measurement noise, and the variance of prior information.
+This is followed by a demonstration of how models with different structures have varying bounds, indicating how well we can possibly estimate the system states.
+The final simulation provides a comparison between bounds when the parameters of the models are set to obtain seizures-like patterns and alpha rhythm-like patterns.
+In the discussion section we provide further intuition into how the Bayesian Cramer Rao lower bound can help lead to advances in our knowledge and the development of theory in unlocking the secrets of the brain.
+
+% If model does not work then it is hard to say why.
 For example, the model itself could be too complicated or an estimation method could not be good enough.
@@
   \begin{center}
-    \includegraphics{./figures/pdf/SingleMass_plos.pdf}
+    \includegraphics{./figures/pdf/SingleMass2.pdf}
   \end{center}
-  \caption{\emph{\textbf{Neural Mass Model.} The somas of the neural mass are represented by the triangle.
-The synaptic connections are represented by the circle.
-The input mean firing rate, $g(v_m(t))$, is convolved with post-synaptic response kernel (PSRK), shown in the inset on the left, to give the mean membrane potential, $v_n(t)$.
-This is transformed by the sigmoid, shown in the inset on the right, to give the output mean firing rate, $g(v_n(t))$.}}
+  \caption{\emph{\textbf{Neural Mass Model.} The soma of the neural mass is represented by the triangle and the synaptic connections are represented by the circles.
+The input mean firing rates, $g(v_{e}(t))$ and $g(v_{i}(t))$, are convolved with post-synaptic response kernels (PSRK), shown in the upper two insets, to give the excitatory and inhibitory post-synaptic membrane potentials (PSPs), $v_{e,n}(t)$ and $v_{i,n}(t)$, respectively.
+The PSPs are then summed and transformed by the sigmoid, shown in the lower inset, to give the output mean firing rate, $g(v_n(t))$.
+The neural mass may have any number of inputs and outputs.
+The firing rate on each of the outputs is equal.}}
   \label{fig:SingleNeuralMass}
 \end{figure}
-Neural mass models map from a mean firing rate of a population of pre-synaptic neurons to a mean membrane potential of a post-synaptic population, which in turn determines that population's output firing rate.
-A graphical description of a single population neural mass model can be seen in Figure~\ref{fig:SingleNeuralMass}, showing the key elements including the post-synaptic response kernel and the sigmoidal activation function.
+Neural mass models describe the mean firing rates and mean membrane potentials of populations of neurons.
+A graphical description of a single population neural mass model can be seen in Figure~\ref{fig:SingleNeuralMass}, showing the key elements including the post-synaptic response kernels and the sigmoidal activation function.
 
-To define a standard neural mass model, we begin by defining the post-synaptic potential of population $n$, $v_n(t)$, as a result of an input firing rate from population $m$, $g(v_m(t))$, as the convolution
+To derive a standard neural mass model, we begin by defining the mean membrane potential of population $n$, $v_n(t)$, as the \brian{(average)} soma as the sum of all $M$ incoming mean post-synaptic membrane potentials
+\begin{align}
+	v_n(t) = \sum_m^M v_{m,n}(t).
+\end{align}
+Each of the post-synaptic membrane potentials results from a convolution between the input firing rate from population $m$, $g(v_m(t))$, with a post-synaptic response kernel, which is defined by
 \begin{equation}\label{eq:conv_eq}
-    v_n(t) = \frac{\alpha_{mn}}{\tau_{mn}}\int_{-\infty}^t  h_{mn}(t-t')g(v_m(t')) \,\mathrm{d}t',
+    v_{m,n}(t) = \frac{\alpha_{m,n}}{\tau_{m,n}}\int_{-\infty}^t  h_{m,n}(t-t')\phi_m(t') \,\mathrm{d}t'.
 \end{equation}
-where $\alpha_{mn}$ is the gain for the post-synaptic response kernel (PSRK), denoted by $h_{mn}(t)$, from neural population $m$ to $n$, and $\tau_{mn}$ is the membrane time constant.
+The parameter $\alpha_{m,n}$ is the gain for the post-synaptic response kernel (PSRK), denoted by $h_{m,n}(t)$, from neural population $m$ to $n$, $\tau_{mn}$ is the membrane time constant, and $\phi(t)$ is the incoming mean firing rate.
 A common form of the PSRK is
@@
 
-The firing rate of each population is related to the mean membrane potential by a sigmoidal activation function; the logistic function sigmoid is typically used: 
+The input firing rate to each population, $\phi_m(t)$, may come from a population that is explicitly modeled or an external input where
 \begin{align}
-    g\left(v_n(t)\right) =& \frac{1}{1+\exp{\left(\varsigma_n\left(v_{0n} - v_n(t)\right)\right)}}.
+	\phi_{m} =& \left\{ \begin{array}{lll} u_m  & \text{if } m \text{ indexes external inputs}\\
+	g(v_m)  & \text{if } m \text{ indexes internal inputs} \end{array}\right. .
+\end{align}
+and $g(\cdot)$ is the sigmoidal activation function that relates the mean membrane potential the the mean firing rate of a population.
+The logistic function sigmoid is typically used and is described by
+\begin{align}
+    g\left(v_n(t)\right) =& \frac{1}{1+\exp{\left(\varsigma_n\left(v_{0} - v_n(t)\right)\right)}}.
     \label{eq:sigmoid}
 \end{align}
-The quantity $\varsigma_n$ describes the slope of the sigmoid (approximating the variance of firing thresholds within the populations) and $v_{0n}$ describes the mean firing threshold relative to a mean rest potential.
-An example of the sigmoidal activation function can be seen in the right inset of Figure~\ref{fig:SingleNeuralMass}.
+The quantity $\varsigma_n$ parametrizes the slope of the sigmoid (approximating the variance of firing thresholds within the populations) and $v_{0}$ describes the mean firing threshold relative to a mean rest potential.
+An example of the sigmoidal activation function can be seen in the lower inset of Figure~\ref{fig:SingleNeuralMass}.
 
@@
 \begin{equation}\label{eq:2ndOrder}
-    \mathrm{D}v_n(t) = \frac{\mathrm{d}^2 v_n(t)}{\mathrm{d}t^2} + \frac{2}{\tau_{mn}}\frac{\mathrm{d} v_n(t)}{\mathrm{d}t} + \frac{1}{\tau_{mn}^2} v_n(t) = \frac{\alpha_{mn}}{\tau_{mn}} g(v_m(t)),
+    \mathrm{D}v_n(t) = \frac{\mathrm{d}^2 v_n(t)}{\mathrm{d}t^2} + \frac{2}{\tau_{mn}}\frac{\mathrm{d} v_n(t)}{\mathrm{d}t} + \frac{1}{\tau_{mn}^2} v_n(t) = \frac{\alpha_{mn}}{\tau_{mn}} \phi_m(t),
 \end{equation}
-where $\mathrm{D}$ is a linear differential operator. Equation~\ref{eq:2ndOrder} can be written as two coupled first-order ODEs by 
+where $\mathrm{D}$ is a linear differential operator. The form a state-space model in a canonical format, equation~\ref{eq:2ndOrder} is written as two coupled first-order ODEs by 
 \begin{equation} \label{eq:2ndOrderNMM}
-    \frac{\mathrm{d} v_n(t)}{\mathrm{d}t} = z_n(t),\,\,\,\,\,    \frac{\mathrm{d}z_n(t)}{\mathrm{d}t} = \frac{\alpha_{mn}}{\tau_{mn}} g(v_m(t)) - \frac{2}{\tau_{mn}}z_n(t) - \frac{1}{\tau_{mn}^2} v_n(t),
+    \frac{\mathrm{d} v_n(t)}{\mathrm{d}t} = z_n(t),\,\,\,\,\,    \frac{\mathrm{d}z_n(t)}{\mathrm{d}t} = \frac{\alpha_{mn}}{\tau_{mn}} \phi_m(t) - \frac{2}{\tau_{mn}}z_n(t) - \frac{1}{\tau_{mn}^2} v_n(t).
 \end{equation}
-where $z_n(t)$ is a dummy variable. This forms the basis of a state-space model in a canonical format.
 
-The parameters of the model, $\tau$, $\alpha$, $v_0$ and $\varsigma$, can be set so the neural mass has characteristics of specific neural populations, such as pyramidal neurons, spiny stellate cells and fast and slow inhibitory interneurons (GABA$_\mathrm{a}$ and GABA$_\mathrm{b}$).
-The neural populations can then be connected to represent the circuitry of a cortical column and further to form networks of columns.
-Various kinds of neural mass models have been developed \cite{Silva1974,Jansen1995,Wendling2002,David2003}, which are depicted in Figure~\ref{fig:NMMs}.
-Each synaptic connection in these networks can be described by the 2$^{\mathrm{nd}}$-order system of Equation~\ref{eq:2ndOrderNMM}, where the resultant dimensions of the networks of neural mass models in Figures~\ref{fig:NMMs} a, b and c are 6, 12 and 18, respectively.
+The parameters of the model, $\tau$, $\alpha$, $v_0$ and $\varsigma$, can be set so the neural mass has characteristics of a specific neural population, such as pyramidal neurons, spiny stellate cells and fast and slow inhibitory interneurons (GABA$_\mathrm{a}$ and GABA$_\mathrm{b}$).
+The neural masses can then be interconnected to represent the circuitry of a cortical column and further to form networks of columns.
+A small subset of the various kinds of neural mass models that have been explored \cite{Silva1974,Jansen1995,Wendling2002,David2003} are depicted in Figure~\ref{fig:NMMs}.
+Each synaptic connection in these networks can be described by the 2$^{\mathrm{nd}}$-order system of Equation~\ref{eq:2ndOrderNMM}, where the resultant dimensions (before simplification) of the networks of neural mass models in Figures~\ref{fig:NMMs} A, B, and C are 6, 12 and 18, respectively.
 \begin{figure}[ht]
 	\centering
-		\includegraphics[scale=1]{./figures/pdf/NeuralMassesHoriz_plos.pdf}
-	\caption{\emph{\textbf{Models of Cortical Columns.} Each model has a pyramidal neural population, with various forms of feedback and inputs (Cx.
-= Cortex, Th.
-= Thalamus).
-The EEG is taken as the mean membrane potential of the pyramidal population, $v_p(t)$, with additive measurement noise.
+		\includegraphics[scale=1]{./figures/pdf/NeuralMassesHoriz_plos2.pdf}
+	\caption{\emph{\textbf{Models of Cortical Columns.} Each model has a pyramidal neural population, with various forms of feedback and inputs.
+Measurements from the model are derived from the mean membrane potential of the pyramidal population, $v_p(t)$, with additive noise.
 a) Minimal model of a column with inhibitory feedback \cite{Silva1974}.
 b) An extension with excitatory feedback and afferent input from thalamus \cite{Jansen1995,David2003}.
-c) Column with inhibition occurring across two times scales, enabling modelling of higher frequency activity observed in seizures~\cite{Wendling2002}.}}
+c) Column with inhibition occurring across two times scales, enabling modeling of higher frequency activity observed in seizures~\cite{Wendling2002}.}}
 	\label{fig:NMMs}
@@
 
-The parameters of the neural masses not only define the population type, but also the behaviour exhibited by the model.
-For example, certain parameter combinations result in a model of a cortical column that will generate alpha wave type activity (normal activity) and, for another set of parameters, we create a model that will exhibit epileptic behaviour \cite{Wendling2002}.
-Therefore, we consider a family of models, which we define generally as
+The higher dimensional models of cortical regions can be written in the form
+\begin{align}
+	\dot{\mathbf{x}} = \mathbf{A}\mathbf{x} + \mathbf{B} g\left(\mathbf{C}\mathbf{x}\right) + \mathbf{D}\mathbf{u},
+\end{align}
+where $\mathbf{x} \in \mathbb{R}^{N_x}$ is a state vector representing the postsynaptic membrane potentials generated by each population synapse and their time derivatives.
+There are two states per synapse and $N_x = 2N$ is the total number of states, where for $N$ synaptic connections in the models the state vector is of the form
+\begin{align}
+	\mathbf{x} = \begin{bmatrix} v_{1} & z_{1} & \hdots & v_{N} & z_{N}  \end{bmatrix}.
+\end{align}
+The matrix $\mathbf{A}$ encodes the dynamics induced by the membrane times and for $N$ synapses has the block diagonal structure
+\begin{align}
+	\mathbf{A} = \mathrm{diag}\left(\begin{array}{ccc}\boldsymbol{\Psi}_1 & \hdots & \boldsymbol{\Psi}_N\end{array}\right),
+\end{align}
+where
+\begin{align}
+	\boldsymbol{\Psi}_n = \begin{bmatrix} 0 & 1 \\ -\frac{1}{\tau_n^2} & -\frac{2}{\tau_n} \end{bmatrix}.
+\end{align}
+The matrix of synaptic gains from internal inputs, $\mathbf{B}$, has the diagonal form 
+\begin{align}
+	\mathbf{B} = \mathrm{diag}\left(\begin{array}{cccccccc} 0 & \alpha_1 & \hdots & 0 & \alpha_{N-K} & \hdots & 0 & 0 \end{array}\right),
+\end{align}
+where $K$ is the number of external inputs.
+The synaptic gains of post-synaptic potentials that are generated from external inputs are captured by the input gain matrix $\mathbf{D}$.
+The adjacency matrix, $\mathbf{C}$, defines the connectivity structure of the model.
+It is a matrix of zeros or ones that defines all the connections between the cell population types (excluding external inputs)
+\begin{align}
+	\mathbf{C} = \begin{bmatrix} 0 & 0 & \hdots & 0 & 0 \\
+	\mathbf{c}_{2,1} & 0 &  & \mathbf{c}_{2,2(N-K)-1} & 0 \\ 
+	\vdots & &  & & \vdots \\
+	0 & 0 & \ddots & 0 & 0 \\
+ 	\mathbf{c}_{2(N-K),1} & 0 &  & \mathbf{c}_{2(N-K),2(N-K)-1} & 0 \\
+	\vdots & & & & \vdots \\
+	0 & 0 & & 0 & 0 \\
+	0 & 0 & \hdots & 0 & 0 \end{bmatrix}. 
+\end{align}
+For example, if the PSPs from synapses 1 and 2 are summed and transformed by the sigmoid to give the input firing rate to synapse $n$, then row $2n$ of $\mathbf{C}$ with have the form
+\begin{align}
+	c_{2n,:} = \begin{bmatrix} 1 & 0 & 1 & 0 & 0 & 0 & \hdots & 0 & 0 \end{bmatrix}.
+\end{align}
+As mentioned above, $\mathbf{D}$ is a diagonal matrix of synaptic gains for connections from external sources where
+\begin{align}
+	\mathbf{D} = \mathrm{diag} \left(\begin{array}{cccccccc}0 & 0 & \hdots & 0 & \alpha_{N-K+1} & \hdots & 0 & \alpha_N\end{array}\right).
+\end{align}
+
+It is necessary to discretize the model to integrate the equations and run simulations. The discrete time version of the model is 
+\begin{align}
+	\dot{\mathbf{x}}_{t+\delta} = \mathbf{A}_{\delta}\mathbf{x}_t + \mathbf{B}_{\delta} g\left(\mathbf{C}\mathbf{x}_t\right) + \mathbf{D}_{\delta}\mathbf{u}_t,
+\end{align}
+where $\delta$ is the integration time step.
+The matrices $\mathbf{A}_{\delta}$, $\mathbf{B}_{\delta}$ and $\mathbf{D}_{\delta}$ are discrete time versions of $\mathbf{A}$, $\mathbf{B}$ and $\mathbf{D}$, respectively, and are defined in Appendix~\ref{app:discretization}.
+For ease of notation, we shall abbreviate increments or decrements in time (by the integration time step) by $t+1$ or $t-1$, respectively.
+For the discrete time version of the model, the input $\mathbf{u}_t$ may be stochastic.
+A discrete time Gaussian input $u_t\sim\mathcal{N}(\boldsymbol{\mu},\boldsymbol{\Sigma})$ represents a Wiener process (random walk) in the continuous time system.
+
+The neural mass model is mapped to electrophysiological measurements by the observation equation
+\begin{align}
+	\mathbf{y}_t =& \mathbf{H}\mathbf{x}_t + \mathbf{e}_t,
+\end{align}
+where $\mathbf{H}\in \mathbb{R}^{N_x \times N_y}$ is the observation matrix, $\mathbf{e}\in \mathbb{R}^{N_y}$ is the observation noise, and $N_y$ is the number of observations.
+In the neural mass modeling literature, the observation matrix, $\mathbf{H}$, typically acts to map the mean membrane potential of pyramidal populations to the electrophysiological measurements.
+This choice of observation is motivated by evidence that the surface potentials in intracranial electroencephalography are proportional to the total synaptic currents of pyramidal neurons (Nunez).
+The incoming PSPs to the pyramidal cells are proportional to the synaptic currents.
+
+To derive the BCRLB, it is useful to write the model in a more compact form as
 \begin{align}\label{eq:NeuralMassModel}
-    \dot{\vec{x}}(t) =& \vec f_{\theta}\left(\vec{x}(t),\vec{u}(t)\right)\\
-    y(t) =& \mat{C}\vec{x}(t) + e(t),
+    \mathbf{x}_t =& \vec f_{\theta}\left(\mathbf{x}_t,\mathbf{u}_t\right)\\
+    \mathbf{y}_t =& \mathbf{H}\mathbf{x}_t + \mathbf{e}_t,
 \end{align}
-where $\vec{x}(t) \in \mathbb{R}^{n_x}$ is a state vector representing the postsynaptic membrane potentials generated by each population synapse and their time derivatives, $n_x$ is the number of states and $\vec{u}(t)$ represents the system input, which may be from afferent connections or other brain regions, exogenous input, or account for model inaccuracies.
-The function $\vec f_{\theta}(\cdot)$ describes the dynamics, where $\theta \in \mathbb{R}^{n_{\theta}}$ determines the model type and the behaviour it exhibits.
-The iEEG is denoted by $y(t)$, $\vec{C}$ is the observation matrix, and $e(t)$ is the observation noise.
+where the function $\vec f_{\theta}(\cdot)$ describes system dynamics and the parameters $\boldsymbol{\theta} \in \mathbb{R}^{n_{\theta}}$ determines the model type and the behavior it exhibits. 
+We explicitly use the subscript $\theta$ because the parameters of the neural masses not only define the population type, but also the behavior exhibited by the model.
+For example, certain parameter combinations result in a model of a cortical column that will generate alpha wave type activity (normal activity) and, for another set of parameters, we create a model that will exhibit epileptic behavior \cite{Wendling2002}.
+Therefore, we consider these systems as family of models.
 
-\unit{Is the following paragraph needed? -}
-The model focused on in the following
-estimation sections is the formulation by Jansen and Rit \cite{Jansen-Rit-95}.
-Given space limitations we refer the reader to the article by Jansen
-and Rit\cite{Jansen-Rit-95} where the original state-space equations
-are presented. The parameters that were described in this section are related to the Jansen and Rit paper by
-\begin{align}
-    A =& \frac{\alpha_{pe}}{2e_0c_1} = \frac{\alpha_{pi}}{2e_0c_3} = \frac{\alpha_{ep}}{2e_0c_2} = \frac{\alpha_{xp}}{2e_0} \\
-    B =& \frac{\alpha_{ip}}{2e_0c_4} \\
-    a =& \frac{1}{\tau_{pe}} = \frac{1}{\tau_{pi}} = \frac{1}{\tau_{ep}} = \frac{1}{\tau_{xp}} \\
-    b =& \frac{1}{\tau_{ip}},
-\end{align}
-where $e_0$ is a parameter that scales the maximum firing rate, $A$ and $B$ are synaptic gains for excitation and inhibition respectively, and $a$ and $b$ are the reciprocals of the synaptic time constants for excitation and inhibition, respectively, and the subscripts $p$, $e$, $x$, and $i$ denote pyramidal ($p$), excitatory interneuron (spiny stellate) ($e$), external ($x$) or inhibitory interneuron ($i$) populations, respectively.
+% \unit{Is the following paragraph needed? -}
+% The model focused on in the following
+% estimation sections is the formulation by Jansen and Rit \cite{Jansen-Rit-95}.
+% Given space limitations we refer the reader to the article by Jansen
+% and Rit\cite{Jansen-Rit-95} where the original state-space equations
+% are presented. The parameters that were described in this section are related to the Jansen and Rit paper by
+% \begin{align}
+%     A =& \frac{\alpha_{pe}}{2e_0c_1} = \frac{\alpha_{pi}}{2e_0c_3} = \frac{\alpha_{ep}}{2e_0c_2} = \frac{\alpha_{xp}}{2e_0} \\
+%     B =& \frac{\alpha_{ip}}{2e_0c_4} \\
+%     a =& \frac{1}{\tau_{pe}} = \frac{1}{\tau_{pi}} = \frac{1}{\tau_{ep}} = \frac{1}{\tau_{xp}} \\
+%     b =& \frac{1}{\tau_{ip}},
+% \end{align}
+% where $e_0$ is a parameter that scales the maximum firing rate, $A$ and $B$ are synaptic gains for excitation and inhibition respectively, and $a$ and $b$ are the reciprocals of the synaptic time constants for excitation and inhibition, respectively, and the subscripts $p$, $e$, $x$, and $i$ denote pyramidal ($p$), excitatory interneuron (spiny stellate) ($e$), external ($x$) or inhibitory interneuron ($i$) populations, respectively.
 By making the assumption that all excitatory synapses share the same time constants and by defining the connectivity constants, $c_1$, $c_2$, $c_3$ and $c_4$, the network of neural masses is the JR NMM.
 
-\section{Theory}
-The Bayesian Cramer-Rao bound (BCRB) can be used to examine the inherent uncertainty in a neural mass model and assess the expected performance of different estimation algorithms.
-The bound was first formulated in \cite{VanTrees1968} for estimation of random parameters, making it well suited to the filtering problem described above when the states (membrane potentials and derivatives) are modelled as random variables.
-The BCRB is a lower bound on the mean squared error (MSE) of an estimator, $\hat{\vec x}(\vec y)$, given by
+\section{Theory}\label{sec:Theory}
+\subsection{Estimation Problem}
+The problem of estimating the distribution of the states given noisy measurements is well developed and described throughout the engineering and statistics literature.
+The problem is formalized by computing the conditional distribution $p\left(\mathbf{x}_t | \mathbf{y}_{1:T}\right)$, or in other words, finding the distribution of states given all observed data on the time interval $(0, T]$.
+Optimal inference can be achieved by applying two recursions.
+First a forward recursion using data up to time $t$,
 \begin{align}
-	\mathsf{MSE} = \mathbb E_{\vec x,\vec y} [(\hat{\vec x}(\vec y) - \vec x)(\hat{\vec x}(\vec y) - \vec x)^T] \ge \mat J^{-1},
+	p\left(\mathbf{x}_t | \mathbf{y}_{1:t}\right) \propto p(\mathbf{y}_t|\mathbf{x_t})\int p(\mathbf{x_t}|\mathbf{x_{t-1}})p(\mathbf{x}_{t-1}|\mathbf{y}_{1:t-1}) \, \mathrm{d}\mathbf{x}_{t-1}
+\end{align}
+for $t = 1,2,\hdots,T$. This is followed by a backward recursion to achieve target distribution
+\begin{align}
+	p\left(\mathbf{x}_t | \mathbf{y}_{1:T}\right) = p\left(\mathbf{x}_t | \mathbf{y}_{1:t}\right) \int \frac{p\left(\mathbf{x}_{t+1} | \mathbf{y}_{1:T}\right) p\left(\mathbf{x}_{t+1} | \mathbf{x}_{t}\right) }{\int p\left(\mathbf{x}_{t+1} | \mathbf{x}_{t}\right) p\left(\mathbf{x}_{t} | \mathbf{y}_{1:t}\right) \mathrm{d}\mathbf{x}_{t} } \mathrm{d}\mathbf{x}_{t+1},
+\end{align}
+for $t = T-1,T-2,\hdots,1,0$.
+If the dynamics $p(\mathbf{x}_t | \mathbf{x}_{t-1})$ and observations $p(\mathbf{y}_t| \mathbf{x_t})$ are linear and Gaussian, then the recursions can be solved exactly by the Kalman filter (Paninski 2010 plus refs from paper).
+
+The nonlinearities in the neural mass model lead to difficulties in solving the integrals, and furthermore, lead to non-Gaussian distributions even with Gaussian disturbances.
+Therefore, approximations are often used for propagating the state distribution through time $p(\mathbf{x}_t | \mathbf{x}_{t-1})$ (ref Freestone, Schiff).
+Two popular approximations are the unscented Kalman filter and the extended Kalman filter.
+The unscented Kalman filter approximates $p(\mathbf{x}_t | \mathbf{x}_{t-1})$ as a Gaussian and the EKF approximates the model $f(\mathbf{x},\mathbf{u})$ as linear, via a truncated Taylor series expansion.
+
+In many Bayesian inference methods, including Kalman filtering, the mean of the posterior distribution is taken as the state estimate
+\begin{align}
+	\hat{\mathbf{x}}\left(\mathbf{y}\right) = \mathbb{E}_\mathbf{x}\left[p\left(\mathbf{x}|\mathbf{y}\right)\right].
+\end{align}
+
+\subsection{Bayesian Cramer-Rao Lower Bound}
+The Bayesian Cramer-Rao bound (BCRB) can be used to examine the inherent uncertainty in a neural mass model and assess the expected (\brian{best possible?}) performance of estimation algorithms.
+The bound was first formulated in \cite{VanTrees1968} for estimation of random parameters, making it well suited to the filtering problem described above when the states (membrane potentials and derivatives) are modeled as random variables.
+The BCRB is a lower bound on the mean squared error (MSE) of state estimates (\brian{of a system?
+}), $\hat{\mathbf{x}}\left(\mathbf{y}\right)$, given by (\brian{do we need so show $\hat x$ as a function of $y$?})
+\begin{align}
+	\mathsf{MSE} &= \mathbb E_{\mathbf{x},\mathbf{y}} \left[\left(\hat{\mathbf{x}}\left(\mathbf{y}\right) - \mathbf{x}\right) \left(\hat{\mathbf{x}}\left(\mathbf{y}\right) - \mathbf{x}\right)^{\top}\right] \nonumber \\
+	& \ge \mathbf{J}^{-1},
 	\label{eqn:mse_bound}
 \end{align}
-where $\mat J$ is the Bayes information matrix defined as
+where $\mathbf{J}$ is the Bayes information matrix. The Bayes information matrix is defined as
 \begin{align}
-	\mat J &= -\mathbb E_{\vec x,\vec y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right],
+	\mathbf{J} &= -\mathbb E_{\mathbf{x},\mathbf{y}}\left[ \nabla_{\mathbf{x}}\nabla_{\mathbf{x}}^{\top} \log p(\mathbf{y},\mathbf{x}) \right],
 	\label{eqn:bayes_matrix}
 \end{align}
-$\nabla_{\vec x} = [\partial/\partial x_1,\ldots,\partial/\partial x_n]^T$, and the expectation is taken over both the state vector $\vec x$ and the measurements $\vec y$.
-In many Bayesian estimation schemes (including Kalman filtering) the estimator is the mean of the posterior distribution, $\hat{\vec x}(\vec y) = \mathbb{E}_{\vec x}\left[ p (\vec x|\vec y)\right]$.
+$\nabla_{\mathbf{x}} = [\partial/\partial x_1,\ldots,\partial/\partial x_n]^{\top}$, and the expectation is taken over both the state vector $\mathbf{x}$ and the measurements $\mathbf{y}$.
+%In many Bayesian estimation schemes (including Kalman filtering) the estimator is the mean of the posterior distribution, $\hat{\vec x}(\vec y) = \mathbb{E}_{\vec x}\left[ p (\vec x|\vec y)\right]$.
 
-To provide some intuition into the Bayes information matrix, we can write the joint distribution $p(\vec y,\vec x)$ as the product of the likelihood function, $\mathcal L$, and the prior distribution, $p_0$.
+To provide some intuition into the Bayes information matrix, we can write the joint distribution $p(\mathbf{y},\mathbf{x})$ as the product of the likelihood function, $\mathcal L$, and the prior distribution, $p_0(\mathbf{x})$.
 Taking the logarithm gives
 \begin{align}
-	\log p(\vec y,\vec x) =& \log\mathcal{L}(\vec x|\vec y) + \log p_0(\vec x).
+	\log p(\mathbf{y},\mathbf{x}) =& \log\mathcal{L}(\mathbf{x}|\mathbf{y}) + \log p_0(\mathbf{x}).
 \end{align}
-Using this relation, the Bayes information matrix in \eqref{eqn:bayes_matrix} can be decomposed into a data component, $\mat J_D$, and a prior component, $\mat J_P$, such that $\mat J = \mat J_D + \mat J_P$.
+Using this relation, the Bayes information matrix in \eqref{eqn:bayes_matrix} can be decomposed into a data component, $\mathbf{J}_D$, and a prior component, $\mathbf{J}_P$, such that $\mathbf{J} = \mathbf{J}_D + \mathbf{J}_P$.
 The two components are defined as
 \begin{align}
-	\mat J_D &= \mathbb E_{\vec x}\left[ \mat J_F(\vec x) \right]  \\
-	\mat J_P &= -\mathbb E_{\vec x}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p_0(\vec x) \right]
+	\mathbf{J}_D &= \mathbb E_{\mathbf{x}}\left[ \mathbf{J}_F(\mathbf{x}) \right]  \\
+	\mathbf{J}_P &= -\mathbb E_{\mathbf{x}}\left[ \nabla_{\mathbf{x}}\nabla_{\mathbf{x}}^{\top} \log p_0(\mathbf{x}) \right],
 \end{align}
-where $\mat J_F$ is the classical Fisher information matrix for deterministic states,
+where $\mathbf{J}_F$ is the classical Fisher information matrix for deterministic states,
 \begin{align}
@@
 \subsection{Recursive Bayesian CRB}
-To calculate a bound for filtering applications we are required to calculate $\mat J$ defined by \eqref{eqn:bayes_matrix} at each time, $k$.
-Fortunately, the structure of the joint density $p(\vec y,\vec x)$ allows for a recursive method to calculate $\mat J_k$ from $\mat J_{k-1}$, first demonstrated in \cite{Tichavsky1998}.
-The recursion presented in that work is not directly applicable to the neural field models described in Section~\ref{sec:background} since noise only enters the system through a single state and the resultant covariance matrix, $\mat Q$, is singular.
-An alternative recursion proposed in \cite{Bergman2001} is suitable to the models considered here and can be summarised as
+To calculate a bound for filtering applications we are required to calculate $\mathbf{J}$ defined by \eqref{eqn:bayes_matrix} at each time, $t$.
+Fortunately, the structure of the joint density $p(\mathbf{y},\mathbf{x})$ allows for a recursive method to calculate $\mathbf{J}_t$ from $\mathbf{J}_{t-1}$, first demonstrated in \cite{Tichavsky1998}.
+The recursion presented in that work is not directly applicable to the neural field models described in Section~\ref{sec:background}, since noise only enters the system through a single state and the resultant covariance matrix, $\mathbf{Q}$, is singular.
+An alternative recursion proposed in \cite{Bergman2001} is suitable to the models considered here and can be summarized as
 \begin{align}
-	\mat J_k = \left( \mat Q + {\mat F}_{k-1} \mat J_{k-1}^{-1} {\mat F}_{k-1}^T\right)^{-1} + \mat C^T \mat R^{-1} \mat C
+	\mathbf{J}_t = \left( \mathbf{Q} + {\mathbf{F}}_{t-1} \mathbf{J}_{t-1}^{-1} {\mathbf{F}}_{t-1}^{\top}\right)^{-1} + \mathbf{H}^{\top} \mathbf{R}^{-1} \mathbf{H}
 \end{align}
@@
 \begin{align}
-		{\mat F}_k &= \mathbb E_{\vec x_k} \left[ \nabla_{\vec x_k}^T \vec f(\vec x_k)\right] 
+		\mathbf{F}_t &= \mathbb E_{\mathbf{x}_t} \left[ \nabla_{\mathbf{x}_t}^{\top} \vec f(\mathbf{x}_t)\right] 
 		\label{eqn:F_matrix_expectation}
@@
 The expectation over the state $\vec x_k$ in \eqref{eqn:F_matrix_expectation} can rarely be computed in closed form for nonlinear transition functions so Monte Carlo techniques must be employed.
-A large number of state realisations are generated at each time point and the expectation is approximated by an average over the realisations.
+A large number of state realizations are generated at each time point and the expectation is approximated by an average over the realizations.
 
@@
 Secondly, different parameter sets can be tested within a single model to assess how the performance of an estimator is expected to change for different types of activity.
-Thirdly, the efficiency of a given estimation algorithm (e.g.
+Thirdly, the accuracy of a given estimation algorithm (e.g.
 EKF, particle filter) can be assessed against the lower bound.
@@
 
-\section{Simulations}
+\section{Simulations}\label{sec:Simulations}
 
@@
 It has been well demonstrated that models of type B and C can generate vastly different oscillations for different parameters \cite{sdf}.
-Bifurcation theory provides a theoretical tool to analyse this feature and has been reported in several places \cite{sdf}.
+Bifurcation theory provides a theoretical tool to analyze this feature and has been reported in several places \cite{sdf}.
 Good qualitative agreement between EEG data and the models has been observed and consequently different parameters sets have been assigned to different types of brain activity such as alpha rhythms or epileptic seizures \cite{sfd}.
@@
 
-Fig.~\ref{fig:AlphaSeizure} illustrates that the states underlying seizure type behaviour is more difficult to estimate than background alpha rhythms.
+Fig.~\ref{fig:AlphaSeizure} illustrates that the states underlying seizure type behavior is more difficult to estimate than background alpha rhythms.
 This is expected since the input noise driving the system is amplified by a larger synaptic gain parameter, $A$, for the seizure state, as listed in Table~\ref{tbl:Parameters}.
@@
 
+\brian{\section*{Dean's Notes}
+\subsection*{Definition of Sigmoid}
+The error function version of the sigmoid is 
+\begin{align}
+	g(x) &= \frac{1}{2}\left(\mathrm{erf}\left(\frac{x-v_{0}}{\sqrt{2}\varsigma}\right) + 1\right).
+\end{align}
+This sigmoid is the cdf of a Gaussian, where $v_0$ is the mean and $\varsigma$ is the variance.
+\subsection*{Transformation of GRV By erf}
+Let
+\begin{align}
+	x\sim\mathcal{N}\left(\mu,\sigma^2\right)
+\end{align}
+and
+\begin{align}
+	y =& g(x) \\
+	=& \frac{1}{2}\left(\mathrm{erf}\left(\frac{x-v_0}{\sqrt{2}\varsigma}\right) + 1\right).
+\end{align}
+Define some quantities to get transformed distribution:
+\begin{align}
+	x =& g^{-1}(y) \\
+	  =& v_0 + \sqrt{2}\varsigma\mathrm{erf}^{-1}\left(2y-1\right) \\
+	\frac{\mathrm{d}x}{\mathrm{d}y} =& \sqrt{2\pi}\varsigma\exp \left( \left[\mathrm{erf}^{-1}\left(2y-1\right)\right]^2\right)
+\end{align}
+The pdf of $y$ is
+\begin{align}
+	f_Y(y) =& \left|\frac{\mathrm{d}x}{\mathrm{d}y}\right|f_X(g^{-1}(y)) \\
+	=& \frac{\varsigma}{\sigma}\exp\left(\left[\mathrm{erf}^{-1}\left(2y-1\right)\right]^2\right)\exp\left(-\frac{\left(v_0 + \sqrt{2}\varsigma\mathrm{erf}^{-1}\left(2y-1\right) - \mu\right)^2}{2\sigma^2}\right)
+\end{align}
+for $0<y<1$. Now let
+\begin{align}
+	z = \mathrm{erf}^{-1}\left(2y-1\right)
+\end{align}
+giving
+\begin{align}
+	f_Y(y)=& \frac{\varsigma}{\sigma}\exp\left(z^2\right)\exp\left(-\frac{\left(v_0 + \sqrt{2}\varsigma z - \mu\right)^2}{2\sigma^2}\right) \\
+	=& \frac{\varsigma}{\sigma}\exp\left(-\frac{2\sigma^2 z^2 - \left(v_0 + \sqrt{2}\varsigma z - \mu\right)^2}{2\sigma^2}\right)
+\end{align}
+Get the numerator of the exponent in quadratic form
+\begin{align}
+	\mathrm{num} =& 2\left(\varsigma^2 - \sigma^2\right)z^2 + 2\sqrt{2}\varsigma\left(v_0-\mu\right)z + \mu^2 + v_0^2 - 2\mu v_0 \\
+	=& az^2 + bz + c \\
+	=& a\left(z-h\right)^2 + k
+\end{align}
+where
+\begin{align}
+	a =& 2\left(\varsigma^2 - \sigma^2\right) \nonumber \\
+	b =& 2\sqrt{2}\varsigma\left(v_0-\mu\right) \nonumber \\
+	c =& \mu^2 + v_0^2 - 2\mu v_0 \nonumber \\
+	k =& c - \frac{b^2}{4a}\nonumber \\
+	h =& - \frac{b}{2a}\nonumber
+\end{align}
+This gives
+\begin{align}
+	f_Y(y)=& \frac{\varsigma}{\sigma}\exp\left(\frac{k}{2\sigma^2}\right)\exp\left(-\frac{ (z - h)^2}{\frac{2\sigma^2}{a}}\right)
+\end{align}
+So we have a Gaussian like structure that may be exploited if we linearize $z$ (via first-order Maclaurin series expansion) such that
+\begin{align}
+	z =& \mathrm{erf}^{-1}\left(2y-1\right) \\
+	\approx& \frac{\sqrt{\pi}}{2}\left(2y-1\right)
+\end{align}
+This may give and interesting alternative for Gaussian approximation filter??
+\subsection*{Mode Derivation}
+The Gaussian transformed by the error function sigmoid has the pdf
+\begin{align}
+	f_X(x)=& a\exp\left(-\frac{ \left(\mathrm{erf}^{-1}\left(2x-1\right) - b\right)^2}{c}\right)
+\end{align}
+Now let's find the mode. To simplify, first we take $\ln$ of $f_X(x)$ (since this will not shift the max) giving
+\begin{align}
+	\ln f_X(x)=& \ln a - \frac{ (\mathrm{erf}^{-1}\left(2x-1\right) - b)^2}{c}
+\end{align}
+Now find max by taking derivative and setting to zero. To do this we define
+\begin{align}
+	v =& 2x-1 \nonumber \\
+	w =& \mathrm{erf}^{-1}(v) \nonumber \\
+	y =& w-b \nonumber \\
+	z =& y^2/c \nonumber
+\end{align}
+\begin{align}
+	\frac{\mathrm{d}}{\mathrm{d}x} \ln f_X(x) =& 0
+\end{align}
+where
+\begin{align}
+	\frac{\mathrm{d}z}{\mathrm{d}y} \frac{\mathrm{d}y}{\mathrm{d}w} \frac{\mathrm{d}w}{\mathrm{d}v} \frac{\mathrm{d}v}{\mathrm{d}x} =& \frac{2y}{c} \times 1 \times \frac{\sqrt{\pi}}{2}\exp\left[\left(\mathrm{erf}^{-1}(v)\right)^2\right] \times 2 \\
+	=& 0.
+\end{align}
+This is true where $y=0$. This implies the mode occurs at
+\begin{align}
+	\mathrm{erf}^{-1}(2x-1) = b.
+\end{align}
+From above
+\begin{align}
+	b = \frac{\sqrt{2}\varsigma\left(\mu-v_0\right)}{2\left(\varsigma^2-\sigma^2\right)}.
+\end{align}
+Therefore the mode is
+\begin{align}
+	\mathrm{mode}\left(f_X(x)\right) = \frac{1}{2}\left(\mathrm{erf}\left(\frac{\sqrt{2}\varsigma\left(\mu-v_0\right)}{2\left(\varsigma^2 - \sigma^2\right)}\right)+1\right)
+\end{align}
+This is correct and can be checked by code below.
+}
 \small

### 2015-01-23 manuscript/manuscript.tex
Some paper comments

@@
 
-\newcommand{\unit}[1]{\textcolor{green}{#1}}
+\newcommand{\unit}[1]{\textcolor{magenta}{#1}}
 \newcommand{\brian}[1]{\textcolor{blue}{#1}}
@@
 Although the methods and models discussed in this paper can be found dispersed in the engineering and computational neuroscience literature, respectively, it is important to link the fields together in order to advance our knowledge in systems neuroscience.
+\unit{this sentence is true, but I worry it downplays the contribution of the paper.
+presumably we will have some new conclusions?
+or maybe not}
 
@@
 Jonathan Touboul).
-Forward models of this kind have led to theory on the emergence of the brains rhythms (Liley), the mechanisms of anesthetic agents (Liley - Dutch guys), the brains response to sensory stimuli (Jansen and Ritt), short-term memory formation (Coombes), visual hallucinations (Cowan Bressloff, Ermentrout) and epileptic seizures (Freestone, Robinson, Wendling, Grimbert?, Terry).
+Forward models of this kind have led to theory on the emergence of the brains rhythms (Liley), the mechanisms of anesthetic agents (Liley - Dutch guys), the brain's response to sensory stimuli (Jansen and Ritt), short-term memory formation (Coombes), visual hallucinations (Cowan Bressloff, Ermentrout) and epileptic seizures (Freestone, Robinson, Wendling, Grimbert?, Terry).
 
 \brian{Introduce the importance of the models, inverse}
-More recently, advances engineering and statistical inference methods along with computing resources have expanded the utility of the mesoscopic models as integral elements of frameworks to estimate system states that can not be directly measured using electrophysiology (Friston, Freestone, Schiff, Terry).
+More recently, advances in engineering and statistical inference methods along with computing resources have expanded the utility of the mesoscopic models as integral elements of frameworks to estimate system states that can not be directly measured using electrophysiology (Friston, Freestone, Schiff, Terry).
 An established framework to reliably estimate physical properties underlying brain dynamics has the potential to revolutionize how we observe, diagnose, treat and cure neurological disorders.
 For example, in epilepsy the dynamics of hidden system and parametric states are thought to be highly patient-specific and unknown (Freestone - Book Chapter).
-Thus, the ability to estimate these quantities will enable a greater understanding of individuals pathologies, leading to more tailored therapies and better patient outcomes.
+Thus, the ability to estimate these quantities will enable a greater understanding of individual pathologies, leading to more tailored therapies and better patient outcomes.
 Furthermore, subject-specific models will enable the application of control theory to the neurosciences, which is typically limited to man-made systems (Schiff Book).
@@
 For example, establishing a match between a generative model's spectral profile with an electrophysiological measurement has typically been used as validation.
-In reality, such a demonstration does little more for neuroscience then generate hypotheses that are currently difficult to test.
-Having said this, many beautiful mathematical results have emerged which capture our imagination, but do not \emph{prove} anything above brain function.
-The reason why this is the case is that the models have inherent uncertainty (unknown inputs) are parameters of these models are difficult (if not impossible) to measure.
-Therefore, parameters are often tuned to get the required output.
+In reality, such a demonstration does little more for neuroscience than generate hypotheses that are currently difficult to test.
+Having said this, many beautiful mathematical results have emerged that capture our imagination, but do not \emph{prove} anything about brain function.
+The reason why this is the case is that the models have inherent uncertainty (unknown inputs) and parameters of these models are difficult (if not impossible) to measure.
+Therefore, parameters are often tuned to obtain the required output.
 In order to truly validate theory, useful predictions must be made.
@@
 A critical step in establishing robust model fitting techniques is understanding the performance and feasibility of estimation algorithms.
-This paper introduces engineering tools to neuroscience community for formally accessing feasibility of estimating system variables.
+This paper introduces engineering tools to the neuroscience community for formally accessing the feasibility of estimating system variables.
 We also postulate how these tools can be used to increase performance of estimation methods that are important for patient-specific modeling.
-The methods introduced in this paper can also be used to improve experiments and maximize the information content of measurements.
+\unit{how does the CRLB help increase algorithm performance?
+seems a bit of a stretch.} The methods introduced in this paper can also be used to improve experiments to maximize the information content of measurements.
 This paper does not deal with model selection.
-Once models have been fit to subjects, a final step in assessing validity will be to examine the within subject predictive power.
+Once models have been fit to subjects, a final step in assessing validity will be to examine the predictive power within a subject.
+\unit{would be good to have a reference on model selection here}
 
 \brian{Layout}
-The rest of the paper is arranged as follows.
+The remainder of the paper is arranged as follows.
 Section~\ref{sec:background} provides a description of the equations for a neural mass model.
-Then a derivation of the Bayesian Cramer Rao lower bound in Section~\ref{sec:Theory}, followed by a description of a recursion required for applications involving filtering methods for state estimation.
-Next we provide a series of simulation that show how the bound is effected by measurement noise, and the variance of prior information.
+A derivation of the Bayesian Cramer Rao lower bound is presented in Section~\ref{sec:Theory}, followed by a description of the recursion required for applications involving filtering methods for state estimation.
+Next we provide a series of simulations that show how the bound is effected by measurement noise, and the variance of prior information.
 This is followed by a demonstration of how models with different structures have varying bounds, indicating how well we can possibly estimate the system states.
@@
 \end{equation}
-where $\mathrm{D}$ is a linear differential operator. The form a state-space model in a canonical format, equation~\ref{eq:2ndOrder} is written as two coupled first-order ODEs by 
+where $\mathrm{D}$ is a linear differential operator. To form a state-space model in a canonical format, Equation~\ref{eq:2ndOrder} is written as two coupled first-order ODEs by 
 \begin{equation} \label{eq:2ndOrderNMM}
@@
 The parameters of the model, $\tau$, $\alpha$, $v_0$ and $\varsigma$, can be set so the neural mass has characteristics of a specific neural population, such as pyramidal neurons, spiny stellate cells and fast and slow inhibitory interneurons (GABA$_\mathrm{a}$ and GABA$_\mathrm{b}$).
-The neural masses can then be interconnected to represent the circuitry of a cortical column and further to form networks of columns.
+\unit{reference here} The neural masses can then be interconnected to represent the circuitry of a cortical column and further to form networks of columns.
 A small subset of the various kinds of neural mass models that have been explored \cite{Silva1974,Jansen1995,Wendling2002,David2003} are depicted in Figure~\ref{fig:NMMs}.
@@
 \begin{align}
-	\mathbf{x} = \begin{bmatrix} v_{1} & z_{1} & \hdots & v_{N} & z_{N}  \end{bmatrix}.
+	\mathbf{x} = \begin{bmatrix} v_{1} & z_{1} & \hdots & v_{N} & z_{N}  \end{bmatrix}\unit{^\top}.
 \end{align}
+\unit{I wonder if x should have time dependence explicit, e.g., x(t) in 8 and 9}
 The matrix $\mathbf{A}$ encodes the dynamics induced by the membrane times and for $N$ synapses has the block diagonal structure
@@
 
-It is necessary to discretize the model to integrate the equations and run simulations. The discrete time version of the model is 
+It is necessary to discretize the model to \unit{numerically} integrate the equations and run simulations. The discrete time version of the model is 
 \begin{align}
@@
 \end{align}
+\unit{what is $\dot x_{t+\delta}$? do you just mean $x_{t+\delta}$? or is the derivative a finite difference here?}
 where $\delta$ is the integration time step.
@@
 For the discrete time version of the model, the input $\mathbf{u}_t$ may be stochastic.
-A discrete time Gaussian input $u_t\sim\mathcal{N}(\boldsymbol{\mu},\boldsymbol{\Sigma})$ represents a Wiener process (random walk) in the continuous time system.
+A discrete time Gaussian input $\mathbf{u}_t\sim\mathcal{N}(\boldsymbol{\mu},\boldsymbol{\Sigma})$ represents a Wiener process (random walk) in the continuous time system.
 
@@
 \begin{align}\label{eq:NeuralMassModel}
-    \mathbf{x}_t =& \vec f_{\theta}\left(\mathbf{x}_t,\mathbf{u}_t\right)\\
+    \mathbf{x}_{t+1} =& \vec f_{\theta}\left(\mathbf{x}_t,\mathbf{u}_t\right)\\
     \mathbf{y}_t =& \mathbf{H}\mathbf{x}_t + \mathbf{e}_t,
 \end{align}
+\unit{I changed it to $t+1$ above}
 where the function $\vec f_{\theta}(\cdot)$ describes system dynamics and the parameters $\boldsymbol{\theta} \in \mathbb{R}^{n_{\theta}}$ determines the model type and the behavior it exhibits. 
@@
 \section{Theory}\label{sec:Theory}
+
 \subsection{Estimation Problem}
@@
 The problem is formalized by computing the conditional distribution $p\left(\mathbf{x}_t | \mathbf{y}_{1:T}\right)$, or in other words, finding the distribution of states given all observed data on the time interval $(0, T]$.
-Optimal inference can be achieved by applying two recursions.
+Optimal inference can be achieved by applying two recursions \unit{reference}.
 First a forward recursion using data up to time $t$,
@@
 If the dynamics $p(\mathbf{x}_t | \mathbf{x}_{t-1})$ and observations $p(\mathbf{y}_t| \mathbf{x_t})$ are linear and Gaussian, then the recursions can be solved exactly by the Kalman filter (Paninski 2010 plus refs from paper).
+\unit{where did you get these equations? i can't follow them. Also think it's better and more common to use $t$ as current time, $t-1$ as previous time (or $t+1$ and $t$), rather than introducing $T$}
 
@@
 The unscented Kalman filter approximates $p(\mathbf{x}_t | \mathbf{x}_{t-1})$ as a Gaussian and the EKF approximates the model $f(\mathbf{x},\mathbf{u})$ as linear, via a truncated Taylor series expansion.
+\unit{My understanding is that both UKF and EKF approximate $p(\mathbf{x}_t | \mathbf{x}_{t-1})$ as a Gaussian.
+They just different methods to define the moments of the Gaussian (unscented transform, linearisation)}
 
@@
 \end{align}
+\unit{add time index to this equation}
 
 \subsection{Bayesian Cramer-Rao Lower Bound}
-The Bayesian Cramer-Rao bound (BCRB) can be used to examine the inherent uncertainty in a neural mass model and assess the expected (\brian{best possible?}) performance of estimation algorithms.
+The Bayesian Cramer-Rao bound (BCRB) can be used to examine the inherent uncertainty in a neural mass model and assess the  performance of estimation algorithms against the best possible.
 The bound was first formulated in \cite{VanTrees1968} for estimation of random parameters, making it well suited to the filtering problem described above when the states (membrane potentials and derivatives) are modeled as random variables.
 The BCRB is a lower bound on the mean squared error (MSE) of state estimates (\brian{of a system?
-}), $\hat{\mathbf{x}}\left(\mathbf{y}\right)$, given by (\brian{do we need so show $\hat x$ as a function of $y$?})
+}\unit{if you like, but i prefer without}), $\hat{\mathbf{x}}\left(\mathbf{y}\right)$, given by (\brian{do we need so show $\hat x$ as a function of $y$?} \unit{I get it looks a bit clunky, but we need it for the expectation over $x$ AND $y$ to make sense below})
 \begin{align}
@@
 \begin{align}
-	\mat J_F(\vec x) &= -\mathbb E_{\vec y| \vec x}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y|\vec x) \right].
+	\mat J_F(\vec x) &= -\mathbb E_{\vec y| \vec x}\left[ \nabla_{\vec x}\nabla_{\vec x}^\top \log p(\vec y|\vec x) \right].
 \end{align}
@@
 
+\unit{maybe we should restate the BCRB explicitly in the filtering context, $\mathsf{MSE}(\hat x_t) \ge J^{-1}_t$ etc. It's a bit implied at the moment}
+
 The bound defined in \eqref{eqn:mse_bound} can be used in several ways.
@@
 Thirdly, the accuracy of a given estimation algorithm (e.g.
-EKF, particle filter) can be assessed against the lower bound.
+EKF, particle filter \cite{something}) can be assessed against the lower bound.
 The results from this third test will provide an indication on whether developing progressively advanced estimation algorithms is warranted, a question that has not been considered in the neural estimation literature thus far.
@@
 \begin{align}
-	\bar{\mat F}_k &= \mathbb E \left[ \nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
-	\bar{\mat R}_k^{-1} &= -\mathbb E\left[ \nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) \right] \\
-	\bar{\mat G}_k &= \mathbb E \left[ \nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
-	\mat Q_k^{-1} &= -\mathbb E \left[ \nabla_{\vec v_k}\nabla_{\vec v_k}^T \log p(\vec v_k) \right]
+	\bar{\mat F}_k &= \mathbb E \left[ \nabla_{\vec x_k}^\top \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
+	\bar{\mat R}_k^{-1} &= -\mathbb E\left[ \nabla_{\vec x_k}\nabla_{\vec x_k}^\top \log p(\vec y_k|\vec x_k) \right] \\
+	\bar{\mat G}_k &= \mathbb E \left[ \nabla_{\vec v_k}^\top \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
+	\mat Q_k^{-1} &= -\mathbb E \left[ \nabla_{\vec v_k}\nabla_{\vec v_k}^\top \log p(\vec v_k) \right]
 \end{align}
@@
 \begin{align}
-	\nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat F_k(\vec x_k) \\
-	\nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat I_{d\times d} \\
-	\nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) &= -\mat H^T\mat R^{-1} \mat H 
+	\nabla_{\vec x_k}^\top \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat F_k(\vec x_k) \\
+	\nabla_{\vec v_k}^\top \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat I_{d\times d} \\
+	\nabla_{\vec x_k}\nabla_{\vec x_k}^\top \log p(\vec y_k|\vec x_k) &= -\mat H^\top\mat R^{-1} \mat H 
 \end{align}
@@
 	\bar{\mat F}_k &= \mathbb E\left[ \mat F_k(\vec x_k)\right] \label{eqn:pcrb_term_F}\\
-	\bar{\mat R}_k^{-1} &= \mat H^T\mat R^{-1} \mat H \\
+	\bar{\mat R}_k^{-1} &= \mat H^\top\mat R^{-1} \mat H \\
 	\bar{\mat G}_k &= \mat I_{d\times d} 
@@
 \begin{align}
-	\mat J_{k|k} = \left( \mat Q + \bar{\mat F}_{k-1} \mat J_{k-1}^{-1} \bar{\mat F}_{k-1}^T\right)^{-1} + \mat H_k^T \mat R^{-1} \mat H_k
+	\mat J_{k|k} = \left( \mat Q + \bar{\mat F}_{k-1} \mat J_{k-1}^{-1} \bar{\mat F}_{k-1}^\top\right)^{-1} + \mat H_k^\top \mat R^{-1} \mat H_k
 \end{align}

### 2015-01-28 code/SetParametersJR.m
added figs

@@ -1,62 +1,0 @@
-% Set parameters for the 6th order JR model
-%
-function params = SetParametersJR(mode)
-
-if nargin<1
-    mode = 'alpha';
-end
-
-switch mode
-    case 'alpha'
-        
-        % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-        % this set is for alpha rhythm
-        % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-        params.mu = 220;        % this is just the mean when this function is used in estimation
-        params.sigma = 5.74;
-        
-        
-        % get current parameters
-        params.e_0 = 2.5;        % maximal firing parameter
-        params.v_0 = 6;        % firing threshold
-        params.r = 0.56;              % sigmoid slope
-        
-        params.A = 3.25;          % excitatory gain
-%         params.A = 5;          % excitatory gain
-        params.a = 100;          % excitatory time constant
-        params.B = 22;         % inhibitory gain
-        params.b = 50;         % inhibitory time constant
-        
-        % connectivity parameters from J and R model
-        C = 135; % 270;        % 135 for alpha activity
-        params.C1 = C;        % connectivity parameter - excitatory feedback PSP
-        params.C2 = 0.8*C;        % connectivity parameter - excitatory feedback firing
-        params.C3 = 0.25*C;        % connectivity parameter - inhibitory feedback PSP
-        params.C4 = 0.25*C;        % connectivity parameter - excitatory feedback firing
-        
-    case 'seizure'
-        % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-        % this set is for seizure
-        % ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-        params.mu = 90;        % this is just the mean when this function is used in estimation
-        params.sigma = 5.74;
-        
-        
-        % get current parameters
-        params.e_0 = 2.5;        % maximal firing parameter
-        params.v_0 = 6;        % firing threshold
-        params.r = 0.56;              % sigmoid slope
-        
-        params.A = 8.25;          % excitatory gain
-%         params.A = 5;          % excitatory gain
-        params.a = 100;          % excitatory time constant
-        params.B = 22;         % inhibitory gain
-        params.b = 50;         % inhibitory time constant
-        
-        % connectivity parameters from J and R model
[…]

### 2015-01-28 code/main_JR.m

@@ -1,170 +1,0 @@
-% This code repricates the structure of the Voss / Schiff FN simulation /
-% estimation
-
-N = 2000;             	% number of samples
-dT = 0.001;          	% sampling time step (global)
-dt = 1*dT;            	% integration time step
-nn = fix(dT/dt);      	% (used in for loop for forward modelling) the integration time step can be small that the sampling (fix round towards zero)
-
-t = 0:dt:(N-1)*dt;
-
-% Intial true parameter values
-mode='alpha';
-% mode='seizure';
-parameters = SetParametersJR(mode);
-parameters.dt = dt;
-A=parameters.A;
-a=parameters.a;
-sigma=parameters.sigma;
-
-% Initialise random number generator for repeatability
-%
-rng(0);
-
-% Transition model
-%
-NStates = 6;                           
-f = @(x)model_JR(x,'transition',parameters);
-F = @(x)model_JR(x,'jacobian',parameters);
-
-% Initialise trajectory state
-%
-x0 = zeros(NStates,1);                   % initial state
-x = zeros(NStates,N); 
-x(:,1) = x0;
-
-
-Q = zeros(NStates);
-Q(4,4) = (sqrt(dt)*A*a*sigma)^2;
-R = 1^2*eye(1);
-
-% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-% Generate trajectory
-% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-% Define input
-%
-v = mvnrnd(zeros(NStates,1),Q,N)';
-
-% Euler-Maruyama integration
-%
-for n=1:N-1
-    x(:,n+1) = f(x(:,n))  + v(:,n);
-end
-
-H = [0 0 1 0 -1 0];           % observation function
-w = mvnrnd(zeros(size(H,1),1),R,N)';
-y = H*x + w;
-figure,plot(t,-H*x)
-
-
-% save(['JR_' mode '.mat'],'y','t') 
-
-
-%%
-% One column = 17.35 max
-% two column = 8.3 cm max
-FS = 8;             % fontsize - point
-font = 'arial';
[…]

### 2015-01-28 code/model_JR.m

@@ -1,100 +1,0 @@
-% MODEL_JR This function implements the state space representation of
-% the nonlinear neural mass model equations.
-% 
-% Inputs: x - the current state
-%         params - a structure of parameters
-%         mode - 'transition' returns the new state,
-%                'jacobian' returns the Jacobian at the current state.
-%
-% Jan 2013
-%
-function out = model_JR(x,mode,params)
-
-dt = params.dt;         % time step
-mu = params.mu;        % this is just the mean when this function is used in estimation
-
-% Get current parameters
-%
-e_0 = params.e_0;        % maximal firing parameter
-v_0 = params.v_0;        % firing threshold
-r = params.r;              % sigmoid slope
-
-A = params.A;          % excitatory gain
-a = params.a;          % excitatory time constant
-B = params.B;         % inhibitory gain
-b = params.b;         % inhibitory time constant
-
-C1 = params.C1;        % connectivity parameter - excitatory feedback PSP
-C2 = params.C2;        % connectivity parameter - excitatory feedback firing
-C3 = params.C3;        % connectivity parameter - inhibitory feedback PSP
-C4 = params.C4;        % connectivity parameter - excitatory feedback firing
-
-
-% Linear component of model
-%
-F = [1, dt, 0, 0, 0, 0; ...
-    -a^2*dt, 1-2*a*dt, 0, 0, 0, 0; ...
-    0, 0, 1, dt, 0, 0; ...
-    0, 0, -a^2*dt, 1-2*a*dt, 0, 0; ...
-    0, 0, 0, 0, 1, dt; ...
-    0, 0, 0, 0, -b^2*dt, 1-2*b*dt];
-
-% Sigmoid functions
-%
-f_p = 1 / (1 + exp(r*(v_0 - (x(3)-x(5)))));   % pyramidal firing
-f_e = 1 / (1 + exp(r*(v_0 - C1*x(1))));       % excitatory feedback firing
-f_i = 1 / (1 + exp(r*(v_0 - C3*x(1))));       % inhibitory feedback firing
-
-if mode(1)=='t';
-    
-    % Nonlinear component
-    %
[…]

### 2015-01-28 code/model_JR_erf.m

@@ -1,106 +1,0 @@
-% MODEL_JR This function implements the state space representation of
-% the nonlinear neural mass model equations.
-% 
-% Inputs: x - the current state
-%         params - a structure of parameters
-%         mode - 'transition' returns the new state,
-%                'jacobian' returns the Jacobian at the current state.
-%
-% Jan 2013
-%
-function out = model_JR_erf(x,mode,params)
-
-dt = params.dt;         % time step
-mu = params.mu;        % this is just the mean when this function is used in estimation
-
-% Get current parameters
-%
-e_0 = params.e_0;        % maximal firing parameter
-v_0 = params.v_0;        % firing threshold
-r = params.r;              % sigmoid slope
-
-A = params.A;          % excitatory gain
-a = params.a;          % excitatory time constant
-B = params.B;         % inhibitory gain
-b = params.b;         % inhibitory time constant
-
-C1 = params.C1;        % connectivity parameter - excitatory feedback PSP
-C2 = params.C2;        % connectivity parameter - excitatory feedback firing
-C3 = params.C3;        % connectivity parameter - inhibitory feedback PSP
-C4 = params.C4;        % connectivity parameter - excitatory feedback firing
-
-
-% Linear component of model
-%
-F = [1, dt, 0, 0, 0, 0; ...
-    -a^2*dt, 1-2*a*dt, 0, 0, 0, 0; ...
-    0, 0, 1, dt, 0, 0; ...
-    0, 0, -a^2*dt, 1-2*a*dt, 0, 0; ...
-    0, 0, 0, 0, 1, dt; ...
-    0, 0, 0, 0, -b^2*dt, 1-2*b*dt];
-
-% Calculate the width of equivalent sigmoid using error function. See
-% test_sigmoid_erf.m function for optimisation results
-%
-rErf = 1.699/r;  
-
-% Sigmoid functions using error function
-%
[…]

### 2015-01-28 code/model_Lopes.m

@@ -43,8 +43,12 @@ F = [1, dt, 0, 0, 0, 0; ...
 
 % Sigmoid functions
 %
-f_v_i = 1 ./ (1 + exp(r*(v_0 - v_i)));      % firing rate of excitatory population
-f_v_e = 1 ./ (1 + exp(r*(v_0 - v_e)));      % firing rate of inhibitory population
+% f_v_i = 1 ./ (1 + exp(r*(v_0 - v_i)));      % firing rate of excitatory population
+% f_v_e = 1 ./ (1 + exp(r*(v_0 - v_e)));      % firing rate of inhibitory population
+
+varsigma = 1.699/r;
+f_v_e = 0.5*(erf((v_e - v_0)/(varsigma*sqrt(2))) + 1);
+f_v_i = 0.5*(erf((v_i - v_0)/(varsigma*sqrt(2))) + 1);
 
 
 if mode(1)=='t';
